### 2018-09-17 manuscript/manuscript.Rmd
started huyck project, loaded raw data

---
title: "A Comparative Ecophysiological Survey of Lycophytes, Ferns, and Seed Plants"
author: 
date: 
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')
```

# Abstract

/
Key Words:

# Introduction

Our understanding of the ecophysiology of ferns and lycophytes is limited. However, recent studies have shown that these two groups possess an unusual set of functional traits that likely influence their ecology (Brodribb, Watkins, Pinterman, Franks). For example, Watkins has demonstrated that the gametophyte generations of many tropical ferns are markedly desiccation-tolerant, with tolerance being closely linked to ecological preference. In seed plants, desiccation tolerance has been essentially lost from vegetative tissues. Additional work on ferns indicates a tight link between carbon and nitrogen relations with species distribution (Watkins 2005). Work on fern hydraulics is in its early stages; however, a study on tropical ferns suggests that ferns have a significantly more resistive vascular system than seed plants. A recent study on temperate species proposes that ferns are more similar to gymnosperms in their hydraulic properties (Watkins).

\

	A primary controlling factor in fern hydraulics is stomatal function. Recently, a series of studies have shown that stomatal control of ferns and lycophytes differs from seed plants. Early evidence on fern and lycophyte stomata indicates that these groups have inefficient stomatal systems. Ongoing work by Brodribb et al. is suggesting that ferns and lycophytes may possess a hydropassive mechanism which influences stomatal opening and closing. Unlike seed plants, fern and lycophyte stomata fail to close in response to abscisic acid (ABA) addition. This significant new discovery suggests that at least two fundamentally different mechanisms of stomatal function have evolved in early plants. This, in turn, may impact the differential distribution of ferns and lycophytes relative to seed plants.

	\

	Studies that examine the comparative physiology of ferns and lycophytes to seed plants are rare. Therefore, it is difficult to derive widely applicable hypotheses of the ecophysiological behavior of these groups. From the few studies we were able to find, ferns and lycophytes appear to have lower photosynthetic rates than seed plants. This would suggest that ferns may be at a competitive disadvantage when growing with seed plants. However, none of these studies have systematically compared the physiology of these groups when growing in similar habitats.

	\

	The goal of the current study is to examine the comparative physiology of a number of temperate ferns and lycophytes to angiosperms. We specifically examined several physiological parameters in a pairwise comparison between a fern or a lycophyte and an angiosperm in identical habitats. Given recent suggestions on inefficient stomatal functions, we hypothesized that ferns would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms. This, in turn, would impact aspects of leaf morphology (stomatal density), physiology (chlorophyll concentration), nutrient relations (N and P), and carbon relations. We speculate that the differences in physiological function may affect the ecology of these ferns, lycophytes, and angiosperms. 
#Methods

## Study site and species

This study was conducted at two sites in upstate New York. The data for most species were gathered at the Edmund Niles Huyck Preserve and Biological Research Station in Rensselaerville, New York.  This site is characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range. Data from the second site were gathered at a natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley. Forests at this site are younger, but consist of a similar species richness and composition as the first site.

\

We examined a series of fern, lycophyte, and angiosperm species at both sites (Table 1). The main focus of this work was to examine ferns relative to light-analogous angiosperms. To accomplish this, we would select a fern species and then locate the nearest neighbor angiosperm growing in similar habitat conditions. We limited our choice of angiosperms to those taxa that were herbaceous and similar in size and stature to the fern in question. At the Huyck Preserve, we sampled across two distinct habitats: understory low-light areas and open, high-light areas.

## Physiological Measurements

To generate comparative physiological data, we followed two procedures. The first was to generate a series of detailed light response curves for a select number of species. The second approach, to increase our comparative sample size, was to make several survey measurements of maximum photosynthetic and conductance rates for additional species; this step was only completed at the Huyck Preserve. All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (Lincoln, NE). For the light response curves, the system was set up to a flow rate of 300 units, and a CO2 concentration of 400 umol m-2 s-1. Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels. These measurements were used to generate light response curves using appropriate light levels for high-light and low-light species separately. The light levels were not identical for each species, with several points being added or subtracted with each run, but they generally followed the same template. For high light species this was light levels of 1500, 1000, 800, 600, 500, 400, 300, 250, 200, 150, 100, 50, 25, 15, 10,5 and 0 (umol m-2 s-1) . Low light curves followed this pattern; 500, 400, 350, 300, 250, 200, 150, 100, 50, 25, 15, 10, 5 and 0 (again umol m-2 s-1). Using the data we generated from the light response curves, we developed an understanding of the levels at which low-light and high-light species reached their maximum photosynthetic rates. These data were then used to complete our survey measurements. All high-light species reached maximum rates at 1200 umol m-2 s-1 and all low-light species reached maximum rates at 600. All other LiCor parameters were identical to those used for the light response curves.

## Foliar chemistry

Air-dried and pulverized tissues from each sample were analyzed for total C and N. Data for percent carbon and percent nitrogen were generated using a costech elemental analyzer blksjdf;jas. Phosphorus data were gathered using alsdjfsklj;

## Statistial analyses

# Results

## Light response curve parameters

In closed canopy environments, angiosperms had significantly higher respiration rates than ferns, which were equal to lycophytes (f2,44= 10.12, p=0.0003). In addition, photosynthetic quantum yield was higher in lycophytes than in angiosperms, whereas ferns had intermediate values between the two (f2,44=5.04, p=0.011). Light compensation points in ferns were equivalent to lycophytes, which both had higher points than angiosperms (f2,44=10.28, p=0.0002). In open canopy environments where lycophytes were lacking, ferns had significantly higher respiration rates (f1,18= 20.87, p= 0.0003), equivalent quantum yields (f1,18= 1.56, p=0.228), and significantly lower light compensation points(f1,18=12.83, p=0.0023).

\

For angiosperms, species growing in open habitats had significantly more negative respiration rates (t22=-6.88, p<0.0001), equivalent quantum yields(t22=1.89, p=0.072), and higher light compensation points (t22=5.23, p<0.0001) than those angiosperms growing in closed habitats.

## Physiological surveys

Ferns growing in open environments had significantly higher maximum area-based photosynthetic rates and stomatal conductances and lower internal leaf CO2 concentrations than those growing in closed canopies (Figure dfg). Angiosperms growing in open canopy environments also had significantly higher maximum area-based photosynthetic rates and stomatal conductances than those growing in closed environments, but there was no significant difference in internal leaf CO2 concentrations between the two light environments.

\

In open canopy environments, angiosperms, on average, had significantly higher maximum photosynthetic rates and stomatal conductances than ferns growing in similar light conditions. There was no significant difference between the internal CO2 concentrations of angiosperm leaves and fern fronds; however, the statistics hint at a strong trend (Figure asldf).

\

Across closed canopy environments where lycophytes, some of these differences between angiosperms and ferns break down. For example, there is no significant difference between the photosynthetic rates of angiosperms and ferns, although these two groups are significantly different than lycophytes. This relationship is also shown in internal CO2 concentrations, with angiosperms and ferns showing no significant difference and lycophytes differing significantly from both groups. For stomatal conductances, however, all three groups are significantly different (Figure adfd).

# Nutrient relations

In open canopy environments, ferns had significantly lower phosphorus concentrations than those growing in closed canopies. Angiosperms had no significant difference across the light gradient. In both environments, ferns and lycophytes do not differ significantly, but both have significantly lower phosphorus levels than angiosperms (Figure asdf). Ferns in open light environments have significantly lower nitrogen levels than those ferns growing in closed light environments; the same is true of angiosperms. Lycophytes have significantly lower nitrogen levels than ferns and angiosperms, which do not statistically differ, in both light conditions (Figure adf).

\

	Neither ferns nor angiosperms differ within their classifications significantly across the two light environments in their nitrogen to phosphorus ratio (N:P). In open light environments, ferns have significantly higher N:P ratios than angiosperms. In closed canopy environments, ferns do not differ from angiosperms significantly in their N:P ratios, but do differ from lycophytes. Also, lycophytes and angiosperms do not differ significantly.

\

	Ferns in open light canopies have significantly higher carbon to nitrogen ratios (C:N) than those growing in closed canopies, while angiosperms do not differ across the light gradient. Also, ferns have significantly higher C:N ratios than corresponding angiosperms in open light environments. In closed canopies, lycophytes have the highest C:N ratios and differ significantly from both ferns and angiosperms, which do not significantly differ.

## Physiological parameters

In open canopy habitats, ferns had significantly lower stomatal densities than those growing in closed canopies, whereas angiosperms showed no significant differences in stomatal densities between the two habitats. In both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms (Figure asdfs). Similarly, ferns displayed a significantly lower SPAD index, a measurement of chlorophyll concentration, in open canopy environments versus closed canopies. Angiosperms showed no significant difference across the light gradient. In open environments, ferns had significantly lower SPAD index values than corresponding angiosperms, while the reverse was true in closed canopy habitats, with ferns displaying significantly higher values than analogous angiosperms (Figure asdfkj).

### 2018-09-19 notes.R
species id's are finished.  ready for plotting

@@ -1,0 +1,8 @@
+
+#1) Figure of light response by group / habitat
+#2) max A
+#3) Group chlorophyll
+#4) group stomatal density
+
+
+#bivariate

### 2018-11-12 manuscript/manuscript.Rmd
added new plots for manuscript

@@
 All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (Lincoln, NE).
-For the light response curves, the system was set up to a flow rate of 300 units, and a CO2 concentration of 400 umol m-2 s-1.
+For the light response curves, the system was set up to a flow rate of 300 units, and a CO~2~ concentration of 400 umol m^-2 s^-1.
 Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels.
@@
 The light levels were not identical for each species, with several points being added or subtracted with each run, but they generally followed the same template.
-For high light species this was light levels of 1500, 1000, 800, 600, 500, 400, 300, 250, 200, 150, 100, 50, 25, 15, 10,5 and 0 (umol m-2 s-1) .
+For high light species this was light levels of 1500, 1000, 800, 600, 500, 400, 300, 250, 200, 150, 100, 50, 25, 15, 10,5 and 0 (umol m-2 s-1).
 Low light curves followed this pattern; 500, 400, 350, 300, 250, 200, 150, 100, 50, 25, 15, 10, 5 and 0 (again umol m-2 s-1).
 Using the data we generated from the light response curves, we developed an understanding of the levels at which low-light and high-light species reached their maximum photosynthetic rates.
-These data were then used to complete our survey measurements.
+These data were then used to complete our light saturated gas exchange survey measurements.
 All high-light species reached maximum rates at 1200 umol m-2 s-1 and all low-light species reached maximum rates at 600.
-All other LiCor parameters were identical to those used for the light response curves.
+All cuvette parameters were identical to those used for the light response curves.
 
@@
 
-Air-dried and pulverized tissues from each sample were analyzed for total C and N.
-Data for percent carbon and percent nitrogen were generated using a costech elemental analyzer blksjdf;jas.
-Phosphorus data were gathered using alsdjfsklj;
+Air-dried and ground tissues from each leaf sample used for gas exchange were analyzed for total C and N.
+Foliar carbon and nitrogen content were generated using a costech elemental analyzer (machine).
+Phosphorus data were gathered using some digest (method)
+
+## Stomatal anatomy
 

### 2018-11-12 AQ_plot.R

@@ -1,0 +1,6 @@
+#plot LRC and analyze differences
+
+lrc <- read.csv("calculated_data/AQ_params.csv")
+
+
+boxplot

### 2018-11-12 functions/light_response_function.R

@@ -1,0 +1,269 @@
+##light response function
+
+fit_AQ_curve <- function(df, group_id, Photo, PARi, fit_type = "onls"){
+  AQ_curve_fits <- data.frame(ID = character(),
+                              Asat = numeric(),
+                              Phi = numeric(),
+                              Rd = numeric(),
+                              theta = numeric(),
+                              resid_SSs = numeric(),
+                              LCP = numeric(),
+                              Q_sat_75 = numeric(),
+                              Q_sat_85 = numeric(),  
+                              stringsAsFactors = FALSE
+  )
+  if(fit_type == "onls"){
+    if(require("onls")){
+      print("onls is loaded correctly")
+    } else {
+      print("trying to install onls")
+      install.packages("onls")
+      if(require("onls")){
+        print("onls installed and loaded")
+      } else {
+        stop("could not install onls")
+      }
+    }
+    library("onls")      
+    for(i in seq_along(unique(df[[group_id]]))){
+      tryCatch({
+        AQ_curve_fits[i, 1] <- unique(df[[group_id]])[i]
+        # Subset by group_ID iteratively:
+        single_curve1 <- df[df[[group_id]] == unique(df[[group_id]])[i],]
+        single_curve1$assim <- single_curve1[[Photo]]
+        single_curve1$PAR <- single_curve1[[PARi]]
+        single_curve = single_curve1[order(single_curve1$PAR),]
+        phi.as.slope <- with(single_curve,
+                             as.numeric(coef(lm(
+                               assim[1:5] ~ PAR[1:5]))[2]))
+        # Fit the curve:
+        temp.fit <- with(single_curve, # use the subset of a single curve
+                         onls(assim ~ ((Phi * PAR + Asat - 
+                                          sqrt((Phi * PAR + Asat)^2 - 
+                                                 4 * Phi * theta * 
+                                                 Asat * PAR ))
+                         )/(2*theta) - Rd,
+                         start=list(
+                           Asat = (max(assim)),
+                           Phi = phi.as.slope,
[…]

### 2018-11-12 master_scripts/cholrophyll.R

@@ -1,0 +1,20 @@
+##chlorophyll content
+
+## need some curvilinear correction for SPAD to CHL
+
+spad <- read.csv("raw_data/chlorophyll.csv")
+
+
+##boxplots (habitat overall)
+
+boxplot(spad ~ plant_group, data=spad, outline = FALSE)
+
+## groups by habitat
+boxplot(spad ~ canopy, data=spad, outline = FALSE)
+
+##by group x habitat
+spad$uniqueid<- paste(spad$plant_group, spad$canopy, sep="-")
+
+boxplot(spad ~ uniqueid, data=spad, outline = FALSE,xaxt='n',at=c(1:2, 4:5))
+axis(1, labels = c("Closed", "Open", "Closed", "Open"), at=c(1:2, 4:5))
+mtext(text = c("Angiosperms","Ferns"), side=1, line=2.5,at=c(1.5, 4.5))

### 2018-11-12 master_scripts/lightresponse.R

@@ -1,56 +1,0 @@
-##light response
-
-lrc <- read.csv("raw_data/lightresponse.csv")
-
-species <- read.csv("raw_data/species_list.csv")
-
-lrc2 <- merge(lrc, species, all=TRUE)
-lightresponse <- lrc2[complete.cases(lrc2$individual),]
-
-
-library(ggplot2)
-angio <- droplevels(lightresponse[lightresponse$plant_group == 'Angiosperm',])
-open <- droplevels(lightresponse[lightresponse$canopy == "Open",])
-closed <- droplevels(lightresponse[lightresponse$canopy == "Closed",])
-
-
-ggplot(data=lightresponse) +
- geom_smooth(mapping = aes(x=PARi, y=Photo, color=plant_group)) +
-   geom_point(mapping = aes(x=PARi, y=Photo, color=plant_group))
-
-
-
-##appears to be some big differences in angiosperm, is it canopy?
-ggplot(data=angio) +
-  # geom_smooth(mapping = aes(x=PARi, y=Photo, linetype=species)) +
-  geom_point(mapping = aes(x=PARi, y=Photo, color=species)) + theme(plot.subtitle = element_text(vjust = 1), 
-    plot.caption = element_text(vjust = 1), 
-    axis.line = element_line(size = 0.3, 
-        linetype = "solid"), panel.grid.minor = element_line( linetype = "blank"), 
-        panel.background = element_rect(fill = NA, 
-        linetype = "solid")) + theme(panel.grid.minor = element_line(colour = NA))
-
-##looks like just one species
-
-
-##differences within canopy types
-ggplot(data=open) +
-  geom_smooth(mapping = aes(x=PARi, y=Photo, color=plant_group)) +
-  geom_point(mapping = aes(x=PARi, y=Photo, shape=species, color=plant_group))
-
-ggplot(data=closed) +
-  geom_smooth(mapping = aes(x=PARi, y=Photo, color=plant_group)) +
-  geom_point(mapping = aes(x=PARi, y=Photo, shape=species, color=plant_group))
-
-ggplot(data=closed, aes(PARi, Photo)) +
-  geom_smooth(mapping=aes(color=plant_group), span=1) 
-  geom_point(mapping = aes(x=PARi, y=Photo, shape=species, color=plant_group))
-  
-ggplot(data=closed, aes(PARi, Photo)) +
-  geom_point(mapping = aes(color=plant_group))
-
-
-
-
-
-

### 2018-11-12 master_scripts/photosynthesis.R

@@ -22,11 +22,11 @@ closed <- photo[photo$canopy == "Closed", "Photo"]
 par(mfrow=c(1,2))
 ##plant group
 par(mar=c(5,5,1,0))
-plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,25), xaxt='n', ylab="",
+plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,25), xaxt='n', ylab="Photosynthesis",
      xlab="")
 vioplot(lyco, fern, angio,at=1:3 ,add=TRUE,
         col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
-axis(2, labels=FALSE, tcl=.25)
+# axis(2, labels=FALSE, tcl=.25)
 axis(1, labels = c("lycophytes","ferns","angio"), at=1:3)
 
 ##habitat
@@ -37,6 +37,19 @@ vioplot(open, closed,at=1:2 ,add=TRUE,
         col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
 axis(2, labels=FALSE, tcl=.25)
 axis(1, labels = c("open","closed"), at=1:2)
+
+##by group x habitat
+photo$uniqueid<- paste(photo$plant_group, photo$canopy, sep="-")
+
+par(mar=c(5,5,1,1))
+boxplot(Photo ~ uniqueid, data=photo, outline = FALSE,xaxt='n',at=c(1:2, 4:5, 7),
+        ylab="Photosynthesis")
+axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5, 7))
+mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, line=2.5,
+      at=c(1.5, 4.5, 7))
+
+
+
 
 ##using ggplot
 # p <- ggplot(photo, aes(x=plant_group, y=Photo, color=bla)) + 

### 2018-11-12 master_scripts/stomata_density.R

@@ -1,8 +1,10 @@
 ## stomatal density by plant type & habitat
 
 stom <- read.csv("raw_data/stomatal_density.csv")
+treatments <- read.csv("raw_data/species_list.csv")
 library(vioplot)
 
+stom2 <- merge(stom, treatments[,4:5])
 
 angio_sd <- stom[stom$plant_group == "Angio", "sto_den"]
 fern_sd <- stom[stom$plant_group == "Fern", "sto_den"]
@@ -17,4 +19,27 @@ axis(2, labels=FALSE, tcl=.25)
 axis(1, labels = c("lycophytes","ferns","angiosperms"), at=1:3)
 
 
-##need to add habitat to this dataset
+## boxplots (overall)
+par(mar=c(5,5,1,1))
+boxplot(sto_den ~ plant_group, data=stom, outline = FALSE, xaxt='n')
+axis(1, labels = c("Angiosperms","Ferns","Lycophytes"), at=1:3)
+
+##boxplots (habitat overall)
+
+boxplot(sto_den ~ plant_group, data=stom2, outline = FALSE, xaxt='n')
+axis(1, labels = c("Angiosperms","Ferns","Lycophytes"), at=1:3)
+
+## groups by habitat
+boxplot(sto_den ~ canopy, data=stom2, outline = FALSE)
+
+##by group x habitat
+stom2$uniqueid<- paste(stom2$plant_group, stom2$canopy, sep="-")
+id_labels <- c("Angiosperm\nClosed", "Angiosperm\nOpen", "Fern\nClosed",
+               "Fern\nOpen", "Lycophyte\nClosed")
+
+boxplot(sto_den ~ uniqueid, data=stom2, outline = FALSE, xaxt='n',at=c(1:2, 4:5, 7))
+axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5, 7))
+mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, line=2.5,
+      at=c(1.5, 4.5, 7))
+
+

### 2018-11-12 scripts/lightresponse.R

@@ -1,0 +1,72 @@
+##light response
+
+lrc <- read.csv("raw_data/lightresponse.csv")
+lrc$id <- paste(lrc$species, lrc$individual, sep="-")
+
+source("light_response_function.R")
+
+##fit light response curves for every individual
+
+hyuck_lrc <- fit_AQ_curve(lrc, 
+      Photo = "Photo", 
+      PARi = "PARi", 
+      group_id = "id") 
+
+diagnostic_AQ_plot(lrc, hyuck_lrc,
+                   Photo = "Photo", 
+                   PARi = "PARi", 
+                   group_id = "id",
+                   save_to_pdf = TRUE,
+                   save_path = "output",
+                   file_name = "LRC_hyuck_plots")
+
+write.csv(hyuck_lrc, "calculated_data/AQ_params.csv", row.names = FALSE)
+
+##trying to learn ggplots
+
+# library(ggplot2)
+# angio <- droplevels(lightresponse[lightresponse$plant_group == 'Angiosperm',])
+# open <- droplevels(lightresponse[lightresponse$canopy == "Open",])
+# closed <- droplevels(lightresponse[lightresponse$canopy == "Closed",])
+# 
+# 
+# ggplot(data=lightresponse) +
+#  geom_smooth(mapping = aes(x=PARi, y=Photo, color=plant_group)) +
+#    geom_point(mapping = aes(x=PARi, y=Photo, color=plant_group))
+# 
+# 
+# 
+# ##appears to be some big differences in angiosperm, is it canopy?
+# ggplot(data=angio) +
+#   # geom_smooth(mapping = aes(x=PARi, y=Photo, linetype=species)) +
+#   geom_point(mapping = aes(x=PARi, y=Photo, color=species)) + theme(plot.subtitle = element_text(vjust = 1), 
+#     plot.caption = element_text(vjust = 1), 
+#     axis.line = element_line(size = 0.3, 
+#         linetype = "solid"), panel.grid.minor = element_line( linetype = "blank"), 
+#         panel.background = element_rect(fill = NA, 
+#         linetype = "solid")) + theme(panel.grid.minor = element_line(colour = NA))
+# 
+# ##looks like just one species
+# 
+# 
+# ##differences within canopy types
+# ggplot(data=open) +
+#   geom_smooth(mapping = aes(x=PARi, y=Photo, color=plant_group)) +
[…]

### 2018-11-13 master_scripts/bivariate_plots.R
added more figures, AQ plots

@@ -1,0 +1,19 @@
+### bivariate relationships
+
+chem <- read.csv("raw_data/photo_chem.csv")
+stom <- read.csv("raw_data/stomatal_density.csv")
+spad <- read.csv("raw_data/chlorophyll.csv")
+
+dat <- merge(chem, spad, all=TRUE)
+
+
+library(doBy)
+stom_agg <- summaryBy(sto_den ~ species + individual, FUN=mean, data=stom, 
+                      keep.names = TRUE)
+
+dat2 <- merge(dat, stom_agg, all=TRUE)
+
+plot(Photo~n_perc, data= chem, col=plant_group, pch=16)
+
+
+plot(n_perc ~ spad, data=dat2, col=plant_group, pch=16)

### 2018-11-13 master_scripts/lrc_plots.R

@@ -1,0 +1,52 @@
+##light response figures
+
+aq <- read.csv("calculated_data/AQ_params.csv")
+  aq$species <- gsub("-\\d+","", aq$ID)
+  aq$individual <- gsub(".*-","", aq$ID)
+
+treatments <- read.csv("raw_data/species_list.csv")
+library(vioplot)
+  
+aq2 <- merge(aq, treatments[,c(2,4:5)])
+
+
+##boxplots (habitat overall)
+
+boxplot(Phi ~ plant_group, data=aq2, outline = FALSE, 
+        ylab =  "Quantum Yield") # Quantum yield
+
+boxplot(Rd ~ plant_group, data=aq2, outline = FALSE, 
+        ylab =  "Rd") # Mitochondrial respiration in the light (Rd)
+
+boxplot(LCP ~ plant_group, data=aq2, outline = FALSE, 
+        ylab =  "Light compensation point") # The light compensation point (LCP)
+
+## groups by habitat
+boxplot(Phi ~ canopy, data=aq2, outline = FALSE, 
+        ylab =  "Quantum Yield") # Quantum yield
+
+boxplot(Rd ~ canopy, data=aq2, outline = FALSE, 
+        ylab =  "Rd") # Mitochondrial respiration in the light (Rd)
+
+boxplot(LCP ~ canopy, data=aq2, outline = FALSE, 
+        ylab =  "Light compensation point") # The light compensation point (LCP)
+
+##by group x habitat
+aq2$uniqueid<- paste(aq2$plant_group, aq2$canopy, sep="-")
+
+boxplot(Phi ~ uniqueid, data=aq2, outline = FALSE,xaxt='n',at=c(1:2, 4:5,7),
+        ylab="Quantum Yield")
+axis(1, labels = c("Closed", "Open", "Closed", "Open"), at=c(1:2, 4:5,7))
+mtext(text = c("Angiosperms","Ferns"), side=1, line=2.5,at=c(1.5, 4.5,7))
+
+boxplot(Rd ~ uniqueid, data=aq2, outline = FALSE,xaxt='n',at=c(1:2, 4:5,7),
+        ylab="Rd")
+axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5,7))
+mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, 
[…]

### 2018-11-13 master_scripts/nitro.R

@@ -30,3 +30,15 @@ vioplot(open, closed,at=1:2 ,add=TRUE,
         col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
 axis(2, labels=FALSE, tcl=.25)
 axis(1, labels = c("open","closed"), at=1:2)
+
+##by group x habitat
+nitro$uniqueid<- paste(nitro$plant_group, nitro$canopy, sep="-")
+id_labels <- c("Angiosperm\nClosed", "Angiosperm\nOpen", "Fern\nClosed",
+               "Fern\nOpen", "Lycophyte\nClosed")
+
+windows()
+boxplot(n_perc ~ uniqueid, data=nitro, outline = FALSE, xaxt='n',at=c(1:2, 4:5, 7))
+axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5, 7))
+mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, line=2.5,
+      at=c(1.5, 4.5, 7))
+

### 2018-11-13 master_scripts/photosynthesis.R

@@ -38,7 +38,7 @@ vioplot(open, closed,at=1:2 ,add=TRUE,
 axis(2, labels=FALSE, tcl=.25)
 axis(1, labels = c("open","closed"), at=1:2)
 
-##by group x habitat
+##by group x habitat = Photosynthesis
 photo$uniqueid<- paste(photo$plant_group, photo$canopy, sep="-")
 
 par(mar=c(5,5,1,1))
@@ -49,7 +49,26 @@ mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, line=2.5,
       at=c(1.5, 4.5, 7))
 
 
+##by group x habitat = Conductance
+photo$uniqueid<- paste(photo$plant_group, photo$canopy, sep="-")
 
+par(mar=c(5,5,1,1))
+boxplot(Cond ~ uniqueid, data=photo, outline = FALSE,xaxt='n',at=c(1:2, 4:5, 7),
+        ylab="Stomatal Conductance")
+axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5, 7))
+mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, line=2.5,
+      at=c(1.5, 4.5, 7))
+
+
+####by group x habitat = WUE
+photo$WUE <- with(photo, Photo/Trmmol)
+
+par(mar=c(5,5,1,1))
+boxplot(WUE ~ uniqueid, data=photo, outline = FALSE,xaxt='n',at=c(1:2, 4:5, 7),
+        ylab="WUE")
+axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5, 7))
+mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, line=2.5,
+      at=c(1.5, 4.5, 7))
 
 ##using ggplot
 # p <- ggplot(photo, aes(x=plant_group, y=Photo, color=bla)) + 

### 2019-02-15 manuscript/manuscript.Rmd
stomatal stats and methods written

@@
 We speculate that the differences in physiological function may affect the ecology of these ferns, lycophytes, and angiosperms.
+
 #Methods
@@
 This study was conducted at two sites in upstate New York.
-The data for most species were gathered at the Edmund Niles Huyck Preserve and Biological Research Station in Rensselaerville, New York.
-This site is characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range.
+Data for most species were gathered at the Edmund Niles Huyck Preserve (ENHP), characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range.
 Data from the second site were gathered at a natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley.
-Forests at this site are younger, but consist of a similar species richness and composition as the first site.
+Forests at this site are younger, but consist of a similar species richness and composition at ENHP.
 
@@
 The main focus of this work was to examine ferns relative to light-analogous angiosperms.
-To accomplish this, we would select a fern species and then locate the nearest neighbor angiosperm growing in similar habitat conditions.
-We limited our choice of angiosperms to those taxa that were herbaceous and similar in size and stature to the fern in question.
-At the Huyck Preserve, we sampled across two distinct habitats: understory low-light areas and open, high-light areas.
+To accomplish this, we selected a fern species and then located the nearest neighboring angiosperm growing in similar habitat conditions.
+Surveyed angiosperm taxa were always herbaceous and similar in size and stature to the neighboring fern species.
+At the ENHP, we sampled across two distinct habitats: understory low-light areas and open, high-light areas.
+At the Colgate University site ......
 
@@
 To generate comparative physiological data, we followed two procedures.
-The first was to generate a series of detailed light response curves for a select number of species.
-The second approach, to increase our comparative sample size, was to make several survey measurements of maximum photosynthetic and conductance rates for additional species; this step was only completed at the Huyck Preserve.
+First, we generated a series of detailed light response curves for a select number of species from each plant group (fern = , angiosperm = ).
+Next, to increase our comparative sample size, we surveyed light saturated rates of photosynthesis (A~n~) and stomatal conductance (g~s~) for additional species at the ENHP.
 All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (Lincoln, NE).
-For the light response curves, the system was set up to a flow rate of 300 units, and a CO~2~ concentration of 400 umol m^-2 s^-1.
+
+\
+
+For the light response curves, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm).
 Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels.
 These measurements were used to generate light response curves using appropriate light levels for high-light and low-light species separately.
-The light levels were not identical for each species, with several points being added or subtracted with each run, but they generally followed the same template.
-For high light species this was light levels of 1500, 1000, 800, 600, 500, 400, 300, 250, 200, 150, 100, 50, 25, 15, 10,5 and 0 (umol m-2 s-1).
-Low light curves followed this pattern; 500, 400, 350, 300, 250, 200, 150, 100, 50, 25, 15, 10, 5 and 0 (again umol m-2 s-1).
-Using the data we generated from the light response curves, we developed an understanding of the levels at which low-light and high-light species reached their maximum photosynthetic rates.
-These data were then used to complete our light saturated gas exchange survey measurements.
-All high-light species reached maximum rates at 1200 umol m-2 s-1 and all low-light species reached maximum rates at 600.
-All cuvette parameters were identical to those used for the light response curves.
+The light levels were not identical for each species, with several points being added or subtracted with each run.
+Broadly, light response curves followed the same template for high light or low light species.
+For high light species, light levels started at a PPFD of 1500 $\mu$mol m^-1^ s^-1^ and then consisted of ~16 additional steps to O $\mu$mol m^-1^ s^-1^.
+For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^.
+From these light response curves, we developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates.
+As a result, additional survey measurments were conducted at at a PPFD of 1200 $\mu$mol m^-1^ s^-1^ for high-light species and 600 $\mu$mol m^-1^ s^-1^ for low-light species.
+All other cuvette parameters were identical to those used for the light response curves.
+
+temperature?
 
@@
 
-In open canopy habitats, ferns had significantly lower stomatal densities than those growing in closed canopies, whereas angiosperms showed no significant differences in stomatal densities between the two habitats.
-In both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms (Figure asdfs).
+Overall, no differences where detected in stomatal densities between species growing open canopy and closed canopy habitats.
+Specifically, differences in stomatal densities were not detected for either fern nor angiosperms species between the two habitats.
+However, in both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms () (Figure asdfs).
+In
+
 Similarly, ferns displayed a significantly lower SPAD index, a measurement of chlorophyll concentration, in open canopy environments versus closed canopies.

### 2019-02-15 master_scripts/ci_plots.R

@@ -1,0 +1,54 @@
+source("master_scripts/plot_objects.R")
+
+## phtosynthetic rate density by plant type & habitat
+
+photo <- read.csv("raw_data/photo_chem.csv")
+library(vioplot)
+library(Hmisc)
+
+angio <- photo[photo$plant_group == "Angio", "Ci"]
+fern <- photo[photo$plant_group == "Fern", "Ci"]
+lyco <- photo[photo$plant_group == "Lyco", "Ci"]
+
+open <- photo[photo$canopy == "Open", "Ci"]
+closed <- photo[photo$canopy == "Closed", "Ci"]
+
+#photosynthesis violin-----
+jpeg(filename = "output/ci.jpeg",
+     width = 10, height = 6, units = "in", res= 400)
+par(mfrow=c(1,2))
+##plant group
+par(mar=c(5,5,1,0))
+plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,400), xaxt='n', ylab=cilab,
+     xlab="")
+vioplot(lyco, fern, angio,at=1:3 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+# axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Lycophytes","Ferns","Angiosperms"), at=1:3)
+
+##habitat
+par(mar=c(5,0,1,1))
+plot(0:1,0:1,type="n",xlim=c(0.5,2.5), ylim=c(0,400), xaxt='n', yaxt='n',ylab="",
+     xlab="")
+vioplot(open, closed,at=1:2 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Open","Closed"), at=1:2)
+dev.off()
+
+##by group x habitat = Photosynthesis
+photo$uniqueid<- paste(photo$plant_group, photo$canopy, sep="-")
+
+
+jpeg(filename = "output/ci2.jpeg",
+     width = 8, height = 6, units = "in", res= 400)
+par(mar=c(5,5,1,1))
+boxplot(Photo ~ uniqueid, data=photo, outline = FALSE,xaxt='n',at=c(1:2, 4:5, 7),
+        ylab=cilab)
[…]

### 2019-02-15 master_scripts/cn_ratio.R

@@ -5,20 +5,22 @@ library(vioplot)
 
 nitro <- chem[complete.cases(chem$n_perc),]
 
-angio <- nitro[nitro$plant_group == "Angio", "n_perc"]
-fern <- nitro[nitro$plant_group == "Fern", "n_perc"]
-lyco <- nitro[nitro$plant_group == "Lyco", "n_perc"]
+nitro$cn <- with(nitro, c_perc/n_perc)
 
-open <- nitro[nitro$canopy == "Open", "n_perc"]
-closed <- nitro[nitro$canopy == "Closed", "n_perc"]
+angio <- nitro[nitro$plant_group == "Angio", "cn"]
+fern <- nitro[nitro$plant_group == "Fern", "cn"]
+lyco <- nitro[nitro$plant_group == "Lyco", "cn"]
 
-jpeg(filename = "output/nitro.jpeg",
+open <- nitro[nitro$canopy == "Open", "cn"]
+closed <- nitro[nitro$canopy == "Closed", "cn"]
+
+jpeg(filename = "output/cn_ratio.jpeg",
      width = 10, height = 6, units = "in", res= 400)
 
 par(mfrow=c(1,2))
 ##plant group
 par(mar=c(5,5,1,0))
-plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,5), xaxt='n', ylab=narealab,
+plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,50),xaxt='n', ylab="CN ratio",
      xlab="")
 vioplot(lyco, fern, angio,at=1:3 ,add=TRUE,
         col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
@@ -27,7 +29,7 @@ axis(1, labels = c("Lycophytes","Ferns","Angio"), at=1:3)
 
 ##habitat
 par(mar=c(5,0,1,1))
-plot(0:1,0:1,type="n",xlim=c(0.5,2.5), ylim=c(0,5), xaxt='n', yaxt='n',ylab="",
+plot(0:1,0:1,type="n",xlim=c(0.5,2.5), ylim=c(0,50),xaxt='n', yaxt='n',ylab="",
      xlab="")
 vioplot(open, closed,at=1:2 ,add=TRUE,
         col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
[…]

### 2019-02-15 master_scripts/lrc_plots.R

@@ -34,19 +34,35 @@ boxplot(LCP ~ canopy, data=aq2, outline = FALSE,
 ##by group x habitat
 aq2$uniqueid<- paste(aq2$plant_group, aq2$canopy, sep="-")
 
+#quantum yield
+jpeg(filename = "output/Phi.jpeg",
+     width = 8, height = 6, units = "in", res= 400)
+par(mar=c(5,5,1,1))
 boxplot(Phi ~ uniqueid, data=aq2, outline = FALSE,xaxt='n',at=c(1:2, 4:5,7),
         ylab="Quantum Yield")
-axis(1, labels = c("Closed", "Open", "Closed", "Open"), at=c(1:2, 4:5,7))
-mtext(text = c("Angiosperms","Ferns"), side=1, line=2.5,at=c(1.5, 4.5,7))
+axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5,7))
+mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, 
+      line=2.5,at=c(1.5, 4.5,7))
+dev.off()
 
+## Mitochondrial respiration in the light (Rd)
+jpeg(filename = "output/Rd.jpeg",
+     width = 8, height = 6, units = "in", res= 400)
+par(mar=c(5,5,1,1))
 boxplot(Rd ~ uniqueid, data=aq2, outline = FALSE,xaxt='n',at=c(1:2, 4:5,7),
         ylab="Rd")
 axis(1, labels = c("Closed", "Open", "Closed", "Open", "Closed"), at=c(1:2, 4:5,7))
 mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, 
       line=2.5,at=c(1.5, 4.5,7))
+dev.off()
 
+#light compensation point
+jpeg(filename = "output/lcp.jpeg",
+     width = 8, height = 6, units = "in", res= 400)
+par(mar=c(5,5,1,1))
 boxplot(LCP ~ uniqueid, data=aq2, outline = FALSE,xaxt='n',at=c(1:2, 4:5,7),
-        ylab="Light Compensation Point")
+        ylab=lcplab)
 axis(1, labels = c("Closed", "Open", "Closed", "Open","Closed"), at=c(1:2, 4:5,7))
 mtext(text = c("Angiosperms","Ferns","Lycophytes"), side=1, 
       line=2.5,at=c(1.5, 4.5,7))
+dev.off()

### 2019-02-15 master_scripts/photosynthesis.R

@@ -1,3 +1,5 @@
+source("master_scripts/plot_objects.R")
+
 ## phtosynthetic rate density by plant type & habitat
 
 photo <- read.csv("raw_data/photo_chem.csv")
@@ -5,12 +7,6 @@ library(vioplot)
 library(ggplot2)
 library(Hmisc)
 
-data_summary <- function(x) {
-  m <- mean(x)
-  ymin <- m-sd(x)
-  ymax <- m+sd(x)
-  return(c(y=m,ymin=ymin,ymax=ymax))
-}
 
 angio <- photo[photo$plant_group == "Angio", "Photo"]
 fern <- photo[photo$plant_group == "Fern", "Photo"]
@@ -19,15 +15,18 @@ lyco <- photo[photo$plant_group == "Lyco", "Photo"]
 open <- photo[photo$canopy == "Open", "Photo"]
 closed <- photo[photo$canopy == "Closed", "Photo"]
 
+#photosynthesis violin-----
+jpeg(filename = "output/photo.jpeg",
+     width = 10, height = 6, units = "in", res= 400)
 par(mfrow=c(1,2))
 ##plant group
 par(mar=c(5,5,1,0))
-plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,25), xaxt='n', ylab="Photosynthesis",
+plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,25), xaxt='n', ylab=photolab,
      xlab="")
 vioplot(lyco, fern, angio,at=1:3 ,add=TRUE,
         col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
 # axis(2, labels=FALSE, tcl=.25)
-axis(1, labels = c("lycophytes","ferns","angio"), at=1:3)
+axis(1, labels = c("Lycophytes","Ferns","Angiosperms"), at=1:3)
 
 ##habitat
 par(mar=c(5,0,1,1))
@@ -36,39 +35,78 @@ plot(0:1,0:1,type="n",xlim=c(0.5,2.5), ylim=c(0,25), xaxt='n', yaxt='n',ylab="",
 vioplot(open, closed,at=1:2 ,add=TRUE,
         col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
 axis(2, labels=FALSE, tcl=.25)
-axis(1, labels = c("open","closed"), at=1:2)
[…]

### 2019-02-15 master_scripts/plot_objects.R

@@ -1,0 +1,10 @@
+#plotobjects
+
+condlab <- expression(italic(g)[s]~~(mol~m^-2~s^-1))
+photolab <- expression(italic(A[n])~~(mu*mol~m^-2~s^-1))
+
+narealab <- expression(Leaf~Nitrogen[area]~~(mg~m^-2))
+cilab <- expression(italic(C)[i]~~(ppm))
+
+lcplab <- expression(LCP~~(mu*mol~m^-2~s^-1))
+denslab <- expression(Stomatal~Density~~(mm^-2))

### 2019-02-15 master_scripts/stomata_density.R

@@ -10,6 +10,8 @@ angio_sd <- stom[stom$plant_group == "Angio", "sto_den"]
 fern_sd <- stom[stom$plant_group == "Fern", "sto_den"]
 lyco_sd <- stom[stom$plant_group == "Lycophyte", "sto_den"]
 
+open <- stom2[stom2$canopy == "Open", "sto_den"]
+closed <- stom2[stom2$canopy == "Closed", "sto_den"]
 
 plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,80), xaxt='n', ylab="",
      xlab="")
@@ -18,28 +20,43 @@ vioplot(lyco_sd, fern_sd, angio_sd,at=1:3 ,add=TRUE,
 axis(2, labels=FALSE, tcl=.25)
 axis(1, labels = c("lycophytes","ferns","angiosperms"), at=1:3)
 
+jpeg(filename = "output/stodens.jpeg",
+     width = 10, height = 6, units = "in", res= 400)
+par(mfrow=c(1,2))
+##plant group
+par(mar=c(5,5,1,0))
+plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,80), xaxt='n', ylab=denslab,
+     xlab="")
+vioplot(lyco_sd, fern_sd, angio_sd,at=1:3 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+# axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Lycophytes","Ferns","Angiosperms"), at=1:3)
 
-## boxplots (overall)
-par(mar=c(5,5,1,1))
-boxplot(sto_den ~ plant_group, data=stom, outline = FALSE, xaxt='n')
-axis(1, labels = c("Angiosperms","Ferns","Lycophytes"), at=1:3)
+##habitat
+par(mar=c(5,0,1,1))
+plot(0:1,0:1,type="n",xlim=c(0.5,2.5), ylim=c(0,80), xaxt='n', yaxt='n',ylab="",
+     xlab="")
+vioplot(open, closed,at=1:2 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Open","Closed"), at=1:2)
+dev.off()
 
-##boxplots (habitat overall)
 
[…]

### 2019-02-15 stats/canopy_cover_stats.R

@@ -1,13 +1,34 @@
-#Stats for species groups
+#Stats for overall difference between open vs closed
 library(visreg)
 library(car)
 library(lme4)
 library(MuMIn)
 library(multcomp)
 
+#cover type treatments by species
+treatments <- read.csv("raw_data/species_list.csv")
+
+#stomatal anatomy
 stom <- read.csv("raw_data/stomatal_density.csv")
+stom2 <- aq2 <- merge(stom, treatments[,c(2,4:5)])
+
+stom_agg <- doBy::summaryBy(sto_den ~ species + individual + plant_group,
+                            FUN=mean, data=stom2)
+
+
+#gas exchange
 photo <- read.csv("raw_data/photo_chem.csv")
-photo$iwue <- with(photo, Photo/Trmmol)
+  photo$iwue <- with(photo, Photo/Trmmol)
+
+#stoichemotry
+chem <- read.csv("raw_data/photo_chem.csv")
+
+#light response curves
+aq <- read.csv("calculated_data/AQ_params.csv")
+  aq$species <- gsub("-\\d+","", aq$ID)
+  aq$individual <- gsub(".*-","", aq$ID)
+aq2 <- merge(aq, treatments[,c(2,4:5)])
+
 
 
 ##each model has species as random effect

### 2019-02-15 stats/stomatal_stats.R

@@ -1,0 +1,105 @@
+#Stats for overall difference between open vs closed
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+
+#cover type treatments by species
+treatments <- read.csv("raw_data/species_list.csv")
+
+#stomatal anatomy
+stom <- read.csv("raw_data/stomatal_density.csv")
+
+#mean for each individual (multiple stomatal counts)
+stom_agg <- doBy::summaryBy(sto_den ~ species + individual + plant_group,
+                            FUN=mean, data=stom, keep.names = TRUE)
+
+stom_dens <- merge(stom_agg, treatments[,c(4:5)])
+stom_dens$canopyplant <- interaction(stom_dens$canopy, stom_dens$plant_group)
+
+###comparative stats-----------------------
+fernangio <- droplevels(stom_dens[!stom_dens$plant_group %in% "Lycophyte", ])
+
+#stomatal density by light (random effect of plant group & species)
+
+#excludes light bc not in both canopies
+stomden_mod <- lmer(sto_den ~ canopy + (1|plant_group/species), data=fernangio)
+visreg(stomden_mod)
+
+Anova(stomden_mod)
+summary(stomden_mod)
+r.squaredGLMM(stomden_mod)
+
+tukey_stom <- glht(stomden_mod, linfct = mcp(canopy = "Tukey"))
+stom_siglets <-cld(tukey_stom)
+stom_siglets2 <- stom_siglets$mcletters$Letters
+# not different overall, tons of group+ species variation
+
+stomden_mod2 <- lmer(sto_den ~ canopy * plant_group + (1|species), data=stom_dens)
+
+visreg(stomden_mod2)
+library(emmeans)
+emmip(stomden_mod2, canopy ~ plant_group)
+emmeans(stomden_mod2, pairwise ~ canopy : plant_group)
+
+Anova(stomden_mod2)
+summary(stomden_mod2)
+r.squaredGLMM(stomden_mod2)
+#interaction was not significant so dropped
+
+
+stomden_mod3 <- lmer(sto_den ~ canopy + plant_group + (1|species), data=stom_dens)
+plot(stomden_mod3) # pretty good
+qqPlot(residuals(stomden_mod3)) # pretty good
+
+Anova(stomden_mod3)
[…]

### 2019-02-18 manuscript/manuscript.Rmd
started stats for photo and gs

@@
 At the ENHP, we sampled across two distinct habitats: understory low-light areas and open, high-light areas.
-At the Colgate University site ......
+At the Colgate University site, additioanl low-light fern and lycophyte species were surveyed.
 
@@
 
-For the light response curves, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm).
+Light response curves were conducted by decreasing the light intensity in the gas exchange cuvette in small step changes.
+For each light response curve, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm).
 Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels.
@@
 For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^.
+The light compensation point (LCP) was calculated as the PPFD at which the net photosynthetic rate equalled zero from the linear phase of each light response curve.
+Quantum yeild (phi) was calculated as the initial slope of the assimilation rate and dark respiration rate was recorded at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.
+
+\
+
 From these light response curves, we developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates.
@@
 
-Air-dried and ground tissues from each leaf sample used for gas exchange were analyzed for total C and N.
-Foliar carbon and nitrogen content were generated using a costech elemental analyzer (machine).
-Phosphorus data were gathered using some digest (method)
+Foliar tissue was sampled following gas exchange measurements for all species at ENHP and used for nutrient analyses.
+Samples were dried to a constant mass and ground using a Wig-L-Bug (Sigma-Aldrich Co.
+St.
+Louis, USA).
+Carbon and nitrogen analyses were measured using a Costech Analytical Elemental Analyzer (Valencia, USA), with the percentage of carbon and nitrogen in samples calculated by comparison with certified standards.
+Foliar phosphorus concentrations were determined using an ash digestion process (D’Angelo et al., 2001) preceded by colour development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon).
+
+Photosynthetic nitrogen use efficiency (PNUE) was defined as the ratio of photosynthesis to leaf nitrogen content on a mass basis.*
 
@@
 
+Stomatal density (SD) was measured by directly counting stomata on the abaxial leaf surface under 40x magnification with a field of view of 0.622 mm^2.
+Stomatal density was calculated from 8 non-overlapping foliar regions for five individuals of each species.
+
 ## Statistial analyses
+
+Linear mixed-effect models responses were used to test responses of functional traits to caetgorial fixed effects of plant group and canopy habitat.
+To test broad differnces in plant groups, species nested within canopy habitats were treated as random effects.
+To test differences between habitat types (ferns and angiosperms only) species was set as a fixed effect.
+Explained variance (R2) of mixed models were computed as in Nakagawa and Schielzeth (2013), in which the marginal R2 represents variance explained by fixed factors and the conditional R2 by both fixed and random factors.Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package (Hothorn et al., 2008).
+All tests of statistical significance were conducted at an α level of 0.05.
+All analyses were performed with R 3.5.1 [R cite].
+
+T-tests were performed to test for differences between functional traits, pooled for species, across canopy habitats within angioperm and fern plant groups.
 
@@
 
+Overall, fern and angiosperm species in open habitats had higher A~n~ than compaable species in low-light habitats (*P* <- 0.0001).
+Specifically, fern species in high-light habitats had 40.6 % greater A~n~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 60 % higher A~n~ (*P* < 0.0001, t=-6.97).
+Ferns and angiosperms growing in low-light habitats had similar A~n~, however, in open habitats angiosperms had 42.6% higher rates of A~n~ than ferns (canopy x plant group, *P* = 0.043).
+
+Specifically, fern species in high-light habitats had 50.7 % greater g~s~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 66.9 % higher g~s~ (*P* < 0.0001, t=-4.157).
+
 Ferns growing in open environments had significantly higher maximum area-based photosynthetic rates and stomatal conductances and lower internal leaf CO2 concentrations than those growing in closed canopies (Figure dfg).
@@
 
-Across closed canopy environments where lycophytes, some of these differences between angiosperms and ferns break down.
-For example, there is no significant difference between the photosynthetic rates of angiosperms and ferns, although these two groups are significantly different than lycophytes.
-This relationship is also shown in internal CO2 concentrations, with angiosperms and ferns showing no significant difference and lycophytes differing significantly from both groups.
-For stomatal conductances, however, all three groups are significantly different (Figure adfd).
+Across closed canopy environments, some of these differences between angiosperms and ferns break down when the model includes lycophytes.
+No differences between A~n~ were detected between between plant groups, with a majority of the variation associated with the random effect of individual species.
+Stomatal conducance was higher in angiosperms than lycophytes, while ferns were not different from either plant group (*P* = 0.005).
+The leaf internal CO~2~ concentrations were not different between plant groups, with a majority of the variation associated with the random effect of individual species.
 
-# Nutrient relations
+# Stoichiometry
 

### 2019-02-18 master_scripts/np_ratio.R

@@ -1,0 +1,55 @@
+## phtosynthetic rate density by plant type & habitat
+
+chem <- read.csv("raw_data/photo_chem.csv")
+library(vioplot)
+
+phos <- chem[complete.cases(chem$p_perc),]
+
+phos$np <- with(phos, c_perc/p_perc)
+
+angio <- phos[phos$plant_group == "Angio", "np"]
+fern <- phos[phos$plant_group == "Fern", "np"]
+lyco <- phos[phos$plant_group == "Lyco", "np"]
+
+open <- phos[phos$canopy == "Open", "np"]
+closed <- phos[phos$canopy == "Closed", "np"]
+
+jpeg(filename = "output/np_ratio.jpeg",
+     width = 10, height = 6, units = "in", res= 400)
+
+par(mfrow=c(1,2))
+##plant group
+par(mar=c(5,5,1,0))
+plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,500),xaxt='n', ylab="NP ratio",
+     xlab="")
+vioplot(lyco, fern, angio,at=1:3 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Lycophytes","Ferns","Angio"), at=1:3)
+
+##habitat
+par(mar=c(5,0,1,1))
+plot(0:1,0:1,type="n",xlim=c(0.5,2.5), ylim=c(0,500),xaxt='n', yaxt='n',ylab="",
+     xlab="")
+vioplot(open, closed,at=1:2 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Open","Closed"), at=1:2)
+
+dev.off()
+
+##by group x habitat-----
+nitro$uniqueid<- paste(nitro$plant_group, nitro$canopy, sep="-")
+id_labels <- c("Angiosperm\nClosed", "Angiosperm\nOpen", "Fern\nClosed",
+               "Fern\nOpen", "Lycophyte\nClosed")
+
+
+jpeg(filename = "output/cn2.jpeg",
+     width = 8, height = 6, units = "in", res= 400)
+par(mar=c(5,5,1,1))
[…]

### 2019-02-18 master_scripts/phosphorous.R

@@ -1,0 +1,54 @@
+## phtosynthetic rate density by plant type & habitat
+source("master_scripts/plot_objects.R")
+
+chem <- read.csv("raw_data/photo_chem.csv")
+library(vioplot)
+
+phos <- chem[complete.cases(chem$p_perc),]
+
+angio <- phos[phos$plant_group == "Angio", "p_perc"]
+fern <- phos[phos$plant_group == "Fern", "p_perc"]
+lyco <- phos[phos$plant_group == "Lyco", "p_perc"]
+
+open <- phos[phos$canopy == "Open", "p_perc"]
+closed <- phos[phos$canopy == "Closed", "p_perc"]
+
+jpeg(filename = "output/phos.jpeg",
+     width = 10, height = 6, units = "in", res= 400)
+
+par(mfrow=c(1,2))
+##plant group
+par(mar=c(5,5,1,0))
+plot(0:1,0:1,type="n",xlim=c(0.5,3.5), ylim=c(0,.75), xaxt='n', ylab=parealab,
+     xlab="")
+vioplot(lyco, fern, angio,at=1:3 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Lycophytes","Ferns","Angio"), at=1:3)
+
+##habitat
+par(mar=c(5,0,1,1))
+plot(0:1,0:1,type="n",xlim=c(0.5,2.5), ylim=c(0,.75), xaxt='n', yaxt='n',ylab="",
+     xlab="")
+vioplot(open, closed,at=1:2 ,add=TRUE,
+        col="grey98", lwd=2,rectCol="grey60", colMed="black", pchMed=16, wex=.75)
+axis(2, labels=FALSE, tcl=.25)
+axis(1, labels = c("Open","Closed"), at=1:2)
+
+dev.off()
+
+##by group x habitat-----
+phos$uniqueid<- paste(phos$plant_group, phos$canopy, sep="-")
+id_labels <- c("Angiosperm\nClosed", "Angiosperm\nOpen", "Fern\nClosed",
+               "Fern\nOpen", "Lycophyte\nClosed")
+
+
+jpeg(filename = "output/phos2.jpeg",
+     width = 8, height = 6, units = "in", res= 400)
+par(mar=c(5,5,1,1))
[…]

### 2019-02-18 master_scripts/photo_stats.R

@@ -1,0 +1,99 @@
+#Stats for overall difference between open vs closed
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+
+
+#photosynthesis
+photo <- read.csv("raw_data/photo_chem.csv")
+  photo$canopyplant <- interaction(photo$canopy, photo$plant_group)
+  
+which(is.na(photo[complete.cases(photo$Photo), "Photo"])) #same length as photo
+  
+#is design balanced?
+fern <- droplevels(photo[photo$plant_group == "Fern", ])
+length(unique(fern$species)) #17
+
+angio <-droplevels(photo[photo$plant_group == "Angio", ])
+length(unique(angio$species))  #10
+
+lyco <-droplevels(photo[photo$plant_group == "Lyco", ])
+length(unique(lyco$species)) #6
+
+##use TYPE 3 sum of squares in Anova for unbalanced design
+
+  
+###comparative stats-----------------------
+shade <- droplevels(photo[photo$canopy == "Closed",])
+shade_mod <-lmer(Photo ~ plant_group + (1|species), data=shade)
+plot(shade_mod)
+
+Anova(shade_mod, type = "3")
+summary(shade_mod)
+r.squaredGLMM(shade_mod) 
+visreg(shade_mod)
+
+tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
+shade_siglets <-cld(tukey_shade)
+shade_siglets2 <- shade_siglets$mcletters$Letters
+  
+
+#canopy comparisions of sun shade with angio and ferns
+fernangio <- droplevels(photo[!photo$plant_group %in% "Lyco", ])
+
+#interaction between canopy and plant group
+photo_mod2 <- lmer(Photo ~ canopy * plant_group + (1|species), data=fernangio)
+
+visreg(photo_mod2, "canopy", by="plant_group")
+
+library(emmeans)
+emmip(photo_mod2, canopy ~ plant_group)
+emmeans(photo_mod2, pairwise ~ canopy : plant_group)
+
+Anova(photo_mod2, type="3")
+summary(photo_mod2)
+r.squaredGLMM(photo_mod2)
+#interaction between plant group and 
+#           R2m       R2c
+#[1,] 0.5301052 0.8926862
+
+mean(fernangio[fernangio$canopy == "Open" & fernangio$plant_group == "Angio", "Photo"])
[…]

### 2019-02-18 master_scripts/plot_objects.R

@@ -4,6 +4,8 @@ condlab <- expression(italic(g)[s]~~(mol~m^-2~s^-1))
 photolab <- expression(italic(A[n])~~(mu*mol~m^-2~s^-1))
 
 narealab <- expression(Leaf~Nitrogen[area]~~(mg~m^-2))
+parealab <- expression(Leaf~Phosphorous[area]~~(mg~m^-2))
+
 cilab <- expression(italic(C)[i]~~(ppm))
 
 lcplab <- expression(LCP~~(mu*mol~m^-2~s^-1))

### 2019-02-18 stats/ci_stats.R

@@ -1,0 +1,21 @@
+#Stats for overall difference between open vs closed
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+
+
+#photosynthesis
+ci <- read.csv("raw_data/photo_chem.csv")
+ci$canopyplant <- interaction(ci$canopy, ci$plant_group)
+
+
+###comparative stats-----------------------
+shade <- droplevels(ci[ci$canopy == "Closed",])
+shade_mod <-lmer(Photo ~ plant_group + (1|species), data=shade)
+Anova(shade_mod)
+summary(shade_mod)
+r.squaredGLMM(shade_mod) 
+visreg(shade_mod)
+

### 2019-02-18 stats/gs_stats.R

@@ -1,0 +1,82 @@
+#Stats for overall difference between open vs closed
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+library(pbkrtest)
+
+
+#photosynthesis
+gs <- read.csv("raw_data/photo_chem.csv")
+gs$canopyplant <- interaction(gs$canopy, gs$plant_group)
+
+fern <- droplevels(gs[gs$plant_group == "Fern", ])
+
+angio <-droplevels(gs[gs$plant_group == "Angio", ])
+
+lyco <-droplevels(gs[gs$plant_group == "Lyco", ])
+
+
+###comparative stats-----------------------
+shade <- droplevels(gs[gs$canopy == "Closed",])
+shade_mod <-lmer(sqrt(Cond) ~ plant_group + (1|species), data=shade)
+plot(shade_mod)
+#used sqrt to deal with residuals versus fitted values
+
+Anova(shade_mod)
+summary(shade_mod)
+r.squaredGLMM(shade_mod) 
+visreg(shade_mod)
+
+#          R2m       R2c
+# [1,] 0.2768772 0.7433387
+# P = 0.005 **
+
+tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
+shade_siglets <-cld(tukey_shade)
+shade_siglets2 <- shade_siglets$mcletters$Letters
+#Angio  Fern  Lyco 
+#"b"  "ab"   "a"
+
+
+#canopy comparisions of sun shade with angio and ferns
+fernangio <- droplevels(gs[!gs$plant_group %in% "Lyco", ])
+agricolae::skewness(fernangio$Cond)
+
+#interaction between canopy and plant group
+gs_mod2 <- lmer(sqrt(Cond) ~ canopy * plant_group + (1|species), data=fernangio)
+plot(gs_mod2)
+
+visreg(photo_mod2, "canopy", by="plant_group")
+
+library(emmeans)
+emmip(gs_mod2, canopy ~ plant_group)
+emmeans(gs_mod2, pairwise ~ canopy : plant_group)
+
+Anova(gs_mod2, type="3") #type 3 sum of sqaures for unbalanced
+
+summary(gs_mod2)
+r.squaredGLMM(gs_mod2)
+#interaction between plant group and 
+#           R2m       R2c
+#[1,] 0.4035651 0.889268
+
[…]

### 2019-02-19 manuscript/manuscript.Rmd
lrc stats

@@
 /
+
 Key Words:
@@
 Data from the second site were gathered at a natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley.
-Forests at this site are younger, but consist of a similar species richness and composition at ENHP.
+Forests at the Colgate University site are younger, but consist of a similar species richness and composition as ENHP.
 
@@
 To generate comparative physiological data, we followed two procedures.
-First, we generated a series of detailed light response curves for a select number of species from each plant group (fern = , angiosperm = ).
+First, we generated a series of detailed light response curves for a select number of species from each plant group (fern = 5, angiosperm = 5, lycophytes = 3).
+Light response curves were conducted for five individuals of each species.
 Next, to increase our comparative sample size, we surveyed light saturated rates of photosynthesis (A~n~) and stomatal conductance (g~s~) for additional species at the ENHP.
@@
 Broadly, light response curves followed the same template for high light or low light species.
-For high light species, light levels started at a PPFD of 1500 $\mu$mol m^-1^ s^-1^ and then consisted of ~16 additional steps to O $\mu$mol m^-1^ s^-1^.
+For high light species, light levels started at a PPFD of 1500 $\mu$mol m^-1^ s^-1^ and then consisted of 16 additional steps to O $\mu$mol m^-1^ s^-1^.
 For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^.
 The light compensation point (LCP) was calculated as the PPFD at which the net photosynthetic rate equalled zero from the linear phase of each light response curve.
-Quantum yeild (phi) was calculated as the initial slope of the assimilation rate and dark respiration rate was recorded at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.
+Quantum yeild ($\phi$) was calculated as the initial slope of the assimilation rate and dark respiration rate was recorded at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.
+
+#explain non-rectagular fit
 
@@
 To test broad differnces in plant groups, species nested within canopy habitats were treated as random effects.
-To test differences between habitat types (ferns and angiosperms only) species was set as a fixed effect.
-Explained variance (R2) of mixed models were computed as in Nakagawa and Schielzeth (2013), in which the marginal R2 represents variance explained by fixed factors and the conditional R2 by both fixed and random factors.Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package (Hothorn et al., 2008).
+To test differences between habitat types (ferns and angiosperms only) species was treated as a fixed effect.
+Explained variance (R^2^) of mixed models were computed as in Nakagawa and Schielzeth (2013), in which the marginal R^26 represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
+Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package (Hothorn et al., 2008).
+T-tests were performed to test for differences between functional traits across canopy habitats within angioperm and fern plant groups.
 All tests of statistical significance were conducted at an α level of 0.05.
 All analyses were performed with R 3.5.1 [R cite].
-
-T-tests were performed to test for differences between functional traits, pooled for species, across canopy habitats within angioperm and fern plant groups.
 
@@
 
-## Physiological surveys
+## Foliar gas exchange
 
-Overall, fern and angiosperm species in open habitats had higher A~n~ than compaable species in low-light habitats (*P* <- 0.0001).
+Overall, fern and angiosperm species in open habitats had higher A~n~ than comparable species in low-light habitats (*P* <- 0.0001).
 Specifically, fern species in high-light habitats had 40.6 % greater A~n~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 60 % higher A~n~ (*P* < 0.0001, t=-6.97).
@@
 
+\
+
+Overall, ferns and angiosperms in open habitats had 64.2% higher g~s~ than comparable species in low-ligh habitats (*P* = 0.002).
 Specifically, fern species in high-light habitats had 50.7 % greater g~s~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 66.9 % higher g~s~ (*P* < 0.0001, t=-4.157).
-
-Ferns growing in open environments had significantly higher maximum area-based photosynthetic rates and stomatal conductances and lower internal leaf CO2 concentrations than those growing in closed canopies (Figure dfg).
-Angiosperms growing in open canopy environments also had significantly higher maximum area-based photosynthetic rates and stomatal conductances than those growing in closed environments, but there was no significant difference in internal leaf CO2 concentrations between the two light environments.
+Across both habitat types, angiosperm species had a 60% higher g~s~ than fern species (*P* = 0.011).
 
@@
 
-In open canopy environments, angiosperms, on average, had significantly higher maximum photosynthetic rates and stomatal conductances than ferns growing in similar light conditions.
-There was no significant difference between the internal CO2 concentrations of angiosperm leaves and fern fronds; however, the statistics hint at a strong trend (Figure asldf).
+The leaf internal CO~2~ concentrations (C~i~) was broadly similar across habitat types and between angiosperms and ferns.
+Across angiosperm species, C~i~ was 12.2 % higher in open habitats than in low-ligh habitats (*P* < 0.031, t=2.2631).
+Across fern species, C~i~ was similar among habitat types.
 
@@
 
-Across closed canopy environments, some of these differences between angiosperms and ferns break down when the model includes lycophytes.
-No differences between A~n~ were detected between between plant groups, with a majority of the variation associated with the random effect of individual species.
-Stomatal conducance was higher in angiosperms than lycophytes, while ferns were not different from either plant group (*P* = 0.005).
-The leaf internal CO~2~ concentrations were not different between plant groups, with a majority of the variation associated with the random effect of individual species.
+In closed canopy environments, variation in gas exchange parameters was largely due to individual species variation instead of functional differences between all three plant groups.
+No differences between A~n~ were detected between between the three plant groups, with a majority of the variation associated with the random effect of individual species.
+Stomatal conductance in angiosperms was higher than in ferns and lycophytes **P* = 0.001).
+The leaf internal CO~2~ concentrations (C~i~) were not different between plant groups, however, C~i~ in lycophytes was broadly lower.
 

### 2019-02-19 master_scripts/photo_stats.R

@@ -24,7 +24,6 @@ length(unique(lyco$species)) #6
 
 ##use TYPE 3 sum of squares in Anova for unbalanced design
 
-  
 ###comparative stats-----------------------
 shade <- droplevels(photo[photo$canopy == "Closed",])
 shade_mod <-lmer(Photo ~ plant_group + (1|species), data=shade)

### 2019-02-19 stats/ci_stats.R

@@ -10,12 +10,59 @@ library(multcomp)
 ci <- read.csv("raw_data/photo_chem.csv")
 ci$canopyplant <- interaction(ci$canopy, ci$plant_group)
 
+hist(ci$Ci)
+
+fern <- droplevels(ci[ci$plant_group == "Fern", ])
+
+angio <-droplevels(ci[ci$plant_group == "Angio", ])
+
+lyco <-droplevels(ci[ci$plant_group == "Lyco", ])
+
 
 ###comparative stats-----------------------
 shade <- droplevels(ci[ci$canopy == "Closed",])
-shade_mod <-lmer(Photo ~ plant_group + (1|species), data=shade)
-Anova(shade_mod)
+shade_mod <-lmer(Ci ~ plant_group + (1|species), data=ci)
+plot(shade_mod)
+Anova(shade_mod, type="3")
 summary(shade_mod)
 r.squaredGLMM(shade_mod) 
 visreg(shade_mod)
 
+#            R2m       R2c  #mostly species effects
+#[1,] 0.09194196 0.8199885
+
+
+tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
+shade_siglets <-cld(tukey_shade)
+shade_siglets2 <- shade_siglets$mcletters$Letters
+
+#plant group by habitat
+
+#canopy comparisions of sun shade with angio and ferns
+fernangio <- droplevels(ci[!ci$plant_group %in% "Lyco", ])
+
+#interaction between canopy and plant group
+ci_mod2 <- lmer(Ci ~ canopy * plant_group + (1|species), data=fernangio)
+plot(ci_mod2)
+
+Anova(ci_mod2, type='3')
+visreg(ci_mod2, "canopy", by="plant_group")
+
+#test simple models
+ci_mod3 <- lmer(Ci ~ canopy + plant_group + (1|species), data=fernangio)
+Anova(ci_mod3, type='3')
+
+#no differences regardless
+
+#t-test between groups
+
+t.test(Ci ~ canopy, data=fern)
+#mean in group Closed   mean in group Open 
+#274.4424             240.7790
+# t = 2.2631, df = 29.753, p-value = 0.03109
+#12.2 % higher in open
+
+t.test(Ci ~ canopy, data=angio)
+#mean in group Closed   mean in group Open 
+#273.9929             269.1036 
+# 0.41531, df = 46.821, p-value = 0.6798

### 2019-02-19 stats/gs_stats.R

@@ -17,6 +17,7 @@ angio <-droplevels(gs[gs$plant_group == "Angio", ])
 
 lyco <-droplevels(gs[gs$plant_group == "Lyco", ])
 
+hist(gs$Cond)
 
 ###comparative stats-----------------------
 shade <- droplevels(gs[gs$canopy == "Closed",])
@@ -24,14 +25,14 @@ shade_mod <-lmer(sqrt(Cond) ~ plant_group + (1|species), data=shade)
 plot(shade_mod)
 #used sqrt to deal with residuals versus fitted values
 
-Anova(shade_mod)
+Anova(shade_mod, type="3")
 summary(shade_mod)
 r.squaredGLMM(shade_mod) 
 visreg(shade_mod)
 
 #          R2m       R2c
-# [1,] 0.2768772 0.7433387
-# P = 0.005 **
+# [1,] 0.3237432 0.7420295
+# P = 0.001 **
 
 tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
 shade_siglets <-cld(tukey_shade)
@@ -42,29 +43,54 @@ shade_siglets2 <- shade_siglets$mcletters$Letters
 
 #canopy comparisions of sun shade with angio and ferns
 fernangio <- droplevels(gs[!gs$plant_group %in% "Lyco", ])
-agricolae::skewness(fernangio$Cond)
+agricolae::skewness(sqrt(fernangio$Cond))
 
 #interaction between canopy and plant group
 gs_mod2 <- lmer(sqrt(Cond) ~ canopy * plant_group + (1|species), data=fernangio)
 plot(gs_mod2)
 
-visreg(photo_mod2, "canopy", by="plant_group")
+visreg(gs_mod2, "canopy", by="plant_group")
 
 library(emmeans)
 emmip(gs_mod2, canopy ~ plant_group)
 emmeans(gs_mod2, pairwise ~ canopy : plant_group)
 
 Anova(gs_mod2, type="3") #type 3 sum of sqaures for unbalanced
+#canopy only, 
 
 summary(gs_mod2)
 r.squaredGLMM(gs_mod2)
-#interaction between plant group and 
+#canopy only 
 #           R2m       R2c
-#[1,] 0.4035651 0.889268
+#[1,] 0.4127405 0.8917534
+# canopy, p = 0.003 **
 
[…]

### 2019-02-19 stats/lrc_stats.R

@@ -1,0 +1,99 @@
+#Stats for overall difference between open vs closed
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+
+library(dplyr)
+library(stringr)
+#photosynthesis
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+       mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+              individual =  str_remove(ID, ".*-"))
+              
+#treatments
+treatments <- read.csv("raw_data/species_list.csv")
+
+lrc_hyuck <- merge(lrc, treatments[,c(2,4:5)])
+  lrc_hyuck$canopyplant <- interaction(lrc_hyuck$canopy, lrc_hyuck$plant_group)
+
+
+fern <- droplevels(lrc_hyuck[lrc_hyuck$plant_group == "Fern", ])
+  #length(unique(levels(fern$species)))
+angio <-droplevels(lrc_hyuck[lrc_hyuck$plant_group == "Angiosperm", ])
+  #length(unique(levels(angio$species))) 
+lyco <-droplevels(lrc_hyuck[lrc_hyuck$plant_group == "Lycophyte", ])
+  #length(unique(levels(lyco$species)))
+
+
+hist(lrc_hyuck$Phi)
+hist(lrc_hyuck$Rd)
+hist(lrc_hyuck$LCP)
+
+###comparative stats-----------------------
+shade <- droplevels(lrc_hyuck[lrc_hyuck$canopy == "Closed",])
+shade2 <- droplevels(shade[!shade$ID == "imp_pal-1",])
+
+#remove imp_pal-1
+
+
+#PHI
+shade_phi <-lmer(Phi ~ plant_group + (1|species), data=shade)
+plot(shade_phi)
+#used sqrt to deal with residuals versus fitted values
+
+Anova(shade_phi)
+summary(shade_phi)
+r.squaredGLMM(shade_phi) 
+visreg(shade_phi)
+
+#          R2m       R2c
+# [1,] 0.1591668 0.394554
+# P = .1463**
+
+tukey_phi <- glht(shade_phi, linfct = mcp(plant_group = "Tukey"))
+phi_siglets <-cld(tukey_phi)
+phi_siglets2 <- phi_siglets$mcletters$Letters
+#all same
+
+
+#Rd
[…]

### 2019-02-19 manuscript/manuscript.Rmd
finsihed lrc stats and started PCA

@@
 For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^.
+Light response data for each curves was fit to a non-rectangular hyperbola model (Equation 6 in Lobo et al.
+2013).
 The light compensation point (LCP) was calculated as the PPFD at which the net photosynthetic rate equalled zero from the linear phase of each light response curve.
-Quantum yeild ($\phi$) was calculated as the initial slope of the assimilation rate and dark respiration rate was recorded at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.
-
-#explain non-rectagular fit
+Quantum yeild ($\phi$) was calculated as the initial slope of the assimilation rate and mitochondrial respiration in the light (R~d~) at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.
 
@@
 
-From these light response curves, we developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates.
+From these light response curves, we also developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates.
 As a result, additional survey measurments were conducted at at a PPFD of 1200 $\mu$mol m^-1^ s^-1^ for high-light species and 600 $\mu$mol m^-1^ s^-1^ for low-light species.
@@
 Carbon and nitrogen analyses were measured using a Costech Analytical Elemental Analyzer (Valencia, USA), with the percentage of carbon and nitrogen in samples calculated by comparison with certified standards.
+Photosynthetic nitrogen use efficiency (PNUE) was defined as the ratio of photosynthesis to leaf nitrogen content on a mass basis.
 Foliar phosphorus concentrations were determined using an ash digestion process (D’Angelo et al., 2001) preceded by colour development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon).
-
-Photosynthetic nitrogen use efficiency (PNUE) was defined as the ratio of photosynthesis to leaf nitrogen content on a mass basis.*
 
@@
 
-In closed canopy environments, angiosperms had significantly higher respiration rates than ferns, which were equal to lycophytes (f2,44= 10.12, p=0.0003).
-In addition, photosynthetic quantum yield was higher in lycophytes than in angiosperms, whereas ferns had intermediate values between the two (f2,44=5.04, p=0.011).
-Light compensation points in ferns were equivalent to lycophytes, which both had higher points than angiosperms (f2,44=10.28, p=0.0002).
-In open canopy environments where lycophytes were lacking, ferns had significantly higher respiration rates (f1,18= 20.87, p= 0.0003), equivalent quantum yields (f1,18= 1.56, p=0.228), and significantly lower light compensation points(f1,18=12.83, p=0.0023).
+Overall, ferns and angiosperms had similar $\phi$ across open and low-light habitats.
+Quantum yields were also equivalent within angiosperm and fern species groups between open and low-light habitats.
+Shifts in R~d~ between open and low-light habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001).
+Specifically, R~d~ in angiosperms was 80.6 % greater in open compared to low-light habitats (t=-4.197, *P* = 0.001).
+Alternatively, R~d~ in ferns was 44.0 % greater in low-light compared to open habitats (t=-2.97, p=0.006).
+Similarly, shifts in LCP between open and low-light habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001).
+Specifically, the LCP in angiosperms was 79.2 % greater in open compared to low-light habitats (t=-3.5998, *P* = 0.005).
+Alternatively, the LCP in ferns was 39.7 % greater in low-light compared to open habitats (t=2.1814, *P* = 0.040).
 
@@
 
-For angiosperms, species growing in open habitats had significantly more negative respiration rates (t22=-6.88, p<0.0001), equivalent quantum yields(t22=1.89, p=0.072), and higher light compensation points (t22=5.23, p<0.0001) than those angiosperms growing in closed habitats.
+In closed canopy environments,the LCP in ferns were equivalent to lycophytes, while both plant groups were marginally higher than angiosperms (*P* = 0.061).
+Photosynthetic quantum was similar between each plant group, but lycophytes and ferns were broadly higher than angiosperms.
+Ferns and lycophytes had equavalent respiration rates, while both groups were higher than angiosperms (*P* = 0.0178)
 
-## Foliar gas exchange
+## Leaf gas exchange
 
@@
 
-## Physiological parameters
+## Stomatal anatomy
 

### 2019-02-19 master_scripts/pca_plot.R

@@ -1,0 +1,80 @@
+##merge all datasets into master for PCA
+library(dplyr)
+
+#read and format data-------
+
+treatments <- read.csv("raw_data/species_list.csv")
+
+#light response curves removing one bad curve
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+  mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+         individual =  as.numeric(str_remove(ID, ".*-"))) %>%
+  merge(treatments[,c(2,4:5)]) %>%
+  filter(!ID ==  "imp_pal-1")
+  
+
+#gas exchange and chemistry
+photo <- read.csv("raw_data/photo_chem.csv") %>%
+         mutate(ID = paste(species, individual, sep="-")) %>%
+         filter(!Site == "Colgate") %>%
+         select(-"plant_group") %>%
+         inner_join(treatments[,c(2,5)], by='species')
+
+##now add stomatal density
+stom <- read.csv("raw_data/stomatal_density.csv") %>%
+        select(-"plant_group") %>%
+        inner_join(treatments[,c(2,4:5)], by='species')
+
+stom_agg <- doBy::summaryBy(sto_den ~ species + individual + plant_group + canopy,
+                          FUN=mean, data=stom, keep.names = TRUE) %>%
+            mutate(ID = paste(species, individual, sep="-"))
+
+#merge all data -----
+hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
+       
+#select data for PCA
+hyuck_nona <- hyuck[complete.cases(hyuck),]
+  hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
+  hyuck_nona$CN <- with(hyuck_nona$c_perc/ hyuck_nona$n_perc)
+  hyuck_nona$NP <- with(hyuck_nona$n_perc/ hyuck_nona$p_perc)
+  
+#get rid of some parameters
+hyuck_pca <- droplevels(hyuck_nona[,-c(1:2,6:7,9:13,20:21)])
+
+
+library(vegan)
+
+##look at variances
+# var(hyuck_nona$Photo)
+# var(hyuck_nona$p_perc)
+#variances among variables are orders of magnitude different so we will rescale
+
+test_dca <- decorana(na.omit(hyuck_pca))
+summary(test_dca, display = 'none')
[…]

### 2019-02-19 stats/lrc_stats.R

@@ -7,39 +7,40 @@ library(multcomp)
 
 library(dplyr)
 library(stringr)
-#photosynthesis
-lrc <- read.csv("calculated_data/AQ_params.csv") %>%
-       mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
-              individual =  str_remove(ID, ".*-"))
-              
 #treatments
 treatments <- read.csv("raw_data/species_list.csv")
 
-lrc_hyuck <- merge(lrc, treatments[,c(2,4:5)])
-  lrc_hyuck$canopyplant <- interaction(lrc_hyuck$canopy, lrc_hyuck$plant_group)
+
+#photosynthesis
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+       mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+              individual =  str_remove(ID, ".*-")) %>%
+       merge(treatments[,c(2,4:5)]) %>%
+       mutate(canopyplant = interaction(canopy, plant_group))
 
 
-fern <- droplevels(lrc_hyuck[lrc_hyuck$plant_group == "Fern", ])
+fern <- droplevels(lrc[lrc$plant_group == "Fern", ])
   #length(unique(levels(fern$species)))
-angio <-droplevels(lrc_hyuck[lrc_hyuck$plant_group == "Angiosperm", ])
+angio <-droplevels(lrc[lrc$plant_group == "Angiosperm" &
+                    !lrc$ID == "imp_pal-1", ])
   #length(unique(levels(angio$species))) 
-lyco <-droplevels(lrc_hyuck[lrc_hyuck$plant_group == "Lycophyte", ])
+lyco <-droplevels(lrc[lrc$plant_group == "Lycophyte", ])
   #length(unique(levels(lyco$species)))
 
 
-hist(lrc_hyuck$Phi)
-hist(lrc_hyuck$Rd)
-hist(lrc_hyuck$LCP)
+hist(lrc$Phi)
+hist(lrc$Rd)
+hist(lrc$LCP)
 
 ###comparative stats-----------------------
-shade <- droplevels(lrc_hyuck[lrc_hyuck$canopy == "Closed",])
+shade <- droplevels(lrc[lrc$canopy == "Closed",])
 shade2 <- droplevels(shade[!shade$ID == "imp_pal-1",])
 
 #remove imp_pal-1
 
 
 #PHI
[…]

### 2019-02-21 master_scripts/pca_canopy.R
pca drafts

@@ -12,51 +12,49 @@ lrc <- read.csv("calculated_data/AQ_params.csv") %>%
          individual =  as.numeric(str_remove(ID, ".*-"))) %>%
   merge(treatments[,c(2,4:5)]) %>%
   filter(!ID ==  "imp_pal-1")
-  
+
 
 #gas exchange and chemistry
 photo <- read.csv("raw_data/photo_chem.csv") %>%
-         mutate(ID = paste(species, individual, sep="-")) %>%
-         filter(!Site == "Colgate") %>%
-         select(-"plant_group") %>%
-         inner_join(treatments[,c(2,5)], by='species')
+  mutate(ID = paste(species, individual, sep="-")) %>%
+  filter(!Site == "Colgate") %>%
+  select(-"plant_group") %>%
+  inner_join(treatments[,c(2,5)], by='species')
 
 ##now add stomatal density
 stom <- read.csv("raw_data/stomatal_density.csv") %>%
-        select(-"plant_group") %>%
-        inner_join(treatments[,c(2,4:5)], by='species')
+  select(-"plant_group") %>%
+  inner_join(treatments[,c(2,4:5)], by='species')
 
 stom_agg <- doBy::summaryBy(sto_den ~ species + individual + plant_group + canopy,
-                          FUN=mean, data=stom, keep.names = TRUE) %>%
-            mutate(ID = paste(species, individual, sep="-"))
+                            FUN=mean, data=stom, keep.names = TRUE) %>%
+  mutate(ID = paste(species, individual, sep="-"))
 
 #merge all data -----
 hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
-       
+
 #select data for PCA
 hyuck_nona <- hyuck[complete.cases(hyuck),]
-  hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
-  hyuck_nona$CN <- with(hyuck_nona, c_perc/ n_perc)
-  hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
-  
-hyuck_shade <- hyuck_nona[hyuck_nona$canopy == "Closed",]  
-  
+hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
+hyuck_nona$CN <- with(hyuck_nona, c_perc/ n_perc)
[…]

### 2019-02-28 manuscript/manuscript.Rmd
stats for elemeents

@@
 Specifically, differences in stomatal densities were not detected for either fern nor angiosperms species between the two habitats.
-However, in both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms () (Figure asdfs).
-In
+However, in both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms (*P* < 0.001).
 
-Similarly, ferns displayed a significantly lower SPAD index, a measurement of chlorophyll concentration, in open canopy environments versus closed canopies.
-Angiosperms showed no significant difference across the light gradient.
-In open environments, ferns had significantly lower SPAD index values than corresponding angiosperms, while the reverse was true in closed canopy habitats, with ferns displaying significantly higher values than analogous angiosperms (Figure asdfkj).
+\
+
+In closed canopy environments, no differenes in stomatal density were detected among species of ferns, angiosperms and lycophytes.
+Variation in stomatal denisty in the shade was largely due to individual species variation.

### 2019-02-28 stats/element_stats.R

@@ -1,0 +1,143 @@
+## elemental stats (including ratios)
+
+#Stats for overall difference between open vs closed
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+
+
+#photosynthesis
+chem <- read.csv("raw_data/photo_chem.csv")
+
+nitro <- chem[complete.cases(chem$n_perc),]
+  nitro$cn_ratio <- with(nitro, c_perc/n_perc)
+
+phos <- chem[complete.cases(chem$p_perc),]
+  phos$np_ratio <- with(phos, n_perc/p_perc)
+
+
+###comparative stats-----------------------
+Nshade <- droplevels(nitro[nitro$canopy == "Closed",])
+  Nshade$percN <- Nshade$n_perc /100
+  
+#arcsin
+shade_N <-lmer(asin(sqrt(percN)) ~ plant_group + (1|species), data=Nshade)
+plot(shade_N)
+qqPlot(residuals(shade_N))
+
+# #glm binomial
+# shade_N2 <- glmer(n_perc/100 ~ plant_group + (1|species), family = binomial, 
+#                   data=Nshade)
+
+#logit
+shade_N3 <-lmer(log(percN/(1-percN)) ~ plant_group + (1|species), data=Nshade)
+plot(shade_N3)
+qqPlot(residuals(shade_N3))
+
+Anova(shade_N, type = "3")
+
+Anova(shade_N3, type = "3")
+summary(shade_N3)
+r.squaredGLMM(shade_N3) 
+visreg(shade_N3)
+
+tukey_Nshade <- glht(shade_N3, linfct = mcp(plant_group = "Tukey"))
+shade_Nsiglets <-cld(tukey_Nshade)
+shade_Nsiglets2 <- shade_Nsiglets$mcletters$Letters
+
+# p < 0.001 (lycophete have less N in shade)
+#Angio  Fern  Lyco 
+#"b"   "b"   "a" 
+
+#R2m       R2c
+#0.4537292 0.8403728
+
+mean(Nshade[Nshade$plant_group == "Angio", "percN"]) #0.0303
+mean(Nshade[Nshade$plant_group == "Fern", "percN"])#0.0295
+
+mean(Nshade[Nshade$plant_group == "Lyco", "percN"]) #0.016
+mean(Nshade[!Nshade$plant_group == "Lyco", "percN"]) #0.02983398
[…]

### 2019-02-28 stats/pnue_stats.R

@@ -1,0 +1,1 @@
+#PNUE stats

### 2019-02-28 stats/stomatal_stats.R

@@ -16,42 +16,74 @@ stom_agg <- doBy::summaryBy(sto_den ~ species + individual + plant_group,
                             FUN=mean, data=stom, keep.names = TRUE)
 
 stom_dens <- merge(stom_agg, treatments[,c(4:5)])
-stom_dens$canopyplant <- interaction(stom_dens$canopy, stom_dens$plant_group)
+  stom_dens$canopyplant <- interaction(stom_dens$canopy, stom_dens$plant_group)
+  
+  mean(stom_dens[stom_dens$plant_group == "Angio", "sto_den"])
+  mean(stom_dens[stom_dens$plant_group == "Fern", "sto_den"])
+  mean(stom_dens[stom_dens$plant_group == "Lycophyte", "sto_den"])
+  
+
+fern <- droplevels(stom_dens[stom_dens$plant_group == "Fern", ])
+  
+angio <-droplevels(stom_dens[stom_dens$plant_group == "Angio", ])
+
+lyco <-droplevels(stom_dens[stom_dens$plant_group == "Lycophyte", ])
+
+
+###comparative stats-----------------------
+shade <- droplevels(stom_dens[stom_dens$canopy == "Closed",])
+
+shade_mod <-lmer(sto_den ~ plant_group + (1|species), data=shade)
+plot(shade_mod) #looks good
+
+Anova(shade_mod, type = "3")
+summary(shade_mod)
+r.squaredGLMM(shade_mod) 
+visreg(shade_mod)
+
+#not different in shade, but less in ferns
+
+tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
+shade_siglets <-cld(tukey_shade)
+shade_siglets2 <- shade_siglets$mcletters$Letters
+
+#Angio      Fern Lycophyte 
+#"a"       "a"       "a"
+
+#lost of species variation
+#R2m       R2c
+#0.2958341 0.9223959
+
+mean(shade[shade$plant_group == "Angio", "sto_den"])
+mean(shade[shade$plant_group == "Fern", "sto_den"])
+mean(shade[shade$plant_group == "Lycophyte", "sto_den"])
+
 
 ###comparative stats-----------------------
 fernangio <- droplevels(stom_dens[!stom_dens$plant_group %in% "Lycophyte", ])
 
-#stomatal density by light (random effect of plant group & species)
[…]

### 2019-03-01 manuscript/manuscript.Rmd
ratio stats

@@
 For each light response curve, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm).
+Temperature was not explictly controlled in the leaf cuvette and leaf temperatures ranged from 27-30 &deg;C.
 Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels.
@@
 All other cuvette parameters were identical to those used for the light response curves.
-
-temperature?
 
@@
 All analyses were performed with R 3.5.1 [R cite].
+
+add vegan
 
@@
 Ferns in open light environments have significantly lower nitrogen levels than those ferns growing in closed light environments; the same is true of angiosperms.
-Lycophytes have significantly lower nitrogen levels than ferns and angiosperms, which do not statistically differ, in both light conditions (Figure adf).
+
+Across closed canopy habitats, Lycophytes have 45% lower foliar nitrogen levels than ferns and angiosperms, which do not statistically differ (*P* < 0.001).
+Conseqently, Lycpohtes had significantly higher CN ratios than ferns and angiosperms (*P* < 0.001).
+No differences were detected  between Lycophytes, ferns and angiosperms across closed canopy habitats.
 

### 2019-03-01 stats/element_stats.R

@@ -18,7 +18,7 @@ phos <- chem[complete.cases(chem$p_perc),]
   phos$np_ratio <- with(phos, n_perc/p_perc)
 
 
-###comparative stats-----------------------
+###N in shade-----------------------
 Nshade <- droplevels(nitro[nitro$canopy == "Closed",])
   Nshade$percN <- Nshade$n_perc /100
   

### 2019-03-05 manuscript/manuscript.Rmd
started bivariate and finshed wue and np ration stats, problem with normalizing pnue

--- 2019-02-28 manuscript/manuscript.Rmd
@@
 ```
+
+1) Differences in ecophysiology of groups in same habitat (seed vs angio)
+2) similarities or differnces between open/closed with ferns and lycophytes
+3) Do these tell us why lycophtes dont go to open ?
 
@@
 Next, to increase our comparative sample size, we surveyed light saturated rates of photosynthesis (A~n~) and stomatal conductance (g~s~) for additional species at the ENHP.
-All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (Lincoln, NE).
+All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (LiCor Biosciences, Lincoln, NE USA).
 
@@
 For each light response curve, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm).
+Temperature was not explictly controlled in the leaf cuvette.
+Cuvette leaf temperatures ranged from 27-30 &deg;C and leaf vapor pressure deficit (VPD) ranged from ***** kPa.
 Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels.
@@
 All other cuvette parameters were identical to those used for the light response curves.
-
-temperature?
+Leaf-level instantaneous water use efficiency (WUE) was calculated as A~n~ divided by transpiration from gas exchange measurements.
 
@@
 Carbon and nitrogen analyses were measured using a Costech Analytical Elemental Analyzer (Valencia, USA), with the percentage of carbon and nitrogen in samples calculated by comparison with certified standards.
-Photosynthetic nitrogen use efficiency (PNUE) was defined as the ratio of photosynthesis to leaf nitrogen content on a mass basis.
+Photosynthetic nitrogen use efficiency (PNUE) was defined as the ratio of A~n~ to leaf nitrogen content.
 Foliar phosphorus concentrations were determined using an ash digestion process (D’Angelo et al., 2001) preceded by colour development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon).
@@
 To test differences between habitat types (ferns and angiosperms only) species was treated as a fixed effect.
-Explained variance (R^2^) of mixed models were computed as in Nakagawa and Schielzeth (2013), in which the marginal R^26 represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
+Explained variance (R^2^) of mixed models were computed as in Nakagawa and Schielzeth (2013), in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
 Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package (Hothorn et al., 2008).
@@
 All analyses were performed with R 3.5.1 [R cite].
+
+Principal component analysis, utilizing the ‘vegan’ package (Oksanen et al., 2018), was used to explore how measured functional traits were distributed and co-varied among plant groups and canopy type.
+
+For bivariate trait relationships, responses of dependent variables were analysed with linear mixed-effect models, with species as a random effect and plant group and/or canopy type as categorical fixed effects.
+Differences in slopes of the relationship of bivariate traits by habitat (Selaginella only) were tested by calculating estimated marginal means and computing pairwise comparisons with the ‘emmeans’ package (Length, 2018).
 
@@
 Ferns and angiosperms growing in low-light habitats had similar A~n~, however, in open habitats angiosperms had 42.6% higher rates of A~n~ than ferns (canopy x plant group, *P* = 0.043).
+Neither ferns nor angiosperms species differed in WUE across the two light environments.
+Across all habitats, the WUE was also statistically similar for ferns and angiosperms.
 
@@
 
-Overall, ferns and angiosperms in open habitats had 64.2% higher g~s~ than comparable species in low-ligh habitats (*P* = 0.002).
+Overall, ferns and angiosperms in open habitats had 64.2% higher g~s~ than comparable species in low-light habitats (*P* = 0.002).
 Specifically, fern species in high-light habitats had 50.7 % greater g~s~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 66.9 % higher g~s~ (*P* < 0.0001, t=-4.157).
 Across both habitat types, angiosperm species had a 60% higher g~s~ than fern species (*P* = 0.011).
-
-\
-
-The leaf internal CO~2~ concentrations (C~i~) was broadly similar across habitat types and between angiosperms and ferns.
-Across angiosperm species, C~i~ was 12.2 % higher in open habitats than in low-ligh habitats (*P* < 0.031, t=2.2631).
-Across fern species, C~i~ was similar among habitat types.
 
@@
 Stomatal conductance in angiosperms was higher than in ferns and lycophytes **P* = 0.001).
-The leaf internal CO~2~ concentrations (C~i~) were not different between plant groups, however, C~i~ in lycophytes was broadly lower.
+Water-use efficiency was 47 % higher in lycophytes than ferns and angiosperms (*P* = 0.005), while ferns and angiosperms where statiscally similar.
 
-# Stoichiometry
+# Stoichiometry (##can use pairwise for specific p values: see leaf P below)
 
-In open canopy environments, ferns had significantly lower phosphorus concentrations than those growing in closed canopies.
-Angiosperms had no significant difference across the light gradient.
-In both environments, ferns and lycophytes do not differ significantly, but both have significantly lower phosphorus levels than angiosperms (Figure asdf).
-Ferns in open light environments have significantly lower nitrogen levels than those ferns growing in closed light environments; the same is true of angiosperms.
-Lycophytes have significantly lower nitrogen levels than ferns and angiosperms, which do not statistically differ, in both light conditions (Figure adf).
+Ferns and angiosperms species in open light environments had marginally lower foliar nitrogen content compared to similar species in growing in closed light environments (-17.2 %, *P* = 0.072), with a large proportion of variation attributed to individual species differences.
+Neither ferns nor angiosperms species differed in their CN ratios across the two light environments.
+Across all habitats, the CN ratio of ferns was 20 % higher than in angiosperms (*P* = 0.0325).
+No differences in foliar P content were detected between open and closed canopies for either ferns or angiosperms.
+Across all habitats, foliar P content was 22.1 % higher in angiosperms than in ferns (*P* =  0.034).
+Neither ferns nor angiosperms species differed in their NP ratios across the two light environments.
+Across all habitats, the NP ratio was statistically similar for ferns and angiosperms.
 
@@
 
-Neither ferns nor angiosperms differ within their classifications significantly across the two light environments in their nitrogen to phosphorus ratio (N:P).
-In open light environments, ferns have significantly higher N:P ratios than angiosperms.
-In closed canopy environments, ferns do not differ from angiosperms significantly in their N:P ratios, but do differ from lycophytes.
-Also, lycophytes and angiosperms do not differ significantly.
-
-\
-
-Ferns in open light canopies have significantly higher carbon to nitrogen ratios (C:N) than those growing in closed canopies, while angiosperms do not differ across the light gradient.
-Also, ferns have significantly higher C:N ratios than corresponding angiosperms in open light environments.
-In closed canopies, lycophytes have the highest C:N ratios and differ significantly from both ferns and angiosperms, which do not significantly differ.
+Across closed canopy habitats, Lycophytes have 45% lower foliar nitrogen content than ferns and angiosperms, which do not statistically differ (*P* < 0.001).
+Conseqently, Lycophytes had significantly higher CN ratios than ferns and angiosperms (*P* < 0.001).
+Lycophytes also had 24% lower foliar phosphorous content than angiosperms (*P* = .026), while ferns were simiilar to both groups.
+The NP ratios were not statistically different across angiosperms, ferns and lycophytes in closed canopies habitats.
 
@@
 Variation in stomatal denisty in the shade was largely due to individual species variation.
+
+## Bivariate relationships between functional traits
+
+The relationship between A~n~ and foliar nitrogen content depended on both plant group and canopy type (canopy x plant group, *P* = 0.028).
+Specifically, the postive slopes of the relationship between A~n~ and N differed for ferns and angiosperms in open canopys, while no relationship beween A~n~ and N was found for either group in closed canopies.
+
+fixed vs random...

### 2019-03-05 master_scripts/bivariate_plots.R

@@ -1,19 +1,88 @@
+source("master_scripts/plot_objects.R")
+
+library(plotrix)
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+
 ### bivariate relationships
 
 chem <- read.csv("raw_data/photo_chem.csv")
-stom <- read.csv("raw_data/stomatal_density.csv")
-spad <- read.csv("raw_data/chlorophyll.csv")
 
-dat <- merge(chem, spad, all=TRUE)
+fern <- droplevels(chem[chem$plant_group == "Fern",])
+lyco <- droplevels(chem[chem$plant_group == "Lyco",])
+angio <- droplevels(chem[chem$plant_group == "Angio",])
 
+fernangio <- droplevels(chem[!chem$plant_group == "Lyco",])
 
-library(doBy)
-stom_agg <- summaryBy(sto_den ~ species + individual, FUN=mean, data=stom, 
-                      keep.names = TRUE)
+cols <- c("forestgreen", "dodgerblue", "firebrick4")
+library(scales)
+cols2 <- c(alpha("forestgreen",.7), alpha("dodgerblue",.7), alpha("firebrick4",.7))
 
-dat2 <- merge(dat, stom_agg, all=TRUE)
+##full model for angio and fern
+# anitro_mod <- lm(Photo~n_perc + canopy * plant_group, data=fernangio)
+anitro_mod2 <- lmer(Photo~n_perc + canopy * plant_group + (1|species), 
+                    data=fernangio)
+plot(anitro_mod2)
 
-plot(Photo~n_perc, data= chem, col=plant_group, pch=16)
+library(car)
+# anova(anitro_mod)
+Anova(anitro_mod2, type=3)
+summary(anitro_mod2)
+visreg(anitro_mod2, "plant_group", by="canopy", overlay=TRUE)
+library(emmeans)
+emmip(anitro_mod2, canopy ~ plant_group)
+emmeans(anitro_mod2, pairwise ~ canopy : plant_group)
+r.squaredGLMM(anitro_mod2)
 
+##simple models for ablineclip, paired with full model
+fern_photoN_open <- lm(Photo~n_perc, data=fern[fern$canopy == "Open",])
+fern_open2 <- lmer(Photo~n_perc + (1|species), data=fern[fern$canopy == "Open",])
+Anova(fern_open2) ###yes p = 0.0002
[…]

### 2019-03-05 stats/element_stats_habitats.R

@@ -1,0 +1,143 @@
+## elemental stats (including ratios)
+
+#Stats for overall difference between open vs closed
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+
+
+#photosynthesis
+chem <- read.csv("raw_data/photo_chem.csv")
+
+nitro <- chem[complete.cases(chem$n_perc),]
+  nitro$cn_ratio <- with(nitro, c_perc/n_perc)
+  nitro$percN <- nitro$n_perc /100
+
+phos <- chem[complete.cases(chem$p_perc),]
+  phos$np_ratio <- with(phos, n_perc/p_perc)
+  phos$percP <- phos$p_perc /100
+  
+N_fernangio <- droplevels(nitro[!nitro$plant_group %in% "Lyco", ])
+P_fernangio <- droplevels(phos[!phos$plant_group %in% "Lyco", ])
+
+
+# fern <- droplevels(N_fernangio[N_fernangio$plant_group == "Fern",])
+# length(levels(fern$species))
+# 
+# angio <- droplevels(N_fernangio[N_fernangio$plant_group == "Angio",])
+# length(levels(angio$species))
+
+###N in openclosed -----------------------
+
+N_mod <- lmer(asin(sqrt(percN)) ~ canopy * plant_group + (1|species), 
+              data=N_fernangio)
+plot(N_mod) #asinsqrt better than logit tranformation
+qqPlot(residuals(N_mod)) #looks good
+
+visreg(N_mod, "canopy", by="plant_group")
+
+library(emmeans)
+emmip(N_mod, canopy ~ plant_group)
+emmeans(N_mod, pairwise ~ canopy : plant_group)
+
+Anova(N_mod, type="3")
+summary(N_mod)
+r.squaredGLMM(N_mod)
+#no effect, lots of species variation
+#           R2m       R2c
+#[1,] 0.1052917 0.7267267
+
+mean(N_fernangio[N_fernangio$canopy == "Open" , "percN"])#0.02417179
+mean(N_fernangio[N_fernangio$canopy == "Closed", "percN"])#0.02983398
+#16.2% higher in shade
+
+#only minor effect of habitat
+
+##phosphorous habitat------
+boxplot(percP ~ plant_group, data=P_fernangio) #trim some outliers
+
[…]

### 2019-03-05 stats/element_stats_shade.R

@@ -74,7 +74,7 @@ Pshade_nooutliers <- Pshade[Pshade$p_perc > .1 &
 boxplot(percP~plant_group, data=Pshade_nooutliers)
 
 #arcsin
-shade_P <-lmer(asin(sqrt(percP)) ~ plant_group + (1|species), data=Pshade_nooutliers)
+shade_P <- lmer(asin(sqrt(percP)) ~ plant_group + (1|species), data=Pshade_nooutliers)
 plot(shade_P)
 qqPlot(residuals(shade_P))
 
@@ -95,8 +95,13 @@ tukey_Pshade <- glht(shade_P, linfct = mcp(plant_group = "Tukey"))
 shade_Psiglets <-cld(tukey_Pshade)
 shade_Psiglets2 <- shade_Psiglets$mcletters$Letters
 
-#not different
+library(emmeans)
+emmeans(shade_P, pairwise ~ plant_group)
 
+mean(Pshade[Pshade$plant_group == "Angio", "percP"]) #0.0025
+mean(Pshade[Pshade$plant_group == "Fern", "percP"]) #0.0021
+mean(Pshade[Pshade$plant_group == "Lyco", "percP"]) #0.0019
+#24%lower in lycopphytes thatn angiosperms
 
 ###cn ratio----
 cn_shade <- droplevels(Nshade[complete.cases(Nshade$cn_ratio), ])
@@ -125,12 +130,15 @@ shade_CNsiglets2 <- shade_CNsiglets$mcletters$Letters
 
 ###NP_SHADE----
 np_shade <- droplevels(Pshade[complete.cases(Pshade$np_ratio), ])
+boxplot(np_ratio~plant_group, data=np_shade)
 
-np_shade_mod <- lmer(sqrt(np_ratio) ~ plant_group + (1|species), data=np_shade)
+np_shade_noout <- np_shade[np_shade$np_ratio < 30,]
+
+np_shade_mod <- lmer(sqrt(np_ratio) ~ plant_group + (1|species), 
+                     data=np_shade_noout)
 plot(np_shade_mod)
 qqPlot(residuals(np_shade_mod))
 
-##outlier
 Anova(np_shade_mod, type = "3")
 
 summary(np_shade_mod)
@@ -141,3 +149,6 @@ tukey_NPshade <- glht(np_shade_mod, linfct = mcp(plant_group = "Tukey"))
 shade_NPsiglets <-cld(tukey_NPshade)
[…]

### 2019-03-05 stats/wue_stats.R

@@ -10,17 +10,17 @@ library(multcomp)
 
 #raw data
 photo <- read.csv("raw_data/photo_chem.csv")
-  #calculate PNUE on am area basis as we do not have sla
-  photo$pnue <- with(photo, Photo/n_perc)
+#calculate PNUE on am area basis as we do not have sla
+photo$wue <- with(photo, Photo/Trmmol)
 
-boxplot(pnue ~ plant_group, data=photo)
+boxplot(wue ~ plant_group, data=photo)
 
-pnue <- photo[complete.cases(photo$pnue),]
+wue <- photo[complete.cases(photo$wue),]
 
 ###comparative stats-----------------------
-shade <- droplevels(pnue[pnue$canopy == "Closed",])
+shade <- droplevels(wue[wue$canopy == "Closed",])
 
-shade_mod <-lmer(pnue~ plant_group + (1|species), data=shade)
+shade_mod <-lmer(sqrt(wue)~ plant_group + (1|species), data=shade)
 plot(shade_mod)
 qqPlot(residuals(shade_mod))
 
@@ -29,32 +29,38 @@ summary(shade_mod)
 r.squaredGLMM(shade_mod) 
 visreg(shade_mod)
 
+emmeans(shade_mod, pairwise ~ plant_group )
+
 tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
 shade_siglets <-cld(tukey_shade)
 shade_siglets2 <- shade_siglets$mcletters$Letters
 
+mean(shade[shade$plant_group == "Lyco" , "wue"])#8.844896
+mean(shade[!shade$plant_group == "Lyco", "wue"])#4.691448
+#47.0% higher in lyco
+#0.005176 
 
 #canopy comparisions of sun shade with angio and ferns------
-fernangio <- droplevels(pnue[!pnue$plant_group %in% "Lyco", ])
-boxplot(pnue ~ species, data=fernangio)
+fernangio <- droplevels(wue[!wue$plant_group %in% "Lyco", ])
+boxplot(wue ~ plant_group, data=fernangio)
 
 #interaction between canopy and plant group
-pnue_mod2 <- lmer(sqrt(pnue) ~ canopy * plant_group + (1|species), data=fernangio)
-plot(pnue_mod2)
-qqPlot(residuals(pnue_mod2))
[…]

### 2019-03-07 bivariate_basic.R
new plots

@@ -1,0 +1,81 @@
+source("master_scripts/plot_objects.R")
+
+library(plotrix)
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+
+### bivariate relationships
+
+chem <- read.csv("raw_data/photo_chem.csv")
+  chem$wue <- with(chem, Photo/Trmmol)
+  
+stom <- read.csv("raw_data/stomatal_density.csv")
+stom_agg <- doBy::summaryBy(sto_den ~ species + individual, data=stom, 
+                            FUN=mean, keep.names = TRUE)
+chem2 <- merge(chem,stom_agg)
+
+fern <- droplevels(chem[chem$plant_group == "Fern",])
+lyco <- droplevels(chem[chem$plant_group == "Lyco",])
+angio <- droplevels(chem[chem$plant_group == "Angio",])
+
+#open closed (no lyco)
+fernangio <- droplevels(chem[!chem$plant_group == "Lyco",])
+#closed only (all groups)
+shade <- droplevels(chem[chem$canopy=="Closed",])
+
+cols <- c("forestgreen", "dodgerblue", "firebrick4")
+library(scales)
+cols2 <- c(alpha("forestgreen",.7), alpha("dodgerblue",.7), alpha("firebrick4",.7))
+leglab <- c("Angio-Closed", "Angio-Open","Ferns-Closed", "Ferns-Open")
+leglab_shade <- c("Angiosperms", "Ferns", "Lycophytes")
+
+#photosynthesis vs nitrogen-----
+
+windows()
+par(mar=c(5,5,1,1))
+plot(Photo~n_perc, data=chem, ylab=photolab, xlab = "Foliar Nitrogen  (%)",
+     type='n', xlim=c(0,5), ylim=c(0,25))
+points(Photo~n_perc, data=angio, pch=c(21,22)[canopy], bg=cols2[1], cex=1.25)
+points(Photo~n_perc, data=fern, pch=c(21,22)[canopy], bg=cols2[2], cex=1.25)
+legend("topleft", leglab, pch=c(21, 22, 21, 22),
+       pt.bg=c(rep(cols2[1],2), rep(cols2[2],2)), bty='n')
+
+##i think: no relationship in shade, different postive slopes angio vs fern open
+windows()
+par(mar=c(5,5,1,1))
[…]

### 2019-03-08 master_scripts/interaction_plots.R
started new figure for interactions

@@ -1,0 +1,56 @@
+source("master_scripts/plot_objects.R")
+##interaction plots for canopy comparisons
+
+chem <- read.csv("raw_data/photo_chem.csv")
+
+fernangio <- droplevels(chem[!chem$plant_group == "Lyco",])
+
+canopycols <- c("forestgreen", "dodgerblue")
+
+library(doBy)
+library(plotrix)
+photo <- fernangio[complete.cases(fernangio$Photo),]
+
+photo_agg <- summaryBy(Photo ~ plant_group + canopy, FUN=c(mean, std.error), 
+                       data=photo)
+photo_agg$xaxis <- c(1,2,1,2)
+
+
+cond <- fernangio[complete.cases(fernangio$Cond),]
+cond_agg <- summaryBy(Cond ~ plant_group + canopy, FUN=c(mean, std.error), 
+                       data=cond)
+cond_agg$xaxis <- c(1,2,1,2)
+
+#photosynthesis (add model stats)
+windows()
+par(mar=c(5,5,1,1))
+plot(Photo.mean ~ xaxis, data=photo_agg, ylim=c(0,15), xlim=c(0.5,2.5), xaxt='n',
+      type='n', ylab=photolab, xlab="")
+with(photo_agg, arrows(xaxis, Photo.mean, xaxis, 
+                           Photo.mean+Photo.std.error, angle=90, 
+                           col=canopycols[plant_group],length=0.03, lwd=2))
+with(photo_agg, arrows(xaxis, Photo.mean, xaxis, 
+                       Photo.mean-Photo.std.error, angle=90, 
+                       col=canopycols[plant_group],length=0.03, lwd=2))
+points(Photo.mean ~ xaxis, data=photo_agg, cex=2.5, col=canopycols[plant_group], 
+       pch=19)
+lines(x=c(1,2), y=c(photo_agg[3,3],photo_agg[4,3]), lwd=2.5, col="dodgerblue")
+lines(x=c(1,2), y=c(photo_agg[1,3],photo_agg[2,3]), lwd=2.5, col="darkgreen")
+axis(1, at=1:2, labels = c("Closed", "Open"))
+
+#cond (add model stats)
+windows()
+par(mar=c(5,5,1,1))
+plot(Cond.mean ~ xaxis, data=cond_agg, ylim=c(0,.35), xlim=c(0.5,2.5), xaxt='n',
+     type='n', ylab=photolab, xlab="")
[…]

### 2019-03-11 manuscript/manuscript.Rmd
pca plots drafted

@@
 ---
-title: "A Comparative Ecophysiological Survey of Lycophytes, Ferns, and Seed Plants"
+title: "Adaptability of ecophysiological traits define how plant lineages make a living in shaded environments "
 author: 
@@
 
-```{r setup, include=FALSE}
-knitr::opts_knit$set(root.dir = '../')
+\
+Courtney E. Campany^1^, coauthors^1^ and Eddie Watkins Jr. ^1^
+
+\
+^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
+
+\
+Corresponding author: C.E. Campany, email: courtneycampany@gmail.com, telephone: +1 607 279 7020
+
+Running Head: Shade ecophysiology of plant lineages
+
+```{r, results="asis", echo=FALSE, warning=FALSE}
+library(knitr)
+opts_knit$set(root.dir = '../')
+```
+
+```{r global data/sourcing, echo=FALSE, message=FALSE, warning=FALSE}
+# Set Flag to include figures in text or not.
+includeFigs <- FALSE
 ```
@@
 
-# Stoichiometry (##can use pairwise for specific p values: see leaf P below)
+## Stoichiometry (##can use pairwise for specific p values: see leaf P below)
 
@@
 fixed vs random...
+
+# Discussion
+
+# Acknowledgements
+We thank
+
+# Tables {.unnumbered}
+**Table 1**.
+
+\
+**Table 2**.
+
+# Figures {.unnumbered}
+
+```{r allometry, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs}
+source("master_scripts/pca_plot.R")
+```  
+**Figure 1**.
+Redundancy analysis of measured functional traits across con-occuring angiosperms and fern species in open and shaded habitats.
+With the 12 measured variables, number% of the trait variation was accounted for in two dimensions.
+Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($$\phi$$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content, the leaf carbon to nitrogen ratio (C:N) and the leaf nitrogen to phosphorus ratio (N:P).
+
+\
+
+```{r allometry, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs}
+source("master_scripts/pca_canopy.R")
+```  
+**Figure 2**. Redundancy analysis of measured functional traits across con-occuring angiosperms fern and lycopyte species in shaded habitats.
+
+# References {.unnumbered}

### 2019-03-11 master_scripts/pca_canopy.R

@@ -1,3 +1,5 @@
+source("master_scripts/plot_objects.R")
+
 ##merge all datasets into master for PCA
 library(dplyr)
 library(stringr)
@@ -35,14 +37,14 @@ hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
 
 #select data for PCA
 hyuck_nona <- hyuck[complete.cases(hyuck),]
-hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
-hyuck_nona$CN <- with(hyuck_nona, c_perc/ n_perc)
-hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
+  hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
+  hyuck_nona$CN <- with(hyuck_nona, c_perc/ n_perc)
+  hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
 
 hyuck_canopy <- hyuck_nona[!hyuck_nona$plant_group == "Lycophyte",]  
 
-#get rid of some parameters
-hyuck_pca <- droplevels(hyuck_canopy[,-c(1:2,6:7,9:13,16,19:21)])
+#get rid of some parameters (Asat is from aqcurve fitting so keep Photo)
+hyuck_pca <- droplevels(hyuck_canopy[,-c(1:3,6:7,9:13,16,19:21)])
 
 library(scales)
 #plant group colors
@@ -72,28 +74,28 @@ library(vegan)
 #principle compoent analysis with scales variances
 hyuck_rda<- rda(hyuck_pca,scale=T)
 # plot(hyuck_rda)
-summary(hyuck_rda)
-
-#nicer plot
+# summary(hyuck_rda)
 
 len <- .8
 
 sites <- scores(hyuck_rda, display='sites')
 spp <- scores(hyuck_rda, display='species')
 
-row.names(spp) <- c("An", "PHI","Rd", "LCP", "Photo", "N", "P", "gs", 
-                    "SD", "WUE", "CN", "NP")
-
 #plotting compare ferns and angio lyco in shade
-windows()
+# windows()
 par(mar=c(5,5,2,2), las=1,cex.axis=0.8)
-plot(sites,ylab="PC 2 (20.6 %)", xlab="PC 1 (36.4%)",type='n',
+plot(sites,ylab="PC 2 (21.8 %)", xlab="PC 1 (32.9%)",type='n',
[…]

### 2019-03-11 master_scripts/pca_plot.R

@@ -1,3 +1,5 @@
+source("master_scripts/plot_objects.R")
+
 ##merge all datasets into master for PCA
 library(dplyr)
 library(stringr)
@@ -38,16 +40,21 @@ hyuck_nona <- hyuck[complete.cases(hyuck),]
   hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
   hyuck_nona$CN <- with(hyuck_nona, c_perc/ n_perc)
   hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
+  hyuck_nona$seed <- as.factor(ifelse(hyuck_nona$plant_group == "Angiosperm",
+                            "seed","non-seed"))
   
 hyuck_shade <- hyuck_nona[hyuck_nona$canopy == "Closed",]  
   
-#get rid of some parameters
-hyuck_pca <- droplevels(hyuck_shade[,-c(1:2,6:7,9:13,16,19:21)])
+#get rid of some parameters (Asat is from aqcurve fitting so keep Photo)
+hyuck_pca <- droplevels(hyuck_shade[,-c(1:3,6:7,9:13,16,19:21,26)])
 
 library(scales)
 #plant group colors
 plantcols <- c(alpha("dodgerblue", .8), alpha("firebrick", .8), 
                alpha("forestgreen", .8))
+
+
+pgcols <- c(plantcols[2], plantcols[3], plantcols[1])
 
 #site variables for ease with ponts in pca
 hyuck_id <- hyuck_shade[,12:13]
@@ -58,7 +65,7 @@ hyuck_id$plantcols <- ifelse(hyuck_id$plant_group == "Angiosperm", plantcols[2],
 hyuck_id$plantcols <- ifelse(hyuck_id$plant_group == "Fern", plantcols[3],
                              hyuck_id$plantcols)
 
-#length of shade and id dfrs should be same!
+#double chekc that length of shade and id dfrs should be same!
 
 library(vegan)
 
@@ -76,28 +83,51 @@ hyuck_rda<- rda(hyuck_pca,scale=T)
 # plot(hyuck_rda)
 # summary(hyuck_rda)
 
-#nicer plot
-
+#manuscript plot
 len <- .8
 
 library(scales)
 sites <- scores(hyuck_rda, display='sites')
[…]

### 2019-03-11 master_scripts/plot_objects.R

@@ -1,7 +1,12 @@
 #plotobjects
 
 condlab <- expression(italic(g)[s]~~(mol~m^-2~s^-1))
+condlab2 <- expression(italic(g)[s])
+
 photolab <- expression(italic(A[n])~~(mu*mol~m^-2~s^-1))
+photolab2 <- expression(italic(A[n]))
+
+resplab <- expression(italic(R[d]))
 
 narealab <- expression(Leaf~Nitrogen[area]~~(mg~m^-2))
 parealab <- expression(Leaf~Phosphorous[area]~~(mg~m^-2))
@@ -10,3 +15,9 @@ cilab <- expression(italic(C)[i]~~(ppm))
 
 lcplab <- expression(LCP~~(mu*mol~m^-2~s^-1))
 denslab <- expression(Stomatal~Density~~(mm^-2))
+
+philab <- expression(phi)
+
+#list of labels for pca plots
+pcalabs <- c(philab, resplab,"LCP", photolab2, "N", "P", condlab2, "SD", "WUE", 
+             "C:N", "N:P")

### 2019-03-11 manuscript/manuscript.Rmd
new plots

@@
 # Set Flag to include figures in text or not.
-includeFigs <- FALSE
+includeFigs <- TRUE
 ```
+
+/
 
@@
 # Tables {.unnumbered}
-**Table 1**.
+
+**Table 1**. Species list.
+
+```{r speciestable, echo=FALSE,warning=FALSE, message=FALSE}}
+library(pixiedust)
+speciestab <- read.csv("raw_data/species_list.csv")
+
+dust(speciestab[1:4])%>%
+  sprinkle_colnames("Species", "Plant Lineage", "Collection Site", 
+                    "Canopy") %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
 
@@
 
-```{r allometry, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs}
+```{r pca1, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs,warning=FALSE, message=FALSE}
 source("master_scripts/pca_plot.R")
@@
 With the 12 measured variables, number% of the trait variation was accounted for in two dimensions.
-Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($$\phi$$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content, the leaf carbon to nitrogen ratio (C:N) and the leaf nitrogen to phosphorus ratio (N:P).
+Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content, the leaf carbon to nitrogen ratio (C:N) and the leaf nitrogen to phosphorus ratio (N:P).
 
@@
 
-```{r allometry, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs}
+```{r pca2, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs,warning=FALSE, message=FALSE}
 source("master_scripts/pca_canopy.R")
@@
 
+```{r interaction, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
+source("master_scripts/interaction_plots.R")
+```  
+**Figure 3**. Differences in mean physiological traits between co-occuring angiosperms and fern species in open light and closed habitats.
+
+```{r lyco, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
+source("master_scripts/lycophytes_plot.R")
+```  
+**Figure 4**. Differences in mean physiological traits between lycophytes species that are constrained to only shade habitats and ferns and angiosperms.
+
+```{r seed, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
+source("master_scripts/seed_nonseed_plots.R")
+```  
+**Figure 5**. Differences in mean physiological traits between co-occuring seed (angiosperms) and non-seed plant species (ferns + lycophytes) in shade habitats.
+
 # References {.unnumbered}

### 2019-03-11 AQ_plot.R

@@ -1,6 +1,0 @@
-#plot LRC and analyze differences
-
-lrc <- read.csv("calculated_data/AQ_params.csv")
-
-
-boxplot

### 2019-03-11 functions/basic_functions.R

@@ -1,0 +1,273 @@
+###mtexti for reading direction arrangement
+mtexti <- function(text, side, off = 0.25,
+                   srt = if(side == 2) 90  else
+                     if(side == 4) 270 else 0, 
+                   outer=TRUE,cex=1,
+                   ...) {
+  # dimensions of plotting region in user units
+  usr <- par('usr')
+  # dimensions of plotting region in inches
+  pin <- par('pin')
+  # user units per inch
+  upi <- c(usr[2]-usr[1],
+           usr[4]-usr[3]) / pin
+  # default x and y positions
+  xpos <- (usr[1] + usr[2])/2
+  ypos <- (usr[3] + usr[4])/2
+  if(1 == side)
+    ypos <- usr[3] - upi[2] * off
+  if(2 == side)
+    xpos <- usr[1] - upi[1] * off
+  if(3 == side)
+    ypos <- usr[4] + upi[2] * off
+  if(4 == side)
+    xpos <- usr[2] + upi[1] * off
+  text(x=xpos, y=ypos, text, xpd=NA, srt=srt, cex=cex,...)
+}
+
+
+###3 functions (fitgam, addpoly, predline, and smooth plot) ----------------------------------------------------------
+
+###fitgam
+fitgam <- function(X,Y,dfr, k=-1, R=NULL){
+  dfr$Y <- dfr[,Y]
+  dfr$X <- dfr[,X]
+  if(!is.null(R)){
+    dfr$R <- dfr[,R]
+    model <- 2
+  } else model <- 1
+  dfr <- droplevels(dfr)
+  
+  
+  if(model ==1){
+    g <- gam(Y ~ s(X, k=k), data=dfr)
+  }
+  if(model ==2){
+    g <- gamm(Y ~ s(X, k=k), random = list(R=~1), data=dfr)
+  }
+  
+  return(g)
+}
+
+
+####addpoly
+addpoly <- function(x,y1,y2,col=alpha("lightgrey",0.75),...){
+  ii <- order(x)
+  y1 <- y1[ii]
+  y2 <- y2[ii]
+  x <- x[ii]
+  polygon(c(x,rev(x)), c(y1, rev(y2)), col=col, border=NA,...)
+}
+
+
+####predline
+predline <- function(fit, from=NULL, to=NULL, ...){
+  
+  if(is.null(from))from <- min(fit$model[,2], na.rm=TRUE)
[…]

### 2019-03-11 master_scripts/bivariate_plots.R

@@ -1,4 +1,5 @@
 source("master_scripts/plot_objects.R")
+source("functions/basic_functions.R")
 
 library(plotrix)
 library(visreg)
@@ -86,3 +87,6 @@ ablineclip(fernopen_fix[[1]], fernopen_fix[[2]],lwd=2,col=cols2[2],
 #angioopen
 ablineclip(angioopen_fix[[1]], angioopen_fix[[2]],lwd=2,col=cols2[1],
            x1=.95, x2=3.66)
+
+#here is the predlines example
+predline(asatgs_mod_fern, col="grey20",lwd=2, lty=2)

### 2019-03-11 master_scripts/interaction_plots.R

@@ -1,56 +1,134 @@
 source("master_scripts/plot_objects.R")
-##interaction plots for canopy comparisons
+source("functions/basic_functions.R")
 
-chem <- read.csv("raw_data/photo_chem.csv")
-
-fernangio <- droplevels(chem[!chem$plant_group == "Lyco",])
-
-canopycols <- c("forestgreen", "dodgerblue")
-
+library(dplyr)
+library(stringr)
 library(doBy)
 library(plotrix)
+
+treatments <- read.csv("raw_data/species_list.csv")
+
+##interaction plots for canopy comparisons
+
+#gas exchange data
+chem <- read.csv("raw_data/photo_chem.csv")
+fernangio <- droplevels(chem[!chem$plant_group == "Lyco",])
+
+#photosyn means
 photo <- fernangio[complete.cases(fernangio$Photo),]
 
 photo_agg <- summaryBy(Photo ~ plant_group + canopy, FUN=c(mean, std.error), 
                        data=photo)
-photo_agg$xaxis <- c(1,2,1,2)
+  photo_agg$xaxis <- c(1,2,1,2)
 
-
+#stomal conductance means
 cond <- fernangio[complete.cases(fernangio$Cond),]
-cond_agg <- summaryBy(Cond ~ plant_group + canopy, FUN=c(mean, std.error), 
-                       data=cond)
-cond_agg$xaxis <- c(1,2,1,2)
+  cond_agg <- summaryBy(Cond ~ plant_group + canopy, FUN=c(mean, std.error), 
+                      data=cond)
+  cond_agg$xaxis <- c(1,2,1,2)
 
-#photosynthesis (add model stats)
-windows()
-par(mar=c(5,5,1,1))
-plot(Photo.mean ~ xaxis, data=photo_agg, ylim=c(0,15), xlim=c(0.5,2.5), xaxt='n',
-      type='n', ylab=photolab, xlab="")
+#light response curves removing one bad curve
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+  mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+         individual =  as.numeric(str_remove(ID, ".*-"))) %>%
+  merge(treatments[,c(2,4:5)]) %>%
[…]

### 2019-03-11 master_scripts/lycophytes_plot.R

@@ -1,0 +1,45 @@
+source("master_scripts/plot_objects.R")
+source("functions/basic_functions.R")
+
+##lycophytes differ
+
+chem <- read.csv("raw_data/photo_chem.csv")
+library(vioplot)
+
+chem_closed <- chem[chem$canopy == "Closed",]
+  chem_closed$cnratio <- with(chem_closed, c_perc/n_perc)
+  chem_closed$wue <- with(chem_closed, Photo/Trmmol)
+
+phos <- chem_closed[chem_closed$p_perc > .1 & chem_closed$p_perc < .4,]
+
+# windows()
+par(mfrow=c(2,2), mgp=c(3,.75,0))
+
+#N content
+par(mar=c(0,5.5,1,0))
+boxplot(n_perc ~ plant_group, data=chem_closed, ylab="Foliar N content  (%)",
+        col=pgcols, ylim=c(0,5), xaxt='n', yaxt='n', outline=FALSE)
+axis(2, las=1)
+
+#P content
+par(mar=c(0,0,1,5.5))
+boxplot(p_perc ~ plant_group, data=phos, ylab="",
+        col=pgcols, ylim=c(0,.425), xaxt='n', yaxt='n')
+axis(4,las=1)
+mtexti("Foliar P content  (%)", 4, outer=TRUE, cex=1, off=.6)
+
+#cn ratio
+par(mar=c(5.5,5.5,0,0))
+boxplot(cnratio ~ plant_group, data=chem_closed, ylab="Foliar C:N ratio",yaxt='n',
+        names=plants,col=pgcols, ylim=c(0,45), outline=FALSE)
+axis(2, las=1)
+
+#water use efficiency
+par(mar=c(5.5,0,0,5.5))
+boxplot(wue ~ plant_group, data=chem_closed, ylab="",names=plants,
+        col=pgcols, ylim=c(0,21), outline=FALSE, yaxt='n')
+axis(4,las=1)
+mtexti(wuelab, 4, outer=TRUE, cex=1, off=.6)
+
+# dev.copy2pdf(file= "output/lycophytes.pdf")
+# dev.off()

### 2019-03-11 master_scripts/lycophytes_plot2.R

@@ -1,0 +1,44 @@
+source("master_scripts/plot_objects.R")
+source("functions/basic_functions.R")
+
+##lycophytes differ
+
+chem <- read.csv("raw_data/photo_chem.csv")
+  chem$cnratio <- with(chem, c_perc/n_perc)
+  chem$wue <- with(chem, Photo/Trmmol)
+
+phos <- chem[chem$p_perc > .1 & chem$p_perc < .4,]
+
+
+##panel plot with non-lycos in both habitats
+windows()
+par(mfrow=c(2,2), mgp=c(3,.75,0))
+
+#N content
+par(mar=c(0,5.5,1,0))
+boxplot(n_perc ~ plant_group, data=chem, ylab="Foliar N content  (%)",
+        col=pgcols, ylim=c(0,5), xaxt='n', yaxt='n', outline=FALSE)
+axis(2, las=1)
+
+#P content
+par(mar=c(0,0,1,5.5))
+boxplot(p_perc ~ plant_group, data=phos, ylab="",outline=FALSE,
+        col=pgcols, ylim=c(0,.425), xaxt='n', yaxt='n')
+axis(4,las=1)
+mtexti("Foliar P content  (%)", 4, outer=TRUE, cex=1, off=.6)
+
+#cn ratio
+par(mar=c(5.5,5.5,0,0))
+boxplot(cnratio ~ plant_group, data=chem, ylab="Foliar C:N ratio",yaxt='n',
+        names=plants,col=pgcols, ylim=c(0,45), outline=FALSE)
+axis(2, las=1)
+
+#water use efficiency
+par(mar=c(5.5,0,0,5.5))
+boxplot(wue ~ plant_group, data=chem, ylab="",names=plants,
+        col=pgcols, ylim=c(0,21), outline=FALSE, yaxt='n')
+axis(4,las=1)
+mtexti(wuelab, 4, outer=TRUE, cex=1, off=.6)
+
+dev.copy2pdf(file= "output/lycophytes2.pdf")
+dev.off()

### 2019-03-11 master_scripts/pca_canopy.R

@@ -46,16 +46,12 @@ hyuck_canopy <- hyuck_nona[!hyuck_nona$plant_group == "Lycophyte",]
 #get rid of some parameters (Asat is from aqcurve fitting so keep Photo)
 hyuck_pca <- droplevels(hyuck_canopy[,-c(1:3,6:7,9:13,16,19:21)])
 
-library(scales)
-#plant group colors
-plantcols <- c(alpha("firebrick", .8), alpha("forestgreen", .8))
-
 #site variables for ease with ponts in pca
 hyuck_id <- hyuck_canopy[,12:13]
 
 ##with alpha
-hyuck_id$plantpcols <- ifelse(hyuck_id$plant_group == "Angiosperm", plantcols[1],
-                             plantcols[2])
+hyuck_id$plantcols <- ifelse(hyuck_id$plant_group == "Angiosperm", plantcols[2],
+                             plantcols[3])
 hyuck_id$canopypch <- ifelse(hyuck_id$canopy == "Closed", 16,1)
 
 #length of shade and id dfrs should be same!
@@ -82,6 +78,7 @@ sites <- scores(hyuck_rda, display='sites')
 spp <- scores(hyuck_rda, display='species')
 
 #plotting compare ferns and angio lyco in shade
+
 # windows()
 par(mar=c(5,5,2,2), las=1,cex.axis=0.8)
 plot(sites,ylab="PC 2 (21.8 %)", xlab="PC 1 (32.9%)",type='n',
@@ -90,11 +87,11 @@ abline(v=0, lty='dashed')
 abline(h=0, lty='dashed')
 ordihull(hyuck_rda, groups = hyuck_canopy$canopy, lwd=2,draw='polygon',
          col="grey",alpha=50, border = "black" ,lty=c(1,3))
-points(sites,cex=1.75, pch=hyuck_id$canopypch, col=hyuck_id$plantpcols)
+points(sites,cex=1.75, pch=hyuck_id$canopypch, col=hyuck_id$plantcols)
 arrows(0, 0, len * spp[, 1],  len * spp[, 2], length = 0.05,lwd=1.5)
 text(spp,labels=pcalabs,cex=1)
 legend("bottomright", legend= c("Angiosperms", "Ferns", "Closed", "Open"),
-       pch=c(16,16, 16, 1), col=c(plantcols[1], plantcols[2], "black", "black"),
[…]

### 2019-03-11 master_scripts/pca_plot.R

@@ -48,13 +48,6 @@ hyuck_shade <- hyuck_nona[hyuck_nona$canopy == "Closed",]
 #get rid of some parameters (Asat is from aqcurve fitting so keep Photo)
 hyuck_pca <- droplevels(hyuck_shade[,-c(1:3,6:7,9:13,16,19:21,26)])
 
-library(scales)
-#plant group colors
-plantcols <- c(alpha("dodgerblue", .8), alpha("firebrick", .8), 
-               alpha("forestgreen", .8))
-
-
-pgcols <- c(plantcols[2], plantcols[3], plantcols[1])
 
 #site variables for ease with ponts in pca
 hyuck_id <- hyuck_shade[,12:13]

### 2019-03-11 master_scripts/plot_objects.R

@@ -7,6 +7,7 @@ photolab <- expression(italic(A[n])~~(mu*mol~m^-2~s^-1))
 photolab2 <- expression(italic(A[n]))
 
 resplab <- expression(italic(R[d]))
+resplab2 <- expression(italic(R[d])~~(mu*mol~m^-2~s^-1))
 
 narealab <- expression(Leaf~Nitrogen[area]~~(mg~m^-2))
 parealab <- expression(Leaf~Phosphorous[area]~~(mg~m^-2))
@@ -17,7 +18,24 @@ lcplab <- expression(LCP~~(mu*mol~m^-2~s^-1))
 denslab <- expression(Stomatal~Density~~(mm^-2))
 
 philab <- expression(phi)
+philab2 <- expression(paste(phi, "  (mol ", " CO" ["2"], "  m",ol^-1, " photons)"))
+
+wuelab <- expression(atop(WUE, paste("(",mu*mol," CO" ["2"]," mmol ",
+                                     H[2],O^-1,")")))
 
 #list of labels for pca plots
 pcalabs <- c(philab, resplab,"LCP", photolab2, "N", "P", condlab2, "SD", "WUE", 
              "C:N", "N:P")
+
+
+library(scales)
+#plant group colors
+plantcols <- c(alpha("dodgerblue", .8), alpha("firebrick", .8), 
+               alpha("forestgreen", .8))
+plantcols2 <- c("dodgerblue", "firebrick","forestgreen")
+
+
+pgcols <- c("firebrick", "forestgreen", "dodgerblue")
+plants <- c("Angiosperms","Ferns","Lycophytes")
+
+seedcols <- c("white", plantcols2[2])

### 2019-03-11 master_scripts/seed_nonseed_plots.R

@@ -1,0 +1,58 @@
+source("master_scripts/plot_objects.R")
+source("functions/basic_functions.R")
+
+treatments <- read.csv("raw_data/species_list.csv")
+
+##lycophytes differ
+
+#gas exchange data
+chem <- read.csv("raw_data/photo_chem.csv")
+
+chem_closed <- chem[chem$canopy == "Closed",]
+  chem_closed$seed <- ifelse(chem_closed$plant_group == "Angio", 
+                             "Seed", "Non-seed")
+  
+#light response curves removing one bad curve
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+    mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+           individual =  as.numeric(str_remove(ID, ".*-"))) %>%
+    merge(treatments[,c(2,4:5)]) %>%
+    filter(!ID ==  "imp_pal-1")
+  
+lrc_shade <-  droplevels(lrc[!lrc$canopy == "Closed",])
+  lrc_shade$seed <- ifelse(lrc_shade$plant_group == "Angiosperm", 
+                           "Seed", "Non-seed")
+
+
+#seed vs nonseed panel figure
+# windows()
+par(mfrow=c(2,2), mgp=c(3,.75,0))
+
+#stomatal conductance
+par(mar=c(0,5,1,0))
+boxplot(Cond ~ seed, data=chem_closed, ylab=condlab,xaxt='n',yaxt='n',
+        ylim=c(0,.3), outline=FALSE, col=seedcols)
+axis(2, las=1)
+
+#quantum yield
+par(mar=c(0,0,1,5))
+boxplot(Phi ~ seed, data=lrc_shade, ylab="", xaxt='n', yaxt='n',
+        ylim=c(0.04,.09),  col=seedcols)
+axis(4,las=1)
+mtexti(philab2, 4, outer=TRUE, cex=1, off=.65)
+
+#light compensation point
+par(mar=c(5,5,0,0))
+boxplot(LCP ~ seed, data=lrc_shade, ylab=lcplab,outline=FALSE,yaxt='n',
+        ylim=c(0.,27),  col=c("white", plantcols2[2]))
+axis(2, las=1)
+
+#dark respiration
+par(mar=c(5,0,0,5))
+boxplot(Rd ~ seed, data=lrc_shade, ylab="", outline=FALSE,
[…]

### 2019-03-13 manuscript/manuscript.Rmd
data table and a vs n plots

@@
 
-```{r speciestable, echo=FALSE,warning=FALSE, message=FALSE}}
+```{r, speciestable, echo=FALSE,warning=FALSE, message=FALSE}
 library(pixiedust)
@@
 \
-**Table 2**.
+**Table 2**. Pooled ecophysiological traits for different plant lineages in open light and closed habitats.
+
+```{r, datatable, echo=FALSE,warning=FALSE, message=FALSE}
+traitstab <- read.csv("manuscript/pooled_data.csv")
+physymb <- '\u03D5'
+
+dust(traitstab)%>%
+  sprinkle_colnames("Plant Lineage", "Canopy", "*A*~n~","*g*~s~","Foliar N",
+                    "Foliar P", "C:N", "N:P", "SD", physymb, "*R*~d~", "LCP") %>%
+  sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle_print_method("markdown")
+```
 
@@
 
+```{r nitro, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
+source("master_scripts/bivariate_plots.R")
+```  
+**Figure 6**. Postive relationship between photosynthesis and foliar nitrogen content exists for angiosperms and fern species in open light habitat, but not in shade habitats
+
 # References {.unnumbered}

### 2019-03-13 functions/basic_functions.R

@@ -1,3 +1,8 @@
+#mean and se functions with na.rm = true as default
+se <- function(x) sqrt(var(x,na.rm=TRUE)/length(na.omit(x)))
+mean2 <- function(x) mean(x, na.rm=TRUE)
+
+
 ###mtexti for reading direction arrangement
 mtexti <- function(text, side, off = 0.25,
                    srt = if(side == 2) 90  else

### 2019-03-13 master_scripts/bivariate_plots.R

@@ -6,9 +6,9 @@ library(visreg)
 library(car)
 library(lme4)
 library(MuMIn)
+library(emmeans)
 
-### bivariate relationships
-
+### bivariate relationships (only A vs Nitrogen is interesting)
 chem <- read.csv("raw_data/photo_chem.csv")
 
 fern <- droplevels(chem[chem$plant_group == "Fern",])
@@ -17,76 +17,79 @@ angio <- droplevels(chem[chem$plant_group == "Angio",])
 
 fernangio <- droplevels(chem[!chem$plant_group == "Lyco",])
 
-cols <- c("forestgreen", "dodgerblue", "firebrick4")
-library(scales)
-cols2 <- c(alpha("forestgreen",.7), alpha("dodgerblue",.7), alpha("firebrick4",.7))
-
 ##full model for angio and fern
 # anitro_mod <- lm(Photo~n_perc + canopy * plant_group, data=fernangio)
 anitro_mod2 <- lmer(Photo~n_perc + canopy * plant_group + (1|species), 
                     data=fernangio)
-plot(anitro_mod2)
+# plot(anitro_mod2)
 
-library(car)
 # anova(anitro_mod)
-Anova(anitro_mod2, type=3)
-summary(anitro_mod2)
-visreg(anitro_mod2, "plant_group", by="canopy", overlay=TRUE)
-library(emmeans)
-emmip(anitro_mod2, canopy ~ plant_group)
-emmeans(anitro_mod2, pairwise ~ canopy : plant_group)
-r.squaredGLMM(anitro_mod2)
+# Anova(anitro_mod2, type=3)
+# summary(anitro_mod2)
+# visreg(anitro_mod2, "plant_group", by="canopy", overlay=TRUE)
+
+# emmip(anitro_mod2, canopy ~ plant_group)
+# emmeans(anitro_mod2, pairwise ~ canopy : plant_group)
+# r.squaredGLMM(anitro_mod2)
 
 ##simple models for ablineclip, paired with full model
 fern_photoN_open <- lm(Photo~n_perc, data=fern[fern$canopy == "Open",])
 fern_open2 <- lmer(Photo~n_perc + (1|species), data=fern[fern$canopy == "Open",])
-Anova(fern_open2) ###yes p = 0.0002
-fernopen_fix <- fixef(fern_open2)
+# Anova(fern_open2) ###yes p = 0.0002
[…]

### 2019-03-13 master_scripts/data_table.R

@@ -1,0 +1,92 @@
+source("functions/basic_functions.R")
+library(doBy)
+library(dplyr)
+library(stringr)
+
+treatments <- read.csv("raw_data/species_list.csv")
+
+##photosynthesis and leaf chemistry data-------------
+photochem <- read.csv("raw_data/photo_chem.csv")
+  photochem$cn_ratio <- with(photochem, c_perc/n_perc)
+  photochem$percN <- photochem$n_perc /100
+  photochem$np_ratio <- with(photochem, n_perc/p_perc)
+  photochem$percP <- photochem$p_perc /100
+
+photochem_mean <- summaryBy(Photo+Cond+percN+percP+cn_ratio+np_ratio ~ 
+                              plant_group+canopy, 
+                              data=photochem, FUN=mean2)
+
+photochem_se <- summaryBy(Photo+Cond+percN+percP+cn_ratio+np_ratio ~ 
+                              plant_group+canopy, 
+                            data=photochem, FUN=se)
+
+#stomatal density data-----------
+stom <- read.csv("raw_data/stomatal_density.csv")
+
+#mean for each individual (multiple stomatal counts)
+stom_agg <- doBy::summaryBy(sto_den ~ species + individual + plant_group,
+                            FUN=mean, data=stom, keep.names = TRUE)
+
+stom_dens <- merge(stom_agg, treatments[,c(4:5)])
+
+stom_mean <- doBy::summaryBy(sto_den ~ plant_group + canopy,
+                             FUN=c(mean2, se), data=stom_dens)
+
+
+##aq curve data
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+  mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+         individual =  str_remove(ID, ".*-")) %>%
+  merge(treatments[,c(2,4:5)])
+
+lrc_agg <- doBy::summaryBy(Phi + Rd + LCP ~ plant_group + canopy, 
+                           data=lrc, FUN=c(mean2, se))
+
+
+#format data table
+photo <- data.frame(paste0(sprintf("%2.1f", round(photochem_mean[,3], 1)), 
+                    " (", sprintf("%2.2f", round(photochem_se[,3],2)),")"))
+
+cond <- data.frame(paste0(sprintf("%1.2f", round(photochem_mean[,4], 2)), 
[…]

### 2019-03-15 manuscript/manuscript.Rmd
pca plots no CN

@@
 \
-Courtney E. Campany^1^, coauthors^1^ and Eddie Watkins Jr. ^1^
+Courtney E. Campany^1^, coauthors^1^ and James E. Watkins Jr. ^1^
 
 \
-^1^ Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith, NSW, Australia
+^1^ Department of Biology, Colgate University, Hamilton, NY, USA
 
 \
-Corresponding author: C.E. Campany, email: courtneycampany@gmail.com, telephone: +1 607 279 7020
+Corresponding author: C.E. Campany, email: courtneycampany@gmail.com,
 
+\
 Running Head: Shade ecophysiology of plant lineages
@@
 
-/
-
-1) Differences in ecophysiology of groups in same habitat (seed vs angio)
-2) similarities or differnces between open/closed with ferns and lycophytes
-3) Do these tell us why lycophtes dont go to open ?
-
 # Abstract
@@
 Our understanding of the ecophysiology of ferns and lycophytes is limited.
-However, recent studies have shown that these two groups possess an unusual set of functional traits that likely influence their ecology (Brodribb, Watkins, Pinterman, Franks).
+However, recent studies have shown that these two groups possess an unusual set of functional traits that likely influence their ecology (Brodribb, Watkins, Pinterman, Franks, campany).
 For example, Watkins has demonstrated that the gametophyte generations of many tropical ferns are markedly desiccation-tolerant, with tolerance being closely linked to ecological preference.
@@
 	A primary controlling factor in fern hydraulics is stomatal function.
-Recently, a series of studies have shown that stomatal control of ferns and lycophytes differs from seed plants.
-Early evidence on fern and lycophyte stomata indicates that these groups have inefficient stomatal systems.
-Ongoing work by Brodribb et al.
-is suggesting that ferns and lycophytes may possess a hydropassive mechanism which influences stomatal opening and closing.
-Unlike seed plants, fern and lycophyte stomata fail to close in response to abscisic acid (ABA) addition.
-This significant new discovery suggests that at least two fundamentally different mechanisms of stomatal function have evolved in early plants.
-This, in turn, may impact the differential distribution of ferns and lycophytes relative to seed plants.
+A series of elegeant studies have shown that stomatal control of ferns and lycophytes differs fundamentally from seed plants.
+Evidence on fern and lycophyte stomata indicates that these groups have inefficient stomatal systems, likely possessing a hydropassive mechanism which influences stomatal opening and closing (many cites).
+Specifically, unlike seed plants, fern and lycophyte stomata fail to close in response to abscisic acid (ABA) addition (many cites).
+Additionally, recent evidence also suggests that co-occuring tropical ferns and lycophytes optimize thier ecophysiology differently (Campany).
+Consequently, the evolution or adaptation of physiolgoical traits may impact the differential distribution of ferns,lycophytes and seed plants to eath other.
 
@@
 Therefore, it is difficult to derive widely applicable hypotheses of the ecophysiological behavior of these groups.
-From the few studies we were able to find, ferns and lycophytes appear to have lower photosynthetic rates than seed plants.
-This would suggest that ferns may be at a competitive disadvantage when growing with seed plants.
+From the limited number of present studies, ferns and lycophytes appear to have lower photosynthetic rates than seed plants (cite).
+This would suggest that non-seed plants may be at a competitive disadvantage when growing alongside seed plants.
 However, none of these studies have systematically compared the physiology of these groups when growing in similar habitats.
@@
 	The goal of the current study is to examine the comparative physiology of a number of temperate ferns and lycophytes to angiosperms.
-We specifically examined several physiological parameters in a pairwise comparison between a fern or a lycophyte and an angiosperm in identical habitats.
-Given recent suggestions on inefficient stomatal functions, we hypothesized that ferns would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms.
-This, in turn, would impact aspects of leaf morphology (stomatal density), physiology (chlorophyll concentration), nutrient relations (N and P), and carbon relations.
-We speculate that the differences in physiological function may affect the ecology of these ferns, lycophytes, and angiosperms.
+We specifically examined several physiological parameters in comparison between co-occurring ferns, lycophytes and  angiosperm in identical closed canopy habitats.
+In a pairwise camoparison, we also evaluated the potential of fern and angiosperm species to adapt to open-light habitats.
+Given increasing evidence of inefficient stomatal functions in non-seed plants, we hypothesized that ferns and lycophytes would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms.
+This, in turn, would impact aspects of leaf anatomy (stomatal density), nutrient relations (N and P), and water use efficiency.
+We speculate that the differences in physiological function also affect the ecology of lycophytes in a manner that prevents them from establishing in full sun environments.
 
@@
 The relationship between A~n~ and foliar nitrogen content depended on both plant group and canopy type (canopy x plant group, *P* = 0.028).
-Specifically, the postive slopes of the relationship between A~n~ and N differed for ferns and angiosperms in open canopys, while no relationship beween A~n~ and N was found for either group in closed canopies.
+Specifically, the postive slopes of the relationship between A~n~ and N differed for ferns and angiosperms in open canopys, while no relationship beween A~n~ and N was found for either group in closed canopies (Figure 6).
 
-fixed vs random...
+No relationships.....
 
 # Discussion
+
+## Comparative ecophysiology in the shade
+
+## Differential adaption to open habitats
+
+## Functional ecology of lycophytes
+
+## Ecophysiological trait variation in seed plants
+
+## Summary
 

### 2019-03-15 master_scripts/pca_canopy.R

@@ -38,7 +38,7 @@ hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
 #select data for PCA
 hyuck_nona <- hyuck[complete.cases(hyuck),]
   hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
-  hyuck_nona$CN <- with(hyuck_nona, c_perc/ n_perc)
+  hyuck_nona$PNUE <- with(hyuck_nona, Photo/ n_perc)
   hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
 
 hyuck_canopy <- hyuck_nona[!hyuck_nona$plant_group == "Lycophyte",]  

### 2019-03-15 master_scripts/pca_plot.R

@@ -38,7 +38,7 @@ hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
 #select data for PCA
 hyuck_nona <- hyuck[complete.cases(hyuck),]
   hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
-  hyuck_nona$CN <- with(hyuck_nona, c_perc/ n_perc)
+  hyuck_nona$PNUE <- with(hyuck_nona, Photo/ n_perc)
   hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
   hyuck_nona$seed <- as.factor(ifelse(hyuck_nona$plant_group == "Angiosperm",
                             "seed","non-seed"))
@@ -103,7 +103,7 @@ legend("topleft", legend= c("Angiosperms", "Ferns", "Lycophytes"),
 # dev.copy2pdf(file= "output/pca_shade1.pdf")
 # dev.off()
 
-##same plot but with seed vs nonseed grouping
+#same plot but with seed vs nonseed grouping
 # windows()
 # par(mar=c(5,5,1,1), las=1,cex.axis=0.8)
 # plot(sites,ylab="PC 2 (16.7 %)", xlab="PC 1 (38.7%)",type='n',
@@ -115,12 +115,11 @@ legend("topleft", legend= c("Angiosperms", "Ferns", "Lycophytes"),
 # points(sites,cex=1.75, bg=hyuck_id$plantcols, pch=21)
 # arrows(0, 0, len * spp[, 1],  len * spp[, 2], length = 0.05, lwd=1.5)
 # text(spp,labels=pcalabs,cex=1)
-# legend("topleft", 
+# legend("topleft",
 #        legend= c("Angiosperms", "Ferns", "Lycophytes", "Seed", "Non-seed"),
 #        inset=0.001, bty='n', cex=1,pt.cex=1.25,lty=c(0,0,0,1,1),
 #        pch=c(rep(21,3),NA,NA),pt.bg=c(pgcols), pt.lwd=1.25,
 #        col=c(rep("black",3),pgcols[1],"black"))
-       
+# 
 # dev.copy2pdf(file= "output/pca_shade2.pdf")
 # dev.off()
-                        

### 2019-03-15 master_scripts/plot_objects.R

@@ -25,7 +25,7 @@ wuelab <- expression(atop(WUE, paste("(",mu*mol," CO" ["2"]," mmol ",
 
 #list of labels for pca plots
 pcalabs <- c(philab, resplab,"LCP", photolab2, "N", "P", condlab2, "SD", "WUE", 
-             "C:N", "N:P")
+             "PNUE", "N:P")
 
 
 library(scales)

### 2019-03-15 scripts/AQ_data_clean.R

@@ -1,0 +1,22 @@
+### clean up aq curve data
+treatments <- read.csv("raw_data/species_list.csv")
+
+
+#photosynthesis
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+  mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+         individual =  str_remove(ID, ".*-")) %>%
+  merge(treatments[,c(2,4:5)]) %>%
+  mutate(canopyplant = interaction(canopy, plant_group))
+
+boxplot(LCP ~ species, data=lrc_clean)
+boxplot(Phi ~ species, data=lrc_clean)
+boxplot(Rd ~ species, data=lrc_clean)
+
+#drop some obvious outliers
+outlierLCP <- max(lrc$LCP)
+#lys_cil-2 appears like a bad curve
+
+lrc_clean <- lrc[!lrc$ID %in% c("lys_cil-2","dip_dig-3","eup_pur-5"),]
+
+write.csv(lrc_clean, "calculated_data/AQ_params_clean.csv", row.names = FALSE)

### 2019-03-15 scripts/ladyfern.R

@@ -1,0 +1,21 @@
+##lady fern
+library(dplyr)
+library(stringr)
+
+treatments <- read.csv("raw_data/species_list.csv")
+
+photochem <- read.csv("raw_data/photo_chem.csv")
+
+#no comparison with AQ curves or stomata
+##no phosphorous data
+
+#subset lady fern 
+lady <- droplevels(photochem[photochem$species %in% c("ath_fil_med", "ath_fil_high"),])
+  lady$wue <- with(lady, Photo/Trmmol)
+  lady$cn <- with(lady, c_perc/n_perc)
+
+boxplot(Photo ~ species, data=lady)
+boxplot(Cond ~ species, data=lady)
+boxplot(n_perc ~ species, data=lady)
+boxplot(cn ~ species, data=lady)
+boxplot(wue ~ species, data=lady)

### 2019-03-15 scripts/lightresponse.R

@@ -3,7 +3,7 @@
 lrc <- read.csv("raw_data/lightresponse.csv")
 lrc$id <- paste(lrc$species, lrc$individual, sep="-")
 
-source("light_response_function.R")
+source("functions/light_response_function.R")
 
 ##fit light response curves for every individual
 

### 2019-03-15 stats/lrc_stats.R

@@ -18,6 +18,10 @@ lrc <- read.csv("calculated_data/AQ_params.csv") %>%
        merge(treatments[,c(2,4:5)]) %>%
        mutate(canopyplant = interaction(canopy, plant_group))
 
+lrc_agg <- doBy::summaryBy(LCP ~ plant_group + canopy, data = lrc, FUN=mean)
+boxplot(LCP ~ species, data=lrc)
+boxplot(LCP ~ plant_group, data=lrc)
+
 
 fern <- droplevels(lrc[lrc$plant_group == "Fern", ])
   #length(unique(levels(fern$species)))

### 2019-03-25 manuscript/manuscript.Rmd
pca text

@@
 
+## Ecophysiology among plant lineages and habitats
+
+We first explored how each of 10 measured functional traits varied and were distributed among co-occuring ferns, Angiosperms and lycophytes in closed canopy habitats.
+In total, 52.9 % of the trait variation was accounted for in two dimensions, encompassing different resource economics, physiology and anatomy of the surveryed taxa.
+Interestingly, there was minimal overlap among component scores for angiosperms species compared to spore-bearing ferns and lycophytes.
+Species associations with functional traits were noticeably divergent between flowering plants and spore-bearing speies in shaded habitats (Fig.
+1).
+Specifically, angiosperms species were associated with gas exchange, anatomical and stoichiometric functional traits related to optimizing carboxylation.
+In contrast, lycophytes and fern species were most associated with traits related to maintaining postive carbon balance under low light.
+
+Next, we explored how the same suite of 10 functional traits were distributed among open and closed habitats in both angiosperms and ferns.
+In total, 55.9 % of the trait variation was accounted for in two dimensions.
+Component scores for angiosperms and fern species broadly diverged across open-light and closed canopy habitats, however, associations with functional traits in open-light habitats were stronger for angiosperms (how do I say this?) (Figure 2).
+Specifically, angiosperms appear more plastic in their ability to upregulate traits related to photosynthetic gas exchange with increased light availability.
+
+Overall, these diverging relationships with measured functional traits suggest a highly variable in situ ecophysiology for coexisting angiosperm, fern and lycophyte taxa.
+Indeed, we detected important differences between....
+(Table 2).
+
 ## Light response curve parameters

### 2019-03-25 master_scripts/pca_canopy.R

@@ -38,7 +38,6 @@ hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
 #select data for PCA
 hyuck_nona <- hyuck[complete.cases(hyuck),]
   hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
-  hyuck_nona$PNUE <- with(hyuck_nona, Photo/ n_perc)
   hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
 
 hyuck_canopy <- hyuck_nona[!hyuck_nona$plant_group == "Lycophyte",]  
@@ -81,7 +80,7 @@ spp <- scores(hyuck_rda, display='species')
 
 # windows()
 par(mar=c(5,5,2,2), las=1,cex.axis=0.8)
-plot(sites,ylab="PC 2 (21.8 %)", xlab="PC 1 (32.9%)",type='n',
+plot(sites,ylab="PC 2 (22.3 %)", xlab="PC 1 (33.6%)",type='n',
      xlim=c(-2.25, 2.25), ylim=c(-2.25, 2.25))
 abline(v=0, lty='dashed')
 abline(h=0, lty='dashed')

### 2019-03-25 master_scripts/pca_plot.R

@@ -38,7 +38,6 @@ hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
 #select data for PCA
 hyuck_nona <- hyuck[complete.cases(hyuck),]
   hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
-  hyuck_nona$PNUE <- with(hyuck_nona, Photo/ n_perc)
   hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
   hyuck_nona$seed <- as.factor(ifelse(hyuck_nona$plant_group == "Angiosperm",
                             "seed","non-seed"))
@@ -46,7 +45,7 @@ hyuck_nona <- hyuck[complete.cases(hyuck),]
 hyuck_shade <- hyuck_nona[hyuck_nona$canopy == "Closed",]  
   
 #get rid of some parameters (Asat is from aqcurve fitting so keep Photo)
-hyuck_pca <- droplevels(hyuck_shade[,-c(1:3,6:7,9:13,16,19:21,26)])
+hyuck_pca <- droplevels(hyuck_shade[,-c(1:3,6:7,9:13,16,19:21,25)])
 
 
 #site variables for ease with ponts in pca
@@ -88,7 +87,7 @@ spp <- scores(hyuck_rda, display='species')
 #plotting compare ferns and angio lyco in shade
 # windows()
 par(mar=c(5,5,1,1), las=1,cex.axis=0.8)
-plot(sites,ylab="PC 2 (16.7 %)", xlab="PC 1 (38.7%)",type='n',
+plot(sites,ylab="PC 2 (18.4 %)", xlab="PC 1 (34.5%)",type='n',
      xlim=c(-2.25, 2.25), ylim=c(-2.25, 2.25))
 abline(v=0, lty='dashed')
 abline(h=0, lty='dashed')

### 2019-03-25 master_scripts/plot_objects.R

@@ -24,8 +24,7 @@ wuelab <- expression(atop(WUE, paste("(",mu*mol," CO" ["2"]," mmol ",
                                      H[2],O^-1,")")))
 
 #list of labels for pca plots
-pcalabs <- c(philab, resplab,"LCP", photolab2, "N", "P", condlab2, "SD", "WUE", 
-             "PNUE", "N:P")
+pcalabs <- c(philab, resplab,"LCP", photolab2, "N", "P", condlab2, "SD", "WUE","N:P")
 
 
 library(scales)

### 2019-03-26 manuscript/manuscript.Rmd
edits to data table and removed C:N

--- 2019-03-15 manuscript/manuscript.Rmd
@@
 csl: new-phytologist.csl
-bibliography: references.bib
+bibliography: hyuck_zotero.bib
 ---
@@
 \
-Running Head: Shade ecophysiology of plant lineages
+Running Head: Ecophysiology of herbaceous plant lineages
 
@@
 
-This study was conducted at two sites in upstate New York.
+This study was conducted at two sites in upstate New York, USA.
 Data for most species were gathered at the Edmund Niles Huyck Preserve (ENHP), characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range.
-Data from the second site were gathered at a natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley.
+Data from the second site were gathered at a forested natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley.
 Forests at the Colgate University site are younger, but consist of a similar species richness and composition as ENHP.
@@
 Surveyed angiosperm taxa were always herbaceous and similar in size and stature to the neighboring fern species.
-At the ENHP, we sampled across two distinct habitats: understory low-light areas and open, high-light areas.
-At the Colgate University site, additioanl low-light fern and lycophyte species were surveyed.
+At the ENHP, we sampled across two distinct habitats: understory low-light areas (closed) and open canopy high-light areas (open).
+At the Colgate University site, additional low-light fern and lycophyte species were surveyed.
 
@@
 
-Light response curves were conducted by decreasing the light intensity in the gas exchange cuvette in small step changes.
+Light response curves were conducted by decreasing the light intensity in the 2x3 cm gas exchange cuvette in small step changes.
 For each light response curve, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm).
 Temperature was not explictly controlled in the leaf cuvette.
-Cuvette leaf temperatures ranged from 27-30 &deg;C and leaf vapor pressure deficit (VPD) ranged from ***** kPa.
+Cuvette leaf temperatures ranged from 27-30 &deg;C and leaf vapor pressure deficit (VPD) averaged 1.37&pm;0.40 kPa.
 Initial light measurements were made using the built-in quantum sensor to determine typical ambient light levels.
@@
 For low light species, light response curves started at a PPFD of 500 $\mu$mol m^-1^ s^-1^, followed by 13 additional steps to O $\mu$mol m^-1^ s^-1^.
-Light response data for each curves was fit to a non-rectangular hyperbola model (Equation 6 in Lobo et al.
-2013).
+Light response data for each curves was fit to a non-rectangular hyperbola model (Equation 6 in @lobo_fitting_2013).
 The light compensation point (LCP) was calculated as the PPFD at which the net photosynthetic rate equalled zero from the linear phase of each light response curve.
@@
 Foliar tissue was sampled following gas exchange measurements for all species at ENHP and used for nutrient analyses.
-Samples were dried to a constant mass and ground using a Wig-L-Bug (Sigma-Aldrich Co.
-St.
+Samples were dried to a constant mass and ground using a Wig-L-Bug (Sigma-Aldrich, St.
 Louis, USA).
 Carbon and nitrogen analyses were measured using a Costech Analytical Elemental Analyzer (Valencia, USA), with the percentage of carbon and nitrogen in samples calculated by comparison with certified standards.
-Photosynthetic nitrogen use efficiency (PNUE) was defined as the ratio of A~n~ to leaf nitrogen content.
-Foliar phosphorus concentrations were determined using an ash digestion process (D’Angelo et al., 2001) preceded by colour development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon).
+Foliar phosphorus concentrations were determined using an ash digestion process [@dangelo_rapid_2001] preceded by color development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon, USA).
 
@@
 
-Stomatal density (SD) was measured by directly counting stomata on the abaxial leaf surface under 40x magnification with a field of view of 0.622 mm^2.
+Stomatal density (SD) was measured by directly counting stomata on the abaxial leaf surface under 40x magnification with a field of view of 0.622 mm^2^.
 Stomatal density was calculated from 8 non-overlapping foliar regions for five individuals of each species.
@@
 
+Principal component analysis, utilizing the ‘vegan’ package [@oksanen_vegan:_2019], was used to explore how measured functional traits were distributed and co-varied among plant groups and canopy type.
 Linear mixed-effect models responses were used to test responses of functional traits to caetgorial fixed effects of plant group and canopy habitat.
@@
 To test differences between habitat types (ferns and angiosperms only) species was treated as a fixed effect.
-Explained variance (R^2^) of mixed models were computed as in Nakagawa and Schielzeth (2013), in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
-Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package (Hothorn et al., 2008).
+Explained variance (R^2^) of mixed models were computed as in @nakagawa_general_2013, in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
+Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package [@hothorn_simultaneous_2008].
 T-tests were performed to test for differences between functional traits across canopy habitats within angioperm and fern plant groups.
+For bivariate trait relationships, responses of dependent variables were analysed with linear mixed-effect models, with species as a random effect and plant group and/or canopy type as categorical fixed effects.
+Differences in slopes of the relationship of bivariate traits by habitat were tested by calculating estimated marginal means and computing pairwise comparisons with the ‘emmeans’ package [@lenth_emmeans:_2019].
 All tests of statistical significance were conducted at an α level of 0.05.
-All analyses were performed with R 3.5.1 [R cite].
-
-Principal component analysis, utilizing the ‘vegan’ package (Oksanen et al., 2018), was used to explore how measured functional traits were distributed and co-varied among plant groups and canopy type.
-
-For bivariate trait relationships, responses of dependent variables were analysed with linear mixed-effect models, with species as a random effect and plant group and/or canopy type as categorical fixed effects.
-Differences in slopes of the relationship of bivariate traits by habitat (Selaginella only) were tested by calculating estimated marginal means and computing pairwise comparisons with the ‘emmeans’ package (Length, 2018).
+All analyses were performed with R 3.5.1 [@r_core_team_r:_2018].
 
@@
 
+## Ecophysiology among plant lineages and habitats
+
+We first explored how each of 10 measured functional traits varied and were distributed among co-occuring ferns, angiosperms and lycophytes in closed habitats.
+In total, 52.9 % of the trait variation was accounted for in two dimensions, encompassing different resource economics, physiology and anatomy of the surveryed taxa.
+Interestingly, there was minimal overlap among component scores for angiosperms species compared to spore-bearing ferns and lycophytes (Figure 1).
+Specifically, angiosperms species were associated with gas exchange, anatomical and stoichiometric functional traits related to optimizing carboxylation.
+In contrast, lycophytes and fern species were most associated with physiological traits related to maintaining postive carbon balance.
+
+\
+
+Next, we explored how the same suite of 10 functional traits were distributed among open and closed habitats, in both angiosperms and ferns.
+In total, 55.9 % of the trait variation was accounted for in two dimensions.
+Component scores for angiosperms and fern species broadly diverged across open and closed canopy habitats, however, associations with physiological traits in open habitats were stronger for angiosperms (Figure 2).
+Specifically, angiosperms appear more plastic in their ability to upregulate traits related to photosynthetic gas exchange with increased light availability.
+Overall, these diverging relationships with measured functional traits suggest a highly variable in situ ecophysiology for coexisting angiosperm, fern and lycophyte taxa.
+This is evident in broad differences in mean values of ecophyioslogical traits across the three plant lineages and among habitat types (Table 2).
+
 ## Light response curve parameters
 
-Overall, ferns and angiosperms had similar $\phi$ across open and low-light habitats.
-Quantum yields were also equivalent within angiosperm and fern species groups between open and low-light habitats.
-Shifts in R~d~ between open and low-light habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001).
-Specifically, R~d~ in angiosperms was 80.6 % greater in open compared to low-light habitats (t=-4.197, *P* = 0.001).
-Alternatively, R~d~ in ferns was 44.0 % greater in low-light compared to open habitats (t=-2.97, p=0.006).
-Similarly, shifts in LCP between open and low-light habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001).
-Specifically, the LCP in angiosperms was 79.2 % greater in open compared to low-light habitats (t=-3.5998, *P* = 0.005).
-Alternatively, the LCP in ferns was 39.7 % greater in low-light compared to open habitats (t=2.1814, *P* = 0.040).
+Overall, ferns and angiosperms had similar $\phi$ across open and closed habitats.
+Quantum yields were also equivalent within angiosperm and fern species groups between open and closed habitats.
+Shifts in R~d~ between open and closed habitats differed for angiosperms and ferns (habitat x plant group, *P* < 0.001).
+Specifically, R~d~ in angiosperms was 80.6 % greater in open compared to closed habitats (t=-4.197, *P* = 0.001).
+Alternatively, R~d~ in ferns was 44.0 % greater in low-light compared to open habitats (t=-2.97, *P* = 0.006).
+Similarly, shifts in LCP between open and closed habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001).
+Specifically, the LCP in angiosperms was 79.2 % greater in open compared to closed habitats (t=-3.5998, *P* = 0.005).
+Alternatively, the LCP in ferns was 39.7 % greater in closed compared to open habitats (t=2.1814, *P* = 0.040).
 
@@
 Ferns and angiosperms species in open light environments had marginally lower foliar nitrogen content compared to similar species in growing in closed light environments (-17.2 %, *P* = 0.072), with a large proportion of variation attributed to individual species differences.
-Neither ferns nor angiosperms species differed in their CN ratios across the two light environments.
-Across all habitats, the CN ratio of ferns was 20 % higher than in angiosperms (*P* = 0.0325).
 No differences in foliar P content were detected between open and closed canopies for either ferns or angiosperms.
@@
 Neither ferns nor angiosperms species differed in their NP ratios across the two light environments.
-Across all habitats, the NP ratio was statistically similar for ferns and angiosperms.
+Across all habitats, the N:P ratio was statistically similar for ferns and angiosperms.
 
@@
 Across closed canopy habitats, Lycophytes have 45% lower foliar nitrogen content than ferns and angiosperms, which do not statistically differ (*P* < 0.001).
-Conseqently, Lycophytes had significantly higher CN ratios than ferns and angiosperms (*P* < 0.001).
 Lycophytes also had 24% lower foliar phosphorous content than angiosperms (*P* = .026), while ferns were simiilar to both groups.
-The NP ratios were not statistically different across angiosperms, ferns and lycophytes in closed canopies habitats.
+The N:P ratios were not statistically different across angiosperms, ferns and lycophytes in closed canopies habitats.
 
@@
 
-**Table 1**. Species list.
+**Table 1**. Surveyed species of angiosperms, ferns and lycophytes across both open canopy high-light and closed canopy low-ligh habitats.
 
@@
 \
-**Table 2**. Pooled ecophysiological traits for different plant lineages in open light and closed habitats.
+**Table 2**. Pooled ecophysiological traits for different plant lineages in open light and closed habitats. Each value reflects the mean (&pm; 1 standard error) for each treatment.
 
@@
 traitstab <- read.csv("manuscript/pooled_data.csv")
-physymb <- '\u03D5'
+# source("manuscript/paper_labels.R")
+# physymb <- '\u03D5'
+
+phototab <- expression(atop(italic(A[n]),(mu*mol~m^-2~s^-1)))
 
 dust(traitstab)%>%
-  sprinkle_colnames("Plant Lineage", "Canopy", "*A*~n~","*g*~s~","Foliar N",
-                    "Foliar P", "C:N", "N:P", "SD", physymb, "*R*~d~", "LCP") %>%
+  sprinkle_colnames("Plant Lineage", "Canopy", 
+                    "*A*~n~ ($\\mu$mol m^-2^ s^-1^)",
+                    "*g*~s~ (mol m^-2^ s^-1^)",
+                    "Foliar Nitrogen(%)", "Foliar Phosphorus (%)","N:P",
+                    "WUE ($\\mu$mol CO~2~ mol H~2~O^-1^)", 
+                    "SD (mm^-2^)", 
+                    "$\\Phi$ (mol CO~2~ mol photons^-1^)", 
+                    "*R*~d~ ($\\mu$mol m^-2^ s^-1^)",
+                    "LCP ($\\mu$mol~m^-2~s^-1)") %>%
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
@@
 **Figure 1**.
-Redundancy analysis of measured functional traits across con-occuring angiosperms and fern species in open and shaded habitats.
-With the 12 measured variables, number% of the trait variation was accounted for in two dimensions.
-Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content, the leaf carbon to nitrogen ratio (C:N) and the leaf nitrogen to phosphorus ratio (N:P).
+Redundancy analysis of measured functional traits across con-occuring similar statured angiosperm, fern and lycophyte species in shaded habitats.
+With the 10 measured variables, 52.9 % of the trait variation was accounted for in two dimensions.
+Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content and the leaf nitrogen to phosphorus ratio (N:P).
 
@@
 ```  
-**Figure 2**. Redundancy analysis of measured functional traits across con-occuring angiosperms fern and lycopyte species in shaded habitats.
+**Figure 2**.
+Redundancy analysis of measured functional traits across con-occuring similar statured angiosperms and fern species in open and shaded habitats.
+With the 10 measured variables, 55.9 % of the trait variation was accounted for in two dimensions.
+Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content and the leaf nitrogen to phosphorus ratio (N:P).
 

### 2019-03-26 manuscript/paper_labels.R

@@ -1,0 +1,22 @@
+#plotobjects
+
+condtab <- expression(atop(italic(g)[s],(mol~m^-2~s^-1)))
+
+phototab <- expression(atop(italic(A[n]),(mu*mol~m^-2~s^-1)))
+
+resptab <- expression(atop(italic(R[d]),(mu*mol~m^-2~s^-1)))
+
+nareatab <- "Foliar~Nitrogen\n (%)"
+pareatab <- "Foliar~Phosphorous\n (%)"
+
+lcptab <- expression(atop(LCP,(mu*mol~m^-2~s^-1)))
+denstab <- expression(atop(Stomatal~Density,(mm^-2)))
+
+phitab<- expression(paste(phi, "  (mol ", " CO" ["2"], "  m",ol^-1, " photons)"))
+
+wuetab <- expression(atop(WUE, paste("(",mu*mol," CO" ["2"]," mmol ",
+                                     H[2],O^-1,")")))
+
+
+variables <- c("Plant Lineage", "Canopy", phototab,condtab,nareatab,
+                    pareatab, "N:P", wuetab, denstab, phitab, resptab, lcptab)

### 2019-03-26 master_scripts/data_table.R

@@ -7,16 +7,14 @@ treatments <- read.csv("raw_data/species_list.csv")
 
 ##photosynthesis and leaf chemistry data-------------
 photochem <- read.csv("raw_data/photo_chem.csv")
-  photochem$cn_ratio <- with(photochem, c_perc/n_perc)
-  photochem$percN <- photochem$n_perc /100
   photochem$np_ratio <- with(photochem, n_perc/p_perc)
-  photochem$percP <- photochem$p_perc /100
+  photochem$wue <- with(photochem, Photo/Trmmol)
 
-photochem_mean <- summaryBy(Photo+Cond+percN+percP+cn_ratio+np_ratio ~ 
+photochem_mean <- summaryBy(Photo+Cond+n_perc+p_perc+np_ratio+wue ~ 
                               plant_group+canopy, 
                               data=photochem, FUN=mean2)
 
-photochem_se <- summaryBy(Photo+Cond+percN+percP+cn_ratio+np_ratio ~ 
+photochem_se <- summaryBy(Photo+Cond+n_perc+p_perc+np_ratio+wue ~ 
                               plant_group+canopy, 
                             data=photochem, FUN=se)
 
@@ -50,17 +48,17 @@ photo <- data.frame(paste0(sprintf("%2.1f", round(photochem_mean[,3], 1)),
 cond <- data.frame(paste0(sprintf("%1.2f", round(photochem_mean[,4], 2)), 
                            " (", sprintf("%1.3f", round(photochem_se[,4],3)),")"))
 
-nitro <- data.frame(paste0(sprintf("%1.2f", round(photochem_mean[,5], 2)), 
-                          " (", sprintf("%1.3f", round(photochem_se[,5],3)),")"))
+nitro <- data.frame(paste0(sprintf("%1.1f", round(photochem_mean[,5], 1)), 
+                          " (", sprintf("%1.2f", round(photochem_se[,5],2)),")"))
 
-phos <- data.frame(paste0(sprintf("%1.3f", round(photochem_mean[,6], 3)), 
-                          " (", sprintf("%1.4f", round(photochem_se[,6],4)),")"))
+phos <- data.frame(paste0(sprintf("%1.2f", round(photochem_mean[,6], 2)), 
+                          " (", sprintf("%1.3f", round(photochem_se[,6],3)),")"))
 
[…]

### 2019-03-26 scripts/vpd.R

@@ -1,0 +1,13 @@
+## quick means of vpd
+source("functions/basic_functions.R")
+
+vpd <- read.csv("raw_data/vpdl.csv")
+
+library(doBy)
+
+mean(vpd$VpdL)
+se(vpd$VpdL)
+sd(vpd$VpdL)
+max(vpd$VpdL)
+min(vpd$VpdL)
+hist(vpd$VpdL)

### 2019-03-26 manuscript/manuscript.Rmd
figures in text, need to complete seed vs spore figures and stats

@@
 
+```{r figtabs, echo=FALSE}
+
+pcagroup <- "1"
+pcashade <- "2"
+interfig <- "3"
+shadefig <- "4"
+seedsporefig <- "5"
+anitro <- "6"
+
+```
 # Abstract
@@
 Louis, USA).
-Carbon and nitrogen analyses were measured using a Costech Analytical Elemental Analyzer (Valencia, USA), with the percentage of carbon and nitrogen in samples calculated by comparison with certified standards.
-Foliar phosphorus concentrations were determined using an ash digestion process [@dangelo_rapid_2001] preceded by color development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon, USA).
+Foliar nitrogen content (N) was measured using a Costech Analytical Elemental Analyzer (Valencia, USA), with the percentage N in samples calculated by comparison with certified standards.
+Foliar phosphorus concentrations (P) were determined using an ash digestion process [@dangelo_rapid_2001] preceded by color development and absorbance measurement on an Astoria Paciﬁc colorimetric autoanalyzer (Clackamas, Oregon, USA).
 
@@
 In total, 52.9 % of the trait variation was accounted for in two dimensions, encompassing different resource economics, physiology and anatomy of the surveryed taxa.
-Interestingly, there was minimal overlap among component scores for angiosperms species compared to spore-bearing ferns and lycophytes (Figure 1).
+Interestingly, there was minimal overlap among component scores for angiosperms species compared to spore-bearing ferns and lycophytes (Figure `r pcagroup`).
 Specifically, angiosperms species were associated with gas exchange, anatomical and stoichiometric functional traits related to optimizing carboxylation.
@@
 In total, 55.9 % of the trait variation was accounted for in two dimensions.
-Component scores for angiosperms and fern species broadly diverged across open and closed canopy habitats, however, associations with physiological traits in open habitats were stronger for angiosperms (Figure 2).
+Component scores for angiosperms and fern species broadly diverged across open and closed canopy habitats, however, associations with physiological traits in open habitats were stronger for angiosperms (Figure `r pcashade`).
 Specifically, angiosperms appear more plastic in their ability to upregulate traits related to photosynthetic gas exchange with increased light availability.
@@
 This is evident in broad differences in mean values of ecophyioslogical traits across the three plant lineages and among habitat types (Table 2).
+
+## Leaf gas exchange
+
+Overall, fern and angiosperm species in open habitats had higher A~n~ than comparable species in closed habitats (*P* < 0.0001).
+Specifically, fern species in high-light habitats had 40.6 % greater A~n~ than in closed habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 60 % higher A~n~ (*P* < 0.0001, t=-6.97).
+Ferns and angiosperms growing in low-light habitats had similar A~n~, however, in open habitats angiosperms had 42.6% higher rates of A~n~ than ferns (canopy x plant group, *P* = 0.043, Figure `r interfig`A).
+Overall, ferns and angiosperms in open habitats also had 64.2% higher g~s~ than comparable species in closed habitats (*P* = 0.002).
+Specifically, fern species in high-light habitats had 50.7 % greater g~s~ than in closed habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 66.9 % higher g~s~ (*P* < 0.0001, t=-4.157).
+Across both habitat types, angiosperm species had a 60% higher g~s~ than fern species (*P* = 0.011, Figure `r interfig`B).
+Neither ferns nor angiosperms species differed in WUE across the two light environments.
+Across all habitats, WUE was also statistically similar for ferns and angiosperms.
+
+\
+
+In closed habitats, variation in gas exchange parameters was largely associated with individual species variation (random effect) instead of functional differences between all three plant groups.
+No differences between A~n~ were detected between between the three plant groups.
+Rates of g~s~ were highest in  angiosperms (*P* = 0.001, Figure `r shadefig`A).
+Overall, the spore-bearing plant lineages had lower g~s~ (), regardless of habitat (Figure `r seedsporefig`A).
+Water-use efficiency was 47 % higher in lycophytes than ferns and angiosperms (*P* = 0.005), while ferns and angiosperms where statiscally similar (Figure `r shadefig`B).
 
@@
 Quantum yields were also equivalent within angiosperm and fern species groups between open and closed habitats.
-Shifts in R~d~ between open and closed habitats differed for angiosperms and ferns (habitat x plant group, *P* < 0.001).
+Shifts in R~d~ between open and closed habitats differed for angiosperms and ferns (habitat x plant group, *P* < 0.001, Figure `r interfig`C).
 Specifically, R~d~ in angiosperms was 80.6 % greater in open compared to closed habitats (t=-4.197, *P* = 0.001).
 Alternatively, R~d~ in ferns was 44.0 % greater in low-light compared to open habitats (t=-2.97, *P* = 0.006).
-Similarly, shifts in LCP between open and closed habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001).
+Similarly, shifts in LCP between open and closed habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001, Figure `r interfig`D).
 Specifically, the LCP in angiosperms was 79.2 % greater in open compared to closed habitats (t=-3.5998, *P* = 0.005).
@@
 
-In closed canopy environments,the LCP in ferns were equivalent to lycophytes, while both plant groups were marginally higher than angiosperms (*P* = 0.061).
+In closed canopy environments, the LCP of spore-bearing plant species were equivalent and marginally higher than angiosperm species (*P* = 0.061, Figure `r seedsporefig`B).
 Photosynthetic quantum was similar between each plant group, but lycophytes and ferns were broadly higher than angiosperms.
@@
 
-## Leaf gas exchange
+## Stoichiometry (##can use pairwise for specific p values: see leaf P below)
 
-Overall, fern and angiosperm species in open habitats had higher A~n~ than comparable species in low-light habitats (*P* <- 0.0001).
-Specifically, fern species in high-light habitats had 40.6 % greater A~n~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 60 % higher A~n~ (*P* < 0.0001, t=-6.97).
-Ferns and angiosperms growing in low-light habitats had similar A~n~, however, in open habitats angiosperms had 42.6% higher rates of A~n~ than ferns (canopy x plant group, *P* = 0.043).
-Neither ferns nor angiosperms species differed in WUE across the two light environments.
-Across all habitats, the WUE was also statistically similar for ferns and angiosperms.
-
+Ferns and angiosperms species in open habitats had marginally lower foliar N content compared to similar species in growing in closed light environments (-17.2 %, *P* = 0.072), with a large proportion of variation attributed to individual species differences.
+No differences in foliar P content were detected between open and closed habitats for either ferns or angiosperms.
+Across all habitats, foliar P content was 22.1 % higher in angiosperms than in ferns (*P* =  0.034).
+Neither ferns nor angiosperms species differed in their N:P ratios across the two habitat types.
 \
 
-Overall, ferns and angiosperms in open habitats had 64.2% higher g~s~ than comparable species in low-light habitats (*P* = 0.002).
-Specifically, fern species in high-light habitats had 50.7 % greater g~s~ than in low-light habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 66.9 % higher g~s~ (*P* < 0.0001, t=-4.157).
-Across both habitat types, angiosperm species had a 60% higher g~s~ than fern species (*P* = 0.011).
-
-\
-
-In closed canopy environments, variation in gas exchange parameters was largely due to individual species variation instead of functional differences between all three plant groups.
-No differences between A~n~ were detected between between the three plant groups, with a majority of the variation associated with the random effect of individual species.
-Stomatal conductance in angiosperms was higher than in ferns and lycophytes **P* = 0.001).
-Water-use efficiency was 47 % higher in lycophytes than ferns and angiosperms (*P* = 0.005), while ferns and angiosperms where statiscally similar.
-
-## Stoichiometry (##can use pairwise for specific p values: see leaf P below)
-
-Ferns and angiosperms species in open light environments had marginally lower foliar nitrogen content compared to similar species in growing in closed light environments (-17.2 %, *P* = 0.072), with a large proportion of variation attributed to individual species differences.
-No differences in foliar P content were detected between open and closed canopies for either ferns or angiosperms.
-Across all habitats, foliar P content was 22.1 % higher in angiosperms than in ferns (*P* =  0.034).
-Neither ferns nor angiosperms species differed in their NP ratios across the two light environments.
-Across all habitats, the N:P ratio was statistically similar for ferns and angiosperms.
-
-\
-
-Across closed canopy habitats, Lycophytes have 45% lower foliar nitrogen content than ferns and angiosperms, which do not statistically differ (*P* < 0.001).
-Lycophytes also had 24% lower foliar phosphorous content than angiosperms (*P* = .026), while ferns were simiilar to both groups.
-The N:P ratios were not statistically different across angiosperms, ferns and lycophytes in closed canopies habitats.
+Within closed habitats, lycophytes have 45% lower foliar N content than ferns and angiosperms, which do not statistically differ (*P* < 0.001, Figure `r shadefig`C).
+Lycophytes also had 24% lower foliar P content than angiosperms (*P* = .026), while ferns were simiilar to both groups (Figure `r shadefig`D).
+The N:P ratios were not statistically different across angiosperms, ferns and lycophytes in closed habitats.
 
@@
 
-Overall, no differences where detected in stomatal densities between species growing open canopy and closed canopy habitats.
-Specifically, differences in stomatal densities were not detected for either fern nor angiosperms species between the two habitats.
-However, in both open and closed environments, ferns had significantly lower stomatal densities than analogous angiosperms (*P* < 0.001).
-
-\
-
-In closed canopy environments, no differenes in stomatal density were detected among species of ferns, angiosperms and lycophytes.
-Variation in stomatal denisty in the shade was largely due to individual species variation.
+Overall, no differences where detected in stomatal density (SD) between species growing in open and closed habitats.
+Specifically, differences in SD were not detected for either fern nor angiosperms species between the two habitats.
+However, in both open and closed environments, ferns had significantly lower SD than analogous angiosperms (*P* < 0.001).
+Additionally, no differenes in SD were detected among species of ferns, angiosperms and lycophytes in closed habitats.
+Variation in SD in the shade was largely due to individual species variation.
 
@@
 
-The relationship between A~n~ and foliar nitrogen content depended on both plant group and canopy type (canopy x plant group, *P* = 0.028).
-Specifically, the postive slopes of the relationship between A~n~ and N differed for ferns and angiosperms in open canopys, while no relationship beween A~n~ and N was found for either group in closed canopies (Figure 6).
+The relationship between A~n~ and foliar N content depended on both plant group and habitat type (habitat x plant group, *P* = 0.028).
+Specifically, the postive slopes of the relationship between A~n~ and N differed for ferns and angiosperms in open habitats, while no relationship beween A~n~ and N was found for either group in closed habitat (Figure `r anitro`).
 
@@
 dust(traitstab)%>%
-  sprinkle_colnames("Plant Lineage", "Canopy", 
+  sprinkle_colnames("Lineage", "Canopy", 
                     "*A*~n~ ($\\mu$mol m^-2^ s^-1^)",
@@
                     "Foliar Nitrogen(%)", "Foliar Phosphorus (%)","N:P",
-                    "WUE ($\\mu$mol CO~2~ mol H~2~O^-1^)", 
+                    "WUE ($\\mu$mol CO~2~ mmol H~2~O^-1^)", 
                     "SD (mm^-2^)", 
@@
   sprinkle(rows= 1, part="head",bold=TRUE)%>%
+  sprinkle(rows = 3, cols = 4,bold=TRUE, part="body")%>%
   sprinkle_print_method("markdown")
@@
 **Figure 1**.
-Redundancy analysis of measured functional traits across con-occuring similar statured angiosperm, fern and lycophyte species in shaded habitats.
+Redundancy analysis of measured functional traits across con-occuring similar statured angiosperm, fern and lycophyte species in closed habitats.
 With the 10 measured variables, 52.9 % of the trait variation was accounted for in two dimensions.
-Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content and the leaf nitrogen to phosphorus ratio (N:P).
+Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content and the foliar nitrogen to phosphorus ratio (N:P).
 
@@
 **Figure 2**.
-Redundancy analysis of measured functional traits across con-occuring similar statured angiosperms and fern species in open and shaded habitats.
+Redundancy analysis of measured functional traits across con-occuring similar statured angiosperms and fern species in open and closed habitats.
 With the 10 measured variables, 55.9 % of the trait variation was accounted for in two dimensions.
-Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content and the leaf nitrogen to phosphorus ratio (N:P).
+Variables include net photosynthesis (A~n~), stomatal conductance (g~s~), instantaneous water use efficiency (WUE), stomatal density (SD), the light compensation point (LCP), quantum yield ($\phi$), dark respiration (R~d~), foliar nitrogen (N) and phosphorus (P) content and the foliar nitrogen to phosphorus ratio (N:P).
 
@@
 ```  
-**Figure 3**. Differences in mean physiological traits between co-occuring angiosperms and fern species in open light and closed habitats.
+**Figure 3**. Differences in mean physiological traits between co-occuring similar statured angiosperms and fern species across open and closed habitats.
 
 ```{r lyco, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
-source("master_scripts/lycophytes_plot.R")
+source("master_scripts/shade_lineages.R")
 ```  
-**Figure 4**. Differences in mean physiological traits between lycophytes species that are constrained to only shade habitats and ferns and angiosperms.
+**Figure 4**. Differences in mean physiological traits between con-occuring similar statured angiosperm, fern and lycophyte species in closed habitats.
 
@@
 ```  
-**Figure 6**. Postive relationship between photosynthesis and foliar nitrogen content exists for angiosperms and fern species in open light habitat, but not in shade habitats
+**Figure 6**.
+Postive relationships between light saturated photosynthesis (A~n~) and foliar nitrogen content exists for small statured angiosperm and fern species in open habitat, but not in closed habitats.
+For significant linear relationships, dashed lines represent model fits between traits and grey shaded areas are 95 % confidence intervals for the mean.
+Conditional and marginal R^2^ are reported for the overall model.
 

### 2019-03-26 master_scripts/bivariate_plots.R

@@ -30,7 +30,9 @@ anitro_mod2 <- lmer(Photo~n_perc + canopy * plant_group + (1|species),
 
 # emmip(anitro_mod2, canopy ~ plant_group)
 # emmeans(anitro_mod2, pairwise ~ canopy : plant_group)
-# r.squaredGLMM(anitro_mod2)
+getr2 <- r.squaredGLMM(anitro_mod2)
+margr2 <- round(getr2[[1]],2)
+condr2 <- round(getr2[[2]],2)
 
 ##simple models for ablineclip, paired with full model
 fern_photoN_open <- lm(Photo~n_perc, data=fern[fern$canopy == "Open",])
@@ -71,6 +73,11 @@ angioopen <- droplevels(chem[chem$plant_group == "Angio" & chem$canopy == "Open"
 par(mar=c(5,5,1,1))
 plot(Photo~n_perc, data=chem, ylab=photolab, xlab = "Foliar Nitrogen  (%)",
      type='n', ylim=c(0,23), xlim=c(0,5.5))
+
+#here is the predlines example
+predline(angio_photoN_open, col=plantcols[2],lwd=2, lty=2)
+predline(fern_photoN_open, col=plantcols[3],lwd=2, lty=2)
+
 points(Photo~n_perc, data=angio, pch=c(21,22)[canopy], bg=plantcols[2], cex=1.25)
 points(Photo~n_perc, data=fern, pch=c(21,22)[canopy], bg=plantcols[3], cex=1.25)
 #fixed effects 
@@ -83,13 +90,12 @@ points(Photo~n_perc, data=fern, pch=c(21,22)[canopy], bg=plantcols[3], cex=1.25)
 # ablineclip(angioopen_fix[[1]], angioopen_fix[[2]],lwd=2,col=cols2[1],
 #            x1=.95, x2=3.66)
 
-#here is the predlines example
-predline(angio_photoN_open, col=plantcols[2],lwd=2, lty=2)
-predline(fern_photoN_open, col=plantcols[3],lwd=2, lty=2)
-
 legend("topleft", legend = c("Angiosperms", "Ferns", "Closed", "Open"),
-       pch = c(21, 21, 16, 15), bty='n', inset=.01,
-       pt.bg = c(plantcols[2],plantcols[3]))
+       pch = c(21, 21, 21, 22), bty='n', inset=.01,
[…]

### 2019-03-26 master_scripts/interaction_plots.R

@@ -72,10 +72,8 @@ lines(x=c(1,2), y=c(photo_agg[3,3],photo_agg[4,3]), lwd=2.5, col="forestgreen")
 lines(x=c(1,2), y=c(photo_agg[1,3],photo_agg[2,3]), lwd=2.5, col="firebrick")
 axis(1, at=1:2, labels = FALSE)
 axis(2, las=1)
-legend("topleft", c("Angiosperms", "Ferns"), pch=19, 
-       col=c(plantcols[2], plantcols[3]),
-                    bty='n', inset=.01)
-# text(x=1, y=14, label = "canopy x plant group,\n P = 0.043")
+
+text(x=0.5, y=15, label = "A", cex=1.25)
 
 #cond (add model stats)
 par(mar=c(0,0,1,5))
@@ -94,6 +92,11 @@ lines(x=c(1,2), y=c(cond_agg[1,3],cond_agg[2,3]), lwd=2.5, col="firebrick")
 axis(1, at=1:2, labels = FALSE)
 axis(4, las=1)
 mtexti(condlab, 4, outer=TRUE, cex=1, off=.6)
+text(x=0.5, y=.4, label = "B", cex=1.25)
+
+legend("topright", c("Angiosperms", "Ferns"), pch=19, 
+        col=c(plantcols[2], plantcols[3]),
+        bty='n', inset=.01)
 
 #rd (add model stats)
 par(mar=c(5,5,0,0))
@@ -111,6 +114,7 @@ lines(x=c(1,2), y=c(rdagg[3,3],rdagg[4,3]), lwd=2.5, col="forestgreen")
 lines(x=c(1,2), y=c(rdagg[1,3],rdagg[2,3]), lwd=2.5, col="firebrick")
 axis(1, at=1:2, labels = c("Closed", "Open"))
 axis(2, las=1)
+text(x=0.5, y=1.09, label = "C", cex=1.25)
 
 #LCP (add model stats)
 par(mar=c(5,0,0,5))
@@ -129,6 +133,7 @@ lines(x=c(1,2), y=c(lcp_agg[1,3],lcp_agg[2,3]), lwd=2.5, col="firebrick")
 axis(1, at=1:2, labels = c("Closed", "Open"))
 axis(4, las=1)
 mtexti(lcplab, 4, outer=TRUE, cex=1, off=.6)
+text(x=0.5, y=20.7, label = "D", cex=1.25)
 
[…]

### 2019-03-26 master_scripts/lycophytes_plot.R

@@ -1,45 +1,0 @@
-source("master_scripts/plot_objects.R")
-source("functions/basic_functions.R")
-
-##lycophytes differ
-
-chem <- read.csv("raw_data/photo_chem.csv")
-library(vioplot)
-
-chem_closed <- chem[chem$canopy == "Closed",]
-  chem_closed$cnratio <- with(chem_closed, c_perc/n_perc)
-  chem_closed$wue <- with(chem_closed, Photo/Trmmol)
-
-phos <- chem_closed[chem_closed$p_perc > .1 & chem_closed$p_perc < .4,]
-
-# windows()
-par(mfrow=c(2,2), mgp=c(3,.75,0))
-
-#N content
-par(mar=c(0,5.5,1,0))
-boxplot(n_perc ~ plant_group, data=chem_closed, ylab="Foliar N content  (%)",
-        col=pgcols, ylim=c(0,5), xaxt='n', yaxt='n', outline=FALSE)
-axis(2, las=1)
-
-#P content
-par(mar=c(0,0,1,5.5))
-boxplot(p_perc ~ plant_group, data=phos, ylab="",
-        col=pgcols, ylim=c(0,.425), xaxt='n', yaxt='n')
-axis(4,las=1)
-mtexti("Foliar P content  (%)", 4, outer=TRUE, cex=1, off=.6)
-
-#cn ratio
-par(mar=c(5.5,5.5,0,0))
-boxplot(cnratio ~ plant_group, data=chem_closed, ylab="Foliar C:N ratio",yaxt='n',
-        names=plants,col=pgcols, ylim=c(0,45), outline=FALSE)
-axis(2, las=1)
-
-#water use efficiency
-par(mar=c(5.5,0,0,5.5))
-boxplot(wue ~ plant_group, data=chem_closed, ylab="",names=plants,
-        col=pgcols, ylim=c(0,21), outline=FALSE, yaxt='n')
-axis(4,las=1)
-mtexti(wuelab, 4, outer=TRUE, cex=1, off=.6)
-
-# dev.copy2pdf(file= "output/lycophytes.pdf")
-# dev.off()

### 2019-03-26 master_scripts/lycophytes_plot2.R

@@ -1,44 +1,0 @@
-source("master_scripts/plot_objects.R")
-source("functions/basic_functions.R")
-
-##lycophytes differ
-
-chem <- read.csv("raw_data/photo_chem.csv")
-  chem$cnratio <- with(chem, c_perc/n_perc)
-  chem$wue <- with(chem, Photo/Trmmol)
-
-phos <- chem[chem$p_perc > .1 & chem$p_perc < .4,]
-
-
-##panel plot with non-lycos in both habitats
-windows()
-par(mfrow=c(2,2), mgp=c(3,.75,0))
-
-#N content
-par(mar=c(0,5.5,1,0))
-boxplot(n_perc ~ plant_group, data=chem, ylab="Foliar N content  (%)",
-        col=pgcols, ylim=c(0,5), xaxt='n', yaxt='n', outline=FALSE)
-axis(2, las=1)
-
-#P content
-par(mar=c(0,0,1,5.5))
-boxplot(p_perc ~ plant_group, data=phos, ylab="",outline=FALSE,
-        col=pgcols, ylim=c(0,.425), xaxt='n', yaxt='n')
-axis(4,las=1)
-mtexti("Foliar P content  (%)", 4, outer=TRUE, cex=1, off=.6)
-
-#cn ratio
-par(mar=c(5.5,5.5,0,0))
-boxplot(cnratio ~ plant_group, data=chem, ylab="Foliar C:N ratio",yaxt='n',
-        names=plants,col=pgcols, ylim=c(0,45), outline=FALSE)
-axis(2, las=1)
-
-#water use efficiency
-par(mar=c(5.5,0,0,5.5))
-boxplot(wue ~ plant_group, data=chem, ylab="",names=plants,
-        col=pgcols, ylim=c(0,21), outline=FALSE, yaxt='n')
-axis(4,las=1)
-mtexti(wuelab, 4, outer=TRUE, cex=1, off=.6)
-
-dev.copy2pdf(file= "output/lycophytes2.pdf")
-dev.off()

### 2019-03-26 master_scripts/seed_nonseed_plots.R

@@ -25,7 +25,7 @@ lrc_shade <-  droplevels(lrc[!lrc$canopy == "Closed",])
 
 
 #seed vs nonseed panel figure
-# windows()
+windows()
 par(mfrow=c(2,2), mgp=c(3,.75,0))
 
 #stomatal conductance
@@ -33,6 +33,7 @@ par(mar=c(0,5,1,0))
 boxplot(Cond ~ seed, data=chem_closed, ylab=condlab,xaxt='n',yaxt='n',
         ylim=c(0,.3), outline=FALSE, col=seedcols)
 axis(2, las=1)
+text(x=0.5, y=.3, label = "A", cex=1.25)
 
 #quantum yield
 par(mar=c(0,0,1,5))
@@ -40,12 +41,14 @@ boxplot(Phi ~ seed, data=lrc_shade, ylab="", xaxt='n', yaxt='n',
         ylim=c(0.04,.09),  col=seedcols)
 axis(4,las=1)
 mtexti(philab2, 4, outer=TRUE, cex=1, off=.65)
+text(x=0.5, y=.09, label = "B", cex=1.25)
 
 #light compensation point
 par(mar=c(5,5,0,0))
 boxplot(LCP ~ seed, data=lrc_shade, ylab=lcplab,outline=FALSE,yaxt='n',
         ylim=c(0.,27),  col=c("white", plantcols2[2]))
 axis(2, las=1)
+text(x=0.5, y=26.95, label = "C", cex=1.25)
 
 #dark respiration
 par(mar=c(5,0,0,5))
@@ -53,6 +56,7 @@ boxplot(Rd ~ seed, data=lrc_shade, ylab="", outline=FALSE,
         ylim=c(0,1.75),  col=c("white", plantcols2[2]),yaxt='n')
 axis(4,las=1)
 mtexti(resplab2, 4, outer=TRUE, cex=1, off=.65)
+text(x=0.5, y=1.74, label = "D", cex=1.25)
 
-# dev.copy2pdf(file= "output/seed_nonseed_plot.pdf")
-# dev.off()
+dev.copy2pdf(file= "output/seed_nonseed_plot.pdf")
+dev.off()

### 2019-03-26 master_scripts/shade_lineages.R

@@ -1,0 +1,47 @@
+source("master_scripts/plot_objects.R")
+source("functions/basic_functions.R")
+
+##lycophytes differ
+
+chem <- read.csv("raw_data/photo_chem.csv")
+library(vioplot)
+
+chem_closed <- chem[chem$canopy == "Closed",]
+chem_closed$cnratio <- with(chem_closed, c_perc/n_perc)
+chem_closed$wue <- with(chem_closed, Photo/Trmmol)
+
+phos <- chem_closed[chem_closed$p_perc > .1 & chem_closed$p_perc < .4,]
+
+#boxplots of significant traits
+
+# windows()
+par(mfrow=c(2,2), mgp=c(3,.75,0))
+
+par(mar=c(0,5.5,1,0))
+boxplot(Cond~ plant_group, data=chem_closed, ylab=condlab, xaxt='n', yaxt='n',
+        col=pgcols, ylim=c(0,0.26), outline=FALSE)
+axis(2, las=1)
+text(x=0.5, y=0.26, label = "A", cex=1.25)
+
+par(mar=c(0,0,1,5.5))
+boxplot(wue ~ plant_group, data=chem_closed, ylab="",xaxt='n', 
+        col=pgcols, ylim=c(0,16), outline=FALSE, yaxt='n')
+axis(4,las=1)
+mtexti(wuelab, 4, outer=TRUE, cex=1, off=.6)
+text(x=0.5, y=16, label = "B", cex=1.25)
+
+par(mar=c(5.5,5.5,0,0))
+boxplot(n_perc ~ plant_group, data=chem_closed, ylab="Foliar N content  (%)",
+        col=pgcols, ylim=c(0,5), yaxt='n', names=plants,outline=FALSE)
+axis(2, las=1)
+text(x=0.5, y=4.9, label = "C", cex=1.25)
+
+par(mar=c(5.5,0,0,5.5))
+boxplot(p_perc ~ plant_group, data=phos, ylab="",
+        col=pgcols, ylim=c(0,.425), yaxt='n',names=plants)
+axis(4,las=1)
+mtexti("Foliar P content  (%)", 4, outer=TRUE, cex=1, off=.6)
+text(x=0.5, y=.42, label = "D", cex=1.25)
+
+# dev.copy2pdf(file= "output/shade_plants.pdf")
+# dev.off()

### 2019-03-26 stats/gs_stats.R

@@ -7,7 +7,7 @@ library(multcomp)
 library(pbkrtest)
 
 
-#photosynthesis
+#gs
 gs <- read.csv("raw_data/photo_chem.csv")
 gs$canopyplant <- interaction(gs$canopy, gs$plant_group)
 
@@ -36,9 +36,8 @@ visreg(shade_mod)
 
 tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
 shade_siglets <-cld(tukey_shade)
-shade_siglets2 <- shade_siglets$mcletters$Letters
 #Angio  Fern  Lyco 
-#"b"  "ab"   "a"
+#"b"  "a"   "a"
 
 
 #canopy comparisions of sun shade with angio and ferns

### 2019-03-26 stats/seed_spore_stats.R

@@ -1,0 +1,132 @@
+### seed vs non-seed stats
+#gs, phi, lcp, rd
+
+library(visreg)
+library(car)
+library(lme4)
+library(MuMIn)
+library(multcomp)
+
+library(dplyr)
+library(stringr)
+#treatments
+treatments <- read.csv("raw_data/species_list.csv")
+
+
+#aq_curves
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+  mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+         individual =  str_remove(ID, ".*-")) %>%
+  merge(treatments[,c(2,4:5)]) %>%
+  mutate(seed = ifelse(plant_group == "Angiosperm", "seed", "spore"))
+
+
+seed <- droplevels(lrc[lrc$seed == "seed" & lrc$canopy == "Closed",])
+spore <- droplevels(lrc[lrc$seed == "spore" & lrc$canopy == "Closed",])
+
+#PHI
+t.test(seed$Phi, spore$Phi, var.equal=FALSE)
+#not significant
+
+#LCP
+t.test(seed$LCP, spore$LCP, var.equal=FALSE)
+
+#Rd
+t.test(seed$Rd, spore$Rd, var.equal=FALSE)
+
+
+
+#LCP models, with species effects
+lrc_shade <- droplevels(lrc[lrc$canopy == "Closed",])
+
+#simple
+lcp_mod <- lm(LCP ~ seed, data=lrc_shade)
+plot(lcp_mod)
+Anova(lcp_mod, type=3)
+visreg(lcp_mod)
+
+#full model
+lcp_mod2 <-lmer(sqrt(LCP) ~ seed + (1|species), data=lrc_shade)
+plot(lcp_mod2)
+#used sqrt to deal with residuals versus fitted values
+
+Anova(lcp_mod2, type=3)
+summary(lcp_mod2)
+r.squaredGLMM(lcp_mod2) 
+
+#          R2m       R2c
+# [1,] 0.2533775 0.4984394
+# P = 0.01073 *  
+
+tukey_lcp <- glht(lcp_mod2, linfct = mcp(seed = "Tukey"))
+lcp_siglets <-cld(tukey_lcp)
+# seed spore 
+# "a"   "b"
+
+
+
+#Rd models-------
+
+#simple
[…]

### 2019-03-27 manuscript/manuscript.Rmd
hyuck methods and results draft

@@
 ---
-title: "Adaptability of ecophysiological traits define how plant lineages make a living in shaded environments "
+title: "Plasticity of ecophysiological traits structure how herbaceous plant lineages make a living in shaded environments "
 author: 
@@
 
-/
+\
 
@@
 A primary controlling factor in fern hydraulics is stomatal function.
-A series of elegeant studies have shown that stomatal control of ferns and lycophytes differs fundamentally from seed plants.
-Evidence on fern and lycophyte stomata indicates that these groups have inefficient stomatal systems, likely possessing a hydropassive mechanism which influences stomatal opening and closing (many cites).
+A series of elegant studies have shown that stomatal control of ferns and lycophytes differs fundamentally from seed plants.
+Evidence on fern and lycophyte stomata indicates that these groups have inefficient stomatal systems, likely possessing a hydro-passive mechanism which influences stomatal opening and closing (many cites).
 Specifically, unlike seed plants, fern and lycophyte stomata fail to close in response to abscisic acid (ABA) addition (many cites).
-Additionally, recent evidence also suggests that co-occuring tropical ferns and lycophytes optimize thier ecophysiology differently (Campany).
-Consequently, the evolution or adaptation of physiolgoical traits may impact the differential distribution of ferns,lycophytes and seed plants to eath other.
+Additionally, recent evidence also suggests that co-occurring tropical ferns and lycophytes optimize their ecophysiology differently (Campany).
+Consequently, the evolution or adaptation of physiological traits may impact the differential distribution of ferns,lycophytes and seed plants to each other.
 
@@
 
-The goal of the current study is to examine the comparative physiology of a number of temperate ferns and lycophytes to angiosperms.
-We specifically examined several physiological parameters in comparison between co-occurring ferns, lycophytes and  angiosperm in identical closed canopy habitats.
-In a pairwise camoparison, we also evaluated the potential of fern and angiosperm species to adapt to open-light habitats.
+The goal of this study is to examine the comparative ecophysiology of a number of temperate ferns and lycophytes to angiosperms.
+First, we compared a suite of plant functional traits between co-occurring herbaceous ferns, lycophytes and  angiosperm in identical closed canopy habitats.
+In a pairwise comparison, we then evaluated the potential of fern and angiosperm species to adapt to open-light habitats.
 Given increasing evidence of inefficient stomatal functions in non-seed plants, we hypothesized that ferns and lycophytes would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms.
-This, in turn, would impact aspects of leaf anatomy (stomatal density), nutrient relations (N and P), and water use efficiency.
-We speculate that the differences in physiological function also affect the ecology of lycophytes in a manner that prevents them from establishing in full sun environments.
+This, in turn, would impact aspects of leaf anatomy, nutrient relations, and water use efficiency.
+We also predict that canalized physiological function affects the ecology of lycophytes in a manner that prevents them from establishing in full sun environments.
 
@@
 This study was conducted at two sites in upstate New York, USA.
-Data for most species were gathered at the Edmund Niles Huyck Preserve (ENHP), characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range.
+Data for most species were gathered at the Edmund Niles Hyuck Preserve (ENHP), characterized by mixed hardwood hemlock forests along the Helderberg escarpment in the foothills of the Catskill mountain range.
 Data from the second site were gathered at a forested natural area on the campus of Colgate University in Hamilton, New York, situated in the foothills surrounding the Chenango Valley.
@@
 We examined a series of fern, lycophyte, and angiosperm species at both sites (Table 1).
-The main focus of this work was to examine ferns relative to light-analogous angiosperms.
-To accomplish this, we selected a fern species and then located the nearest neighboring angiosperm growing in similar habitat conditions.
+To examine ferns relative to light-analogous angiosperms, we first selected a fern species and then located the nearest neighboring angiosperm growing in similar habitat conditions.
 Surveyed angiosperm taxa were always herbaceous and similar in size and stature to the neighboring fern species.
@@
 Light response curves were conducted for five individuals of each species.
-Next, to increase our comparative sample size, we surveyed light saturated rates of photosynthesis (A~n~) and stomatal conductance (g~s~) for additional species at the ENHP.
+Second, to increase our comparative sample size, we surveyed light saturated rates of photosynthesis (A~n~) and stomatal conductance (g~s~) for additional species at the ENHP.
 All gas exchange parameters were generated using the LiCor 6400 XT Photosynthesis System (LiCor Biosciences, Lincoln, NE USA).
@@
 For each light response curve, The CO~2~ in the leaf cuvette was set to a flow rate of 300 $\mu$mol s^-1^ and at ambient atmospheric [CO~2~] (400 ppm).
-Temperature was not explictly controlled in the leaf cuvette.
+Temperature was not explicitly controlled in the leaf cuvette.
 Cuvette leaf temperatures ranged from 27-30 &deg;C and leaf vapor pressure deficit (VPD) averaged 1.37&pm;0.40 kPa.
@@
 Light response data for each curves was fit to a non-rectangular hyperbola model (Equation 6 in @lobo_fitting_2013).
-The light compensation point (LCP) was calculated as the PPFD at which the net photosynthetic rate equalled zero from the linear phase of each light response curve.
-Quantum yeild ($\phi$) was calculated as the initial slope of the assimilation rate and mitochondrial respiration in the light (R~d~) at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.
+The light compensation point (LCP) was calculated as the PPFD at which the net photosynthetic rate equaled zero from the linear phase of each light response curve.
+Quantum yield ($\phi$) was calculated as the initial slope of the assimilation rate and mitochondrial respiration in the light (R~d~) at the step change where PPFD was 0 $\mu$mol m^-1^ s^-1^.
 
@@
 From these light response curves, we also developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates.
-As a result, additional survey measurments were conducted at at a PPFD of 1200 $\mu$mol m^-1^ s^-1^ for high-light species and 600 $\mu$mol m^-1^ s^-1^ for low-light species.
+As a result, additional point measurements of leaf gas exchange were conducted at at a PPFD of 1200 $\mu$mol m^-1^ s^-1^ for high-light species and 600 $\mu$mol m^-1^ s^-1^ for low-light species.
 All other cuvette parameters were identical to those used for the light response curves.
-Leaf-level instantaneous water use efficiency (WUE) was calculated as A~n~ divided by transpiration from gas exchange measurements.
+Leaf-level instantaneous water use efficiency (WUE) was calculated as A~n~ divided by leaf transpiration from gas exchange measurements.
 
@@
 Principal component analysis, utilizing the ‘vegan’ package [@oksanen_vegan:_2019], was used to explore how measured functional traits were distributed and co-varied among plant groups and canopy type.
-Linear mixed-effect models responses were used to test responses of functional traits to caetgorial fixed effects of plant group and canopy habitat.
-To test broad differnces in plant groups, species nested within canopy habitats were treated as random effects.
-To test differences between habitat types (ferns and angiosperms only) species was treated as a fixed effect.
+Linear mixed-effect models were used to test responses of functional traits to categorical fixed effects of plant group,  habitat type and/or reproduction type (spore vs seed).
+The interaction between plant group and habitat was tested for functional trait comparisions between analagous fern and angiosperm species.
+To test for broad differences among plant groups, individual plant species were treated as random effects in each model.
+Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among fixed effect treatments with the ‘multcomp’ package [@hothorn_simultaneous_2008].
+Pairwise comparisons with the ‘emmeans’ package [@lenth_emmeans:_2019] were calculated to investigate interactions between plant group and habitat, if present.
+For bivariate trait relationships, responses of dependent variables were analysed with linear mixed-effect models, with species as a random effect and plant group and/or canopy type as categorical fixed effects.
 Explained variance (R^2^) of mixed models were computed as in @nakagawa_general_2013, in which the marginal R^2^ represents variance explained by fixed factors and the conditional R^2^ by both fixed and random factors.
-Tukey’s post-hoc test were performed in conjunction with ANOVA to determine which mean values of functional traits were different among plant groups and canopy habitats with the ‘multcomp’ package [@hothorn_simultaneous_2008].
-T-tests were performed to test for differences between functional traits across canopy habitats within angioperm and fern plant groups.
-For bivariate trait relationships, responses of dependent variables were analysed with linear mixed-effect models, with species as a random effect and plant group and/or canopy type as categorical fixed effects.
-Differences in slopes of the relationship of bivariate traits by habitat were tested by calculating estimated marginal means and computing pairwise comparisons with the ‘emmeans’ package [@lenth_emmeans:_2019].
-All tests of statistical significance were conducted at an α level of 0.05.
+All tests of statistical significance were conducted at an $\alpha$ level of 0.05.
 All analyses were performed with R 3.5.1 [@r_core_team_r:_2018].
@@
 
-We first explored how each of 10 measured functional traits varied and were distributed among co-occuring ferns, angiosperms and lycophytes in closed habitats.
-In total, 52.9 % of the trait variation was accounted for in two dimensions, encompassing different resource economics, physiology and anatomy of the surveryed taxa.
-Interestingly, there was minimal overlap among component scores for angiosperms species compared to spore-bearing ferns and lycophytes (Figure `r pcagroup`).
-Specifically, angiosperms species were associated with gas exchange, anatomical and stoichiometric functional traits related to optimizing carboxylation.
-In contrast, lycophytes and fern species were most associated with physiological traits related to maintaining postive carbon balance.
+We first explored how each of 10 measured functional traits varied and were distributed among co-occurring ferns, angiosperms and lycophytes in closed habitats.
+In total, 52.9 % of the trait variation was accounted for in two dimensions, encompassing differences in resource economics, physiology and anatomy of the surveyed taxa.
+Interestingly, there was minimal overlap among component scores for angiosperms compared to spore-bearing ferns and lycophytes (Figure `r pcagroup`).
+Specifically, angiosperm species were associated with gas exchange, anatomical and stoichiometric functional traits related to optimizing carboxylation.
+In contrast, lycophyte and fern species were most associated with physiological traits related to maintaining postie carbon balance.
 
@@
 Component scores for angiosperms and fern species broadly diverged across open and closed canopy habitats, however, associations with physiological traits in open habitats were stronger for angiosperms (Figure `r pcashade`).
-Specifically, angiosperms appear more plastic in their ability to upregulate traits related to photosynthetic gas exchange with increased light availability.
+Specifically, angiosperms appear more plastic in their ability to up-regulate traits related to photosynthetic gas exchange with increased light availability.
 Overall, these diverging relationships with measured functional traits suggest a highly variable in situ ecophysiology for coexisting angiosperm, fern and lycophyte taxa.
@@
 
-Overall, fern and angiosperm species in open habitats had higher A~n~ than comparable species in closed habitats (*P* < 0.0001).
-Specifically, fern species in high-light habitats had 40.6 % greater A~n~ than in closed habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 60 % higher A~n~ (*P* < 0.0001, t=-6.97).
-Ferns and angiosperms growing in low-light habitats had similar A~n~, however, in open habitats angiosperms had 42.6% higher rates of A~n~ than ferns (canopy x plant group, *P* = 0.043, Figure `r interfig`A).
-Overall, ferns and angiosperms in open habitats also had 64.2% higher g~s~ than comparable species in closed habitats (*P* = 0.002).
-Specifically, fern species in high-light habitats had 50.7 % greater g~s~ than in closed habitats(*P* < 0.0001, t=-4.72), while angiosperm species had 66.9 % higher g~s~ (*P* < 0.0001, t=-4.157).
-Across both habitat types, angiosperm species had a 60% higher g~s~ than fern species (*P* = 0.011, Figure `r interfig`B).
-Neither ferns nor angiosperms species differed in WUE across the two light environments.
-Across all habitats, WUE was also statistically similar for ferns and angiosperms.
+Ferns and angiosperms growing in low-light habitats had similar A~n~, however, in open habitats angiosperms had 74.0 % higher rates of A~n~ than ferns (canopy x plant group, *P* = 0.043, Figure `r interfig`A).
+Rates of A~n~ for angiosperms in open habitats were 154.2 % higher than angiosperms in closed habitats (*P* =0.001, t=-4.742), while rates of A~n~ for ferns were statistically similar across habitat types.
+Overall, ferns and angiosperms in open habitats had nearly 3-fold higher g~s~ than comparable species in closed habitats (*P* = 0.003).
+Rates of g~s~ for angiosperms in open habitats increased nearly two-fold compared to angiosperms in closed habitats (*P* = 0.0341, t=-2.973), while rates of g~s~ for ferns were statistically similar across habitat types (Figure `r interfig`B).
+Neither ferns nor angiosperms differed in WUE across the open and closed habitats, and WUE was  statistically similar bewteen the two plant groups.
 
@@
 
-In closed habitats, variation in gas exchange parameters was largely associated with individual species variation (random effect) instead of functional differences between all three plant groups.
-No differences between A~n~ were detected between between the three plant groups.
+In closed habitats, variation in gas exchange parameters was largely associated with individual species variation (random effect) instead of functional differences between plant groups.
+No differences between A~n~ were detected between between angiosperms, ferns or lycophytes.
 Rates of g~s~ were highest in angiosperms (*P* = 0.001, Figure `r shadefig`A).
-Overall, the spore-bearing plant lineages had lower g~s~ (), regardless of habitat (Figure `r seedsporefig`A).
-Water-use efficiency was 47 % higher in lycophytes than ferns and angiosperms (*P* = 0.005), while ferns and angiosperms where statiscally similar (Figure `r shadefig`B).
+Overall, the spore-bearing plant species had 50.5 % lower g~s~ than angiosperms (*P* = 0.004, Figure `r seedsporefig`A).
+Water-use efficiency was 88.5 % higher in lycophytes than ferns and angiosperms (*P* = 0.005), while ferns and angiosperms where statistically similar (Figure `r shadefig`B).
 
@@
 Overall, ferns and angiosperms had similar $\phi$ across open and closed habitats.
-Quantum yields were also equivalent within angiosperm and fern species groups between open and closed habitats.
-Shifts in R~d~ between open and closed habitats differed for angiosperms and ferns (habitat x plant group, *P* < 0.001, Figure `r interfig`C).
-Specifically, R~d~ in angiosperms was 80.6 % greater in open compared to closed habitats (t=-4.197, *P* = 0.001).
-Alternatively, R~d~ in ferns was 44.0 % greater in low-light compared to open habitats (t=-2.97, *P* = 0.006).
-Similarly, shifts in LCP between open and closed habitats differed for angiosperms and ferns (habitats x plant group, *P* < 0.001, Figure `r interfig`D).
-Specifically, the LCP in angiosperms was 79.2 % greater in open compared to closed habitats (t=-3.5998, *P* = 0.005).
-Alternatively, the LCP in ferns was 39.7 % greater in closed compared to open habitats (t=2.1814, *P* = 0.040).
+Quantum yields were also equivalent across angiosperms and ferns in open and closed habitats.
+Shifts in R~d~ between open and closed habitats differed for angiosperms but not for ferns (habitat x plant group, *P* < 0.001, Figure `r interfig`C).
+Specifically, R~d~ in angiosperms was 4 fold greater in open compared to closed habitats (t=-4.802, *P* = 0.013).
+Similarly, shifts in LCP between open and closed habitats differed for angiosperms but not for ferns (habitats x plant group, *P* < 0.001, Figure `r interfig`D).
+Specifically, the LCP in angiosperms increased nearly 4-fold in open compared to closed habitats (t=-4.382, *P* = 0.019).
 
@@
 
-In closed canopy environments, the LCP of spore-bearing plant species were equivalent and marginally higher than angiosperm species (*P* = 0.061, Figure `r seedsporefig`B).
-Photosynthetic quantum was similar between each plant group, but lycophytes and ferns were broadly higher than angiosperms.
-Ferns and lycophytes had equavalent respiration rates, while both groups were higher than angiosperms (*P* = 0.0178)
+In closed canopy habitats, the LCP of spore-bearing plant species were equivalent and 161.9 % higher than angiosperm species (*P* = 0.0195, Figure `r seedsporefig`C).
+Photosynthetic quantum was statistically similar between each plant group, but lycophytes and ferns were broadly higher than angiosperms.
+Spore-bearing plant species had equivalent R~d~, and R~d~ was nearly 4-fold lower than in angiosperms (*P* = 0.0178, Figure `r seedsporefig`D)
 
-## Stoichiometry (##can use pairwise for specific p values: see leaf P below)
+## Stoichiometry
 
-Ferns and angiosperms species in open habitats had marginally lower foliar N content compared to similar species in growing in closed light environments (-17.2 %, *P* = 0.072), with a large proportion of variation attributed to individual species differences.
+Ferns and angiosperms in open habitats had marginally lower foliar N content compared to similar species growing in closed habitats (-17.2 %, *P* = 0.073), with a large proportion of variation attributed to individual species differences.
 No differences in foliar P content were detected between open and closed habitats for either ferns or angiosperms.
@@
 Neither ferns nor angiosperms species differed in their N:P ratios across the two habitat types.
+
 \
 
-Within closed habitats, lycophytes have 45% lower foliar N content than ferns and angiosperms, which do not statistically differ (*P* < 0.001, Figure `r shadefig`C).
-Lycophytes also had 24% lower foliar P content than angiosperms (*P* = .026), while ferns were simiilar to both groups (Figure `r shadefig`D).
+Within closed habitats, lycophytes have 46.3 % lower foliar N content than ferns and angiosperms, which do not statistically differ (*P* < 0.001, Figure `r shadefig`C).
+Lycophytes also had 24.0 % lower foliar P content than angiosperms (*P* = .026), while ferns were similar to both groups (Figure `r shadefig`D).
+Overall, spore-bearing species had 18.7 % lower foliar P content than angiosperm species (*P* = 0.008, Figure `r seedsporefig`B).
 The N:P ratios were not statistically different across angiosperms, ferns and lycophytes in closed habitats.
+
+\
+
+The relationship between A~n~ and foliar N content depended on both plant group and habitat type (habitat x plant group, *P* = 0.028).
+Specifically, the positive slopes of the relationship between A~n~ and N differed for ferns and angiosperms in open habitats, while no relationship between A~n~ and N was found for either group in closed habitat (Figure `r anitro`).
 
@@
 
-Overall, no differences where detected in stomatal density (SD) between species growing in open and closed habitats.
-Specifically, differences in SD were not detected for either fern nor angiosperms species between the two habitats.
-However, in both open and closed environments, ferns had significantly lower SD than analogous angiosperms (*P* < 0.001).
-Additionally, no differenes in SD were detected among species of ferns, angiosperms and lycophytes in closed habitats.
-Variation in SD in the shade was largely due to individual species variation.
-
-## Bivariate relationships between functional traits
-
-The relationship between A~n~ and foliar N content depended on both plant group and habitat type (habitat x plant group, *P* = 0.028).
-Specifically, the postive slopes of the relationship between A~n~ and N differed for ferns and angiosperms in open habitats, while no relationship beween A~n~ and N was found for either group in closed habitat (Figure `r anitro`).
-
-No relationships.....
+Overall, differences in stomatal density (SD) were not detected for either ferns or angiosperms between open and closed habitats.
+Across habitats, however, ferns had a significantly lower SD than analogous angiosperms (-52.0 %, *P* < 0.001).
+Additionally, no differences in SD were detected among species of ferns, angiosperms and lycophytes in closed habitats.
+Variation in SD in closed habitats was largely due to individual species variation.
 
@@
 
-## Comparative ecophysiology in the shade
-
-## Differential adaption to open habitats
-
-## Functional ecology of lycophytes
-
-## Ecophysiological trait variation in seed plants
-
-## Summary
+* which way should we tell the story?
+1) fundamental differences in shade, then how angiosperms succeed in light
+2) or vice versa
 
@@
 
-**Table 1**. Surveyed species of angiosperms, ferns and lycophytes across both open canopy high-light and closed canopy low-ligh habitats.
+**Table 1**.
+Surveyed species of angiosperms, ferns and lycophytes across both open canopy high-light and closed canopy low-light habitats.
+All surveyed taxa were  herbaceous and similar in size and stature.
 
@@
 traitstab <- read.csv("manuscript/pooled_data.csv")
-# source("manuscript/paper_labels.R")
-# physymb <- '\u03D5'
-
-phototab <- expression(atop(italic(A[n]),(mu*mol~m^-2~s^-1)))
 
@@
 
-```{r pca1, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs,warning=FALSE, message=FALSE}
+```{r pca1, fig.width=5, fig.height=5, echo=FALSE,  dev="png", dpi=400,eval=includeFigs,warning=FALSE, message=FALSE}
 source("master_scripts/pca_plot.R")
@@
 **Figure 1**.
-Redundancy analysis of measured functional traits across con-occuring similar statured angiosperm, fern and lycophyte species in closed habitats.
+Redundancy analysis of measured functional traits across con-occurring herbaceous angiosperm, fern and lycophyte species in closed habitats.
 With the 10 measured variables, 52.9 % of the trait variation was accounted for in two dimensions.
@@
 
-```{r pca2, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs,warning=FALSE, message=FALSE}
+```{r pca2, fig.width=5, fig.height=5, echo=FALSE,  dev="png", dpi=400,eval=includeFigs,warning=FALSE, message=FALSE}
 source("master_scripts/pca_canopy.R")
@@
 **Figure 2**.
-Redundancy analysis of measured functional traits across con-occuring similar statured angiosperms and fern species in open and closed habitats.
+Redundancy analysis of measured functional traits across con-occurring herbaceous angiosperms and fern species in open and closed habitats.
 With the 10 measured variables, 55.9 % of the trait variation was accounted for in two dimensions.
@@
 
-```{r interaction, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
+```{r interaction, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=400,eval=includeFigs, warning=FALSE, message=FALSE}
 source("master_scripts/interaction_plots.R")
 ```  
-**Figure 3**. Differences in mean physiological traits between co-occuring similar statured angiosperms and fern species across open and closed habitats.
+**Figure 3**. Differences in mean physiological traits between co-occurring herbaceous angiosperms and fern species across open and closed habitats.
 
-```{r lyco, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
-source("master_scripts/shade_lineages.R")
+```{r lyco, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=400,eval=includeFigs, warning=FALSE, message=FALSE}
+source("master_scripts/shade_lineages2.R")
 ```  
-**Figure 4**. Differences in mean physiological traits between con-occuring similar statured angiosperm, fern and lycophyte species in closed habitats.
+**Figure 4**. Differences in mean physiological traits between con-occurring herbaceous angiosperm, fern and lycophyte species in closed habitats.
 
-```{r seed, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
-source("master_scripts/seed_nonseed_plots.R")
+```{r seed, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=400,eval=includeFigs, warning=FALSE, message=FALSE}
+source("master_scripts/seed_nonseed_plots2.R")
 ```  
-**Figure 5**. Differences in mean physiological traits between co-occuring seed (angiosperms) and non-seed plant species (ferns + lycophytes) in shade habitats.
+**Figure 5**. Differences in mean physiological traits between co-occurring herbaceous seed plants (angiosperms) and spore-bearing plants (ferns + lycophytes) in closed habitats.
 
-```{r nitro, fig.width=7, fig.height=7, echo=FALSE,  dev="png", dpi=600,eval=includeFigs, warning=FALSE, message=FALSE}
+```{r nitro, fig.width=5, fig.height=5, echo=FALSE,  dev="png", dpi=400,eval=includeFigs, warning=FALSE, message=FALSE}
 source("master_scripts/bivariate_plots.R")
@@
 **Figure 6**.
-Postive relationships between light saturated photosynthesis (A~n~) and foliar nitrogen content exists for small statured angiosperm and fern species in open habitat, but not in closed habitats.
+Positive relationships between light saturated photosynthesis (A~n~) and foliar nitrogen content existed for herbaceous angiosperm and fern species in open habitat, but not in closed habitats.
 For significant linear relationships, dashed lines represent model fits between traits and grey shaded areas are 95 % confidence intervals for the mean.
-Conditional and marginal R^2^ are reported for the overall model.
+Conditional and marginal R^2^ are reported for the overall model fit.
 

### 2019-03-27 master_scripts/seed_nonseed_plots2.R

@@ -29,17 +29,18 @@ lrc_shade <-  droplevels(lrc[!lrc$canopy == "Closed",])
 
 #seed vs nonseed panel figure
 # windows()
-par(mfrow=c(2,2), mgp=c(3,.75,0))
+par(mfrow=c(2,2), mgp=c(3,.75,0),oma=c(4,5,1,5))
 
 #stomatal conductance
-par(mar=c(0,5,1,0))
-boxplot(Cond ~ seed, data=chem_closed, ylab=condlab,xaxt='n',yaxt='n',
+par(mar=c(0,0,0,0))
+boxplot(Cond ~ seed, data=chem_closed, ylab="",xaxt='n',yaxt='n',
         ylim=c(0,.3), outline=FALSE, col=seedcols)
 axis(2, las=1)
 text(x=0.5, y=.3, label = "A", cex=1.25)
+mtext(side=2, at=.15, line=3,text=condlab, xpd=TRUE, las=3, cex=.9)
 
 #phosphorus
-par(mar=c(0,0,1,5))
+par(mar=c(0,0,0,0))
 boxplot(p_perc ~ seed, data=phos_closed, ylab="", xaxt='n', yaxt='n',
         ylim=c(0,0.5),  col=seedcols)
 axis(4,las=1)
@@ -47,19 +48,20 @@ mtexti("Foliar Phosphorus  (%)", 4, outer=TRUE, cex=1, off=.65)
 text(x=0.5, y=0.5, label = "B", cex=1.25)
 
 #light compensation point
-par(mar=c(5,5,0,0))
-boxplot(LCP ~ seed, data=lrc_shade, ylab=lcplab,outline=FALSE,yaxt='n',
-        ylim=c(0.,27),  col=c("white", plantcols2[2]))
+par(mar=c(0,0,0,0))
+boxplot(LCP ~ seed, data=lrc_shade, ylab="",outline=FALSE,yaxt='n',
+        ylim=c(0,27),  col=c("white", plantcols2[2]))
 axis(2, las=1)
 text(x=0.5, y=26.95, label = "C", cex=1.25)
+mtext(side=2, at=13.5, line=3,text=lcplab, xpd=TRUE, las=3, cex=.9)
 
 #dark respiration
-par(mar=c(5,0,0,5))
+par(mar=c(0,0,0,0))
 boxplot(Rd ~ seed, data=lrc_shade, ylab="", outline=FALSE,
[…]

### 2019-03-27 master_scripts/shade_lineages2.R

@@ -1,0 +1,52 @@
+source("master_scripts/plot_objects.R")
+source("functions/basic_functions.R")
+
+##lycophytes differ
+
+chem <- read.csv("raw_data/photo_chem.csv")
+
+chem_closed <- chem[chem$canopy == "Closed",]
+chem_closed$cnratio <- with(chem_closed, c_perc/n_perc)
+chem_closed$wue <- with(chem_closed, Photo/Trmmol)
+
+phos <- chem_closed[chem_closed$p_perc > .1 & chem_closed$p_perc < .4,]
+
+#boxplots of significant traits
+
+# windows()
+par(mfrow=c(2,2), mgp=c(3,.75,0),oma=c(4,5.5,1,5.5))
+
+#gs
+par(mar=c(0,0,0,0))
+boxplot(Cond~ plant_group, data=chem_closed, ylab="", xaxt='n', yaxt='n',
+        col=pgcols, ylim=c(0,0.26), outline=FALSE)
+axis(2, las=1)
+text(x=0.5, y=0.26, label = "A", cex=1.25)
+mtext(side=2, at=.13, line=3,text=condlab, xpd=TRUE, las=3, cex=.9)
+
+#wue
+par(mar=c(0,0,0,0))
+boxplot(wue ~ plant_group, data=chem_closed, ylab="",xaxt='n', 
+        col=pgcols, ylim=c(0,16), outline=FALSE, yaxt='n')
+axis(4,las=1)
+mtexti(wuelab, 4, outer=TRUE, cex=1, off=.6)
+text(x=0.5, y=16, label = "B", cex=1.25)
+
+#nitro
+par(mar=c(0,0,0,0))
+boxplot(n_perc ~ plant_group, data=chem_closed, ylab="",
+        col=pgcols, ylim=c(0,5), yaxt='n', names=plants,outline=FALSE)
+axis(2, las=1)
+text(x=0.5, y=4.9, label = "C", cex=1.25)
+mtext(side=2, at=2.5, line=3,text="Foliar N content  (%)", xpd=TRUE, las=3, cex=.9)
+
+#phos
+par(mar=c(0,0,0,0))
+boxplot(p_perc ~ plant_group, data=phos, ylab="",
+        col=pgcols, ylim=c(0,.425), yaxt='n',names=plants)
+axis(4,las=1)
+mtexti("Foliar P content  (%)", 4, outer=TRUE, cex=1, off=.6)
[…]

### 2019-03-27 stats/element_stats_habitats.R

@@ -41,6 +41,7 @@ visreg(N_mod, "canopy", by="plant_group")
 library(emmeans)
 emmip(N_mod, canopy ~ plant_group)
 emmeans(N_mod, pairwise ~ canopy : plant_group)
+emmeans(N_mod, "canopy")
 
 Anova(N_mod, type="3")
 summary(N_mod)

### 2019-03-27 stats/gs_stats.R

@@ -40,7 +40,8 @@ shade_siglets <-cld(tukey_shade)
 #"b"  "a"   "a"
 
 
-#canopy comparisions of sun shade with angio and ferns
+#canopy comparisions of sun shade with angio and ferns-------
+
 fernangio <- droplevels(gs[!gs$plant_group %in% "Lyco", ])
 agricolae::skewness(sqrt(fernangio$Cond))
 
@@ -105,3 +106,17 @@ t.test(Cond ~ canopy, data=angio)
 #0.09781456           0.29560109 
 # t = -4.157, df = 21.105, p-value = 0.0004425
 #66.9%
+
+
+#match figure yes
+mean(fernangio[fernangio$canopy == "Open" & 
+                 fernangio$plant_group == "Fern", "Cond"])
+mean(fernangio[fernangio$canopy == "Closed" & 
+                 fernangio$plant_group == "Fern", "Cond"], na.rm=TRUE)
+
+#match figure yes
+mean(fernangio[fernangio$canopy == "Open" & 
+                 fernangio$plant_group == "Angio", "Cond"])
+mean(fernangio[fernangio$canopy == "Closed" & 
+                 fernangio$plant_group == "Angio", "Cond"], na.rm=TRUE)
+#61%higher in angio open

### 2019-03-27 stats/lrc_stats.R

@@ -11,7 +11,7 @@ library(stringr)
 treatments <- read.csv("raw_data/species_list.csv")
 
 
-#photosynthesis
+#aq curve
 lrc <- read.csv("calculated_data/AQ_params.csv") %>%
        mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
               individual =  str_remove(ID, ".*-")) %>%
@@ -193,9 +193,10 @@ r.squaredGLMM(rd_mod2)
 #           R2m       R2c
 #[1,] 0.4931456 0.5970595
 
-mean(fernangio[fernangio$plant_group == "Fern" , "Phi"], na.rm=TRUE) #.0692
-mean(fernangio[fernangio$plant_group == "Angiosperm" , "Phi"], na.rm=TRUE) #.0620
-#really close
+mean(fernangio[fernangio$plant_group == "Angiosperm" & fernangio$canopy == "Open", 
+               "Rd"], na.rm=TRUE) #0.7687267
+mean(fernangio[fernangio$plant_group == "Angiosperm" & fernangio$canopy == "Closed", 
+               "Rd"], na.rm=TRUE) #0.1487808
 
 #lcp
 
@@ -218,3 +219,7 @@ mean(fernangio[fernangio$plant_group == "Fern" , "LCP"], na.rm=TRUE)
 mean(fernangio[fernangio$plant_group == "Angiosperm" , "LCP"], na.rm=TRUE)
 #really close
 
+mean(fernangio[fernangio$plant_group == "Angiosperm" & fernangio$canopy == "Open", 
+               "LCP"], na.rm=TRUE) #12.99204
+mean(fernangio[fernangio$plant_group == "Angiosperm" & fernangio$canopy == "Closed", 
+               "LCP"], na.rm=TRUE) #2.698899

### 2019-03-27 stats/photo_stats.R

@@ -36,7 +36,7 @@ visreg(shade_mod)
 
 tukey_shade <- glht(shade_mod, linfct = mcp(plant_group = "Tukey"))
 shade_siglets <-cld(tukey_shade)
-shade_siglets2 <- shade_siglets$mcletters$Letters
+#similar between plant groups, large random species effects
   
 
 #canopy comparisions of sun shade with angio and ferns
@@ -58,8 +58,10 @@ r.squaredGLMM(photo_mod2)
 #           R2m       R2c
 #[1,] 0.5301052 0.8926862
 
-mean(fernangio[fernangio$canopy == "Open" & fernangio$plant_group == "Angio", "Photo"])
-mean(fernangio[fernangio$canopy == "Open" & fernangio$plant_group == "Fern", "Photo"])
+mean(fernangio[fernangio$canopy == "Open" & 
+                 fernangio$plant_group == "Angio", "Photo"])
+mean(fernangio[fernangio$canopy == "Open" & 
+                 fernangio$plant_group == "Fern", "Photo"])
 
 
 #t-test between groups
@@ -76,23 +78,15 @@ t.test(Photo ~ canopy, data=angio)
 # t = -6.9651, df = 20.368, p-value = 8.347e-07
 
 
-#excludes light bc not in both canopies
-# photo_mod <- lmer(Photo ~ canopy + (1|plant_group/species), data=fernangio)
-# visreg(photo_mod)
-#   
-# Anova(photo_mod)
-# summary(photo_mod)
-# r.squaredGLMM(photo_mod)
-#   
-# tukey_photo <- glht(photo_mod, linfct = mcp(canopy = "Tukey"))
-# photo_siglets <-cld(tukey_photo)
-# photo_siglets2 <- photo_siglets$mcletters$Letters
-#   # photosynthesis way higher in open canopies
-# #Closed   Open 
-# # "a"    "b"
-# #p < 0.0001
-# #R2m       R2c
-# #[1,] 0.3440435 0.8933528
-# mean(fernangio[fernangio$canopy == "Open", "Photo"])
-# mean(fernangio[fernangio$canopy == "Closed", "Photo"])
-# #54.8% higher in full sun
+#match figure yes
+mean(fernangio[fernangio$canopy == "Open" & 
+                 fernangio$plant_group == "Fern", "Photo"])
+mean(fernangio[fernangio$canopy == "Closed" & 
+                 fernangio$plant_group == "Fern", "Photo"], na.rm=TRUE)
+
[…]

### 2019-03-27 stats/seed_spore_stats.R

@@ -24,54 +24,39 @@ lrc <- read.csv("calculated_data/AQ_params.csv") %>%
 seed <- droplevels(lrc[lrc$seed == "seed" & lrc$canopy == "Closed",])
 spore <- droplevels(lrc[lrc$seed == "spore" & lrc$canopy == "Closed",])
 
-#PHI
-t.test(seed$Phi, spore$Phi, var.equal=FALSE)
-#not significant
 
-#LCP
-t.test(seed$LCP, spore$LCP, var.equal=FALSE)
+#LCP models, with species effects and outlier
+lrc_shade <- droplevels(lrc[lrc$canopy == "Closed" &  !lrc$ID == "imp_pal-1",])
 
-#Rd
-t.test(seed$Rd, spore$Rd, var.equal=FALSE)
+boxplot(LCP~seed, data=lrc_shade)
+boxplot(Rd~seed, data=lrc_shade)
 
 
-
-#LCP models, with species effects
-lrc_shade <- droplevels(lrc[lrc$canopy == "Closed",])
-
-#simple
-lcp_mod <- lm(LCP ~ seed, data=lrc_shade)
+#lcp model ----------
+lcp_mod <-lmer(sqrt(LCP) ~ seed + (1|species), data=lrc_shade)
 plot(lcp_mod)
-Anova(lcp_mod, type=3)
-visreg(lcp_mod)
-
-#full model
-lcp_mod2 <-lmer(sqrt(LCP) ~ seed + (1|species), data=lrc_shade)
-plot(lcp_mod2)
 #used sqrt to deal with residuals versus fitted values
 
-Anova(lcp_mod2, type=3)
-summary(lcp_mod2)
-r.squaredGLMM(lcp_mod2) 
+Anova(lcp_mod, type=3)
+summary(lcp_mod)
+r.squaredGLMM(lcp_mod) 
 
 #          R2m       R2c
 # [1,] 0.2533775 0.4984394
 # P = 0.01073 *  
 
-tukey_lcp <- glht(lcp_mod2, linfct = mcp(seed = "Tukey"))
+tukey_lcp <- glht(lcp_mod, linfct = mcp(seed = "Tukey"))
 lcp_siglets <-cld(tukey_lcp)
 # seed spore 
 # "a"   "b"
 
+mean(lrc_shade[lrc_shade$seed == "seed","LCP"])
+mean(lrc_shade[lrc_shade$seed == "spore","LCP"])
 
[…]

### 2021-06-03 manuscript/manuscript.Rmd
slide figure edit

@@
 ---
-title: "Plasticity of ecophysiological traits structure how herbaceous plant lineages make a living in shaded environments "
+title: "Plasticity of ecophysiological traits structure how herbaceous plant lineages persist in shaded environments "
 author: 
@@
 
-Our understanding of the ecophysiology of ferns and lycophytes is limited.
-However, recent studies have shown that these two groups possess an unusual set of functional traits that likely influence their ecology (Brodribb, Watkins, Pinterman, Franks, campany).
-For example, Watkins has demonstrated that the gametophyte generations of many tropical ferns are markedly desiccation-tolerant, with tolerance being closely linked to ecological preference.
+The earliest angiosperm ancestors shared ecophysiological traits associated with shady, disturbed, and wet habitats [@feild_dark_2004].
+Plasticity in these traits, developed through natural selection, then allowed angiosperms to expand into brighter and more evaporative terrestrial habitats characterized as water-limited instead of light limited [@feild_ecophysiology_2007].
+The same degree of flexibility has not been widely detected in the modern diversification of spore-bearing plant taxa.
+As a consequence, the Cretaceous radiation of angiosperms initiated a tremendous shift in of terrestrial plant communities, which included reciprocal trends in diversity of spore-bearing plants [@lupia_comparing_1999].
+Most modern taxa of spore-bearing plants are now believed to have diversified following this angiosperm radiation, so that angiosperm dominant communities now define the ecological niches that spore-bearing plants must inhabit [@schneider_ferns_2004].
+
+\
+Our understanding of the ecophysiology of the modern spore-bearing plants, including ferns and lycophytes, is limited.
+However, studies have shown that these two groups possess unique functional traits and physiological behaviors that likely influence their ecology.
+Ferns have been shown to be more conservative in stomatal closure, relative to xylem dysfunction, than angiosperms [@brodribb_stomatal_2004].
+Xylem of ferns have been shown to have greater hydraulic efficiency than conifer or angiosperm xylem, while still maintaining comparable resistance to cavitation [@pittermann_structure-function_2011].
+Microphylls of lycophytes also likely led to distinct ecophysiology between ferns with megaphylls, even in shared habitats [@campany_convergence_2019].
+Last, ferns and lycophytes have limited flexibility to water stress, compared to angiosperms, as a consequence of a passive stomatal function [@mcadam_ancestral_2013].
+These studies reveal several remaining canalized functions in ferns and lycophytes, which undoubtedly affect how spore-bearing taxa persist in a now angiosperm dominated landscape.
+
+\
+Studies that examine that broadly comparate the physiology of ferns and lycophytes to seed plants are rare.
+Therefore, it is difficult to derive widely applicable hypotheses of the ecophysiological behavior of these groups.
+From the present studies available, ferns and lycophytes appear to have lower photosynthetic rates than seed plants [@brodribb_leaf_2005; @brodribb_leaf_2007; @carriqui_diffusional_2015; @campany_convergence_2019].
+This would suggest that non-seed plants may be at a competitive disadvantage when growing alongside seed plants.
+However, only @carriqui_diffusional_2015 has systematically compared the physiology of these groups when growing in similar habitats (a common garden experiment).
+
+ Free-sporing plants exhibit trends reciprocal to those of angiosperms of equivalent latitudes
+
+@watkins_ecological_2007 has demonstrated that the gametophyte generations of many tropical ferns are markedly desiccation-tolerant, with tolerance being closely linked to ecological preference.
 In seed plants, desiccation tolerance has been essentially lost from vegetative tissues.
@@
 
+**add these cites as evidence, not as cites in opening sentence.
+
 \
-
 A primary controlling factor in fern hydraulics is stomatal function.
-A series of elegant studies have shown that stomatal control of ferns and lycophytes differs fundamentally from seed plants.
-Evidence on fern and lycophyte stomata indicates that these groups have inefficient stomatal systems, likely possessing a hydro-passive mechanism which influences stomatal opening and closing (many cites).
-Specifically, unlike seed plants, fern and lycophyte stomata fail to close in response to abscisic acid (ABA) addition (many cites).
-Additionally, recent evidence also suggests that co-occurring tropical ferns and lycophytes optimize their ecophysiology differently (Campany).
+A series of elegant studies have shown that stomatal control of ferns and lycophytes differs fundamentally from seed plants [@ruszala_land_2011; @mcadam_fern_2012] .
+Evidence on fern and lycophyte stomata indicates that these groups have inefficient stomatal systems, likely possessing a hydro-passive mechanism which influences stomatal opening and closing.
+Specifically, unlike seed plants, fern and lycophyte stomata fail to close in response to abscisic acid (ABA) addition [@mcadam_fern_2012].
+Additionally, recent evidence also suggests that co-occurring tropical ferns and lycophytes optimize their ecophysiology differently [@campany_convergence_2019].
 Consequently, the evolution or adaptation of physiological traits may impact the differential distribution of ferns,lycophytes and seed plants to each other.
@@
 \
-
-Studies that examine the comparative physiology of ferns and lycophytes to seed plants are rare.
-Therefore, it is difficult to derive widely applicable hypotheses of the ecophysiological behavior of these groups.
-From the limited number of present studies, ferns and lycophytes appear to have lower photosynthetic rates than seed plants (cite).
-This would suggest that non-seed plants may be at a competitive disadvantage when growing alongside seed plants.
-However, none of these studies have systematically compared the physiology of these groups when growing in similar habitats.
-
-\
-
 The goal of this study is to examine the comparative ecophysiology of a number of temperate ferns and lycophytes to angiosperms.
@@
 In a pairwise comparison, we then evaluated the potential of fern and angiosperm species to adapt to open-light habitats.
-Given increasing evidence of inefficient stomatal functions in non-seed plants, we hypothesized that ferns and lycophytes would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms.
+Given increasing evidence of canalized physiological function in non-seed plants, we hypothesized that ferns and lycophytes would have lower photosynthetic rates and stomatal conductance values in comparison to the nearby angiosperms.
 This, in turn, would impact aspects of leaf anatomy, nutrient relations, and water use efficiency.
-We also predict that canalized physiological function affects the ecology of lycophytes in a manner that prevents them from establishing in full sun environments.
+We also predict that more severe canalized physiological function affects the ecology of lycophytes in a manner that prevents them from establishing in full sun environments.
 
-#Methods
+# Methods
 
@@
 \
-
 We examined a series of fern, lycophyte, and angiosperm species at both sites (Table 1).
@@
 At the Colgate University site, additional low-light fern and lycophyte species were surveyed.
+
+**HOW MANY**
 
@@
 \
-
 Light response curves were conducted by decreasing the light intensity in the 2x3 cm gas exchange cuvette in small step changes.
@@
 \
-
 From these light response curves, we also developed an understanding of the levels at which low-light and high-light species reached light saturated photosynthetic rates.
@@
 Interestingly, there was minimal overlap among component scores for angiosperms compared to spore-bearing ferns and lycophytes (Figure `r pcagroup`).
-Specifically, angiosperm species were associated with gas exchange, anatomical and stoichiometric functional traits related to optimizing carboxylation.
-In contrast, lycophyte and fern species were most associated with physiological traits related to maintaining postie carbon balance.
+Specifically, component scores for angiosperm species were associated with gas exchange, anatomical and stoichiometric functional traits related to optimizing carboxylation.
+In contrast, lycophyte and fern species were most associated with physiological traits related to maintaining positive carbon balance.
 
 \
-
 Next, we explored how the same suite of 10 functional traits were distributed among open and closed habitats, in both angiosperms and ferns.
@@
 \
-
 In closed habitats, variation in gas exchange parameters was largely associated with individual species variation (random effect) instead of functional differences between plant groups.
@@
 \
-
 In closed canopy habitats, the LCP of spore-bearing plant species were equivalent and 161.9 % higher than angiosperm species (*P* = 0.0195, Figure `r seedsporefig`C).
@@
 \
-
 Within closed habitats, lycophytes have 46.3 % lower foliar N content than ferns and angiosperms, which do not statistically differ (*P* < 0.001, Figure `r shadefig`C).

### 2021-06-03 master_scripts/interaction_plots_talk.R

@@ -57,7 +57,7 @@ jpeg(filename = "manuscript/interactionplot.jpeg",
      width = 8.4, height = 8.4, units = "in", res= 500)
 
 # windows()
-par(mfrow=c(2,2), las=1,mgp=c(3,1,0),oma=c(6,5,1,5))
+par(mfrow=c(2,2), las=1,mgp=c(3,1,0),oma=c(6,5,1,5), cex.axis = 1.25)
 
 #photosynthesis
 par(mar=c(0,0,0,0))
@@ -71,11 +71,11 @@ with(photo_agg, arrows(xaxis, Photo.mean, xaxis,
                        col=canopycols[plant_group],length=0.03, lwd=2))
 points(Photo.mean ~ xaxis, data=photo_agg, cex=1.5, col=canopycols[plant_group], 
        pch=19)
-lines(x=c(1,2), y=c(photo_agg[3,3],photo_agg[4,3]), lwd=3.5, col="forestgreen")
-lines(x=c(1,2), y=c(photo_agg[1,3],photo_agg[2,3]), lwd=3.5, col="firebrick")
+lines(x=c(1,2), y=c(photo_agg[3,3],photo_agg[4,3]), lwd=4, col="forestgreen")
+lines(x=c(1,2), y=c(photo_agg[1,3],photo_agg[2,3]), lwd=4, col="firebrick")
 axis(1, at=1:2, labels = FALSE, tcl=.5)
 axis(2, las=1)
-mtext(side=2, at=7.5, line=3,text=photolab, xpd=TRUE, las=3, cex=1)
+mtext(side=2, at=7.5, line=3,text=photolab, xpd=TRUE, las=3, cex=1.25)
 
 text(x=0.5, y=15, label = "A", cex=1.25)
 
@@ -91,16 +91,16 @@ with(cond_agg, arrows(xaxis, Cond.mean, xaxis,
                        col=canopycols[plant_group],length=0.03, lwd=2))
 points(Cond.mean ~ xaxis, data=cond_agg, cex=1.5, col=canopycols[plant_group], 
        pch=19)
-lines(x=c(1,2), y=c(cond_agg[3,3],cond_agg[4,3]), lwd=3.5, col="forestgreen")
-lines(x=c(1,2), y=c(cond_agg[1,3],cond_agg[2,3]), lwd=3.5, col="firebrick")
[…]

### 2021-06-03 master_scripts/pca_canopy_talk.R

@@ -1,0 +1,98 @@
+source("master_scripts/plot_objects.R")
+
+##merge all datasets into master for PCA
+library(dplyr)
+library(stringr)
+
+#read and format data-------
+
+treatments <- read.csv("raw_data/species_list.csv")
+
+#light response curves removing one bad curve
+lrc <- read.csv("calculated_data/AQ_params.csv") %>%
+  mutate(species = as.factor(str_remove(ID, "-[[:digit:]]")),
+         individual =  as.numeric(str_remove(ID, ".*-"))) %>%
+  merge(treatments[,c(2,4:5)]) %>%
+  filter(!ID ==  "imp_pal-1")
+
+
+#gas exchange and chemistry
+photo <- read.csv("raw_data/photo_chem.csv") %>%
+  mutate(ID = paste(species, individual, sep="-")) %>%
+  filter(!Site == "Colgate") %>%
+  select(-"plant_group") %>%
+  inner_join(treatments[,c(2,5)], by='species')
+
+##now add stomatal density
+stom <- read.csv("raw_data/stomatal_density.csv") %>%
+  select(-"plant_group") %>%
+  inner_join(treatments[,c(2,4:5)], by='species')
+
+stom_agg <- doBy::summaryBy(sto_den ~ species + individual + plant_group + canopy,
+                            FUN=mean, data=stom, keep.names = TRUE) %>%
+  mutate(ID = paste(species, individual, sep="-"))
+
+#merge all data -----
+hyuck <- inner_join(lrc, photo) %>% inner_join(stom_agg)
+
+#select data for PCA
+hyuck_nona <- hyuck[complete.cases(hyuck),]
+  hyuck_nona$WUE <- with(hyuck_nona, Photo/Trmmol)
+  hyuck_nona$NP <- with(hyuck_nona, n_perc/ p_perc)
+
+hyuck_canopy <- hyuck_nona[!hyuck_nona$plant_group == "Lycophyte",]  
+
+#get rid of some parameters (Asat is from aqcurve fitting so keep Photo)
+hyuck_pca <- droplevels(hyuck_canopy[,-c(1:3,6:7,9:13,16,19:21)])
+
+#site variables for ease with ponts in pca
+hyuck_id <- hyuck_canopy[,12:13]
+
+##with alpha
+hyuck_id$plantcols <- ifelse(hyuck_id$plant_group == "Angiosperm", plantcols[2],
[…]

### 2021-06-03 master_scripts/pca_plot_talk.R

@@ -86,21 +86,21 @@ spp <- scores(hyuck_rda, display='species')
 
 #plotting compare ferns and angio lyco in shade
 # windows()
-# jpeg(filename = "output/pcashade.jpeg",
-#      width = 6, height = 6, units = "in", res= 600)
+jpeg(filename = "manuscript/pcashade.jpeg",
+     width = 7, height = 7, units = "in", res= 600)
 
-par(mar=c(5,5,1,1), las=1,cex.axis=0.8)
+par(mar=c(5,5,1,1), las=1,cex.axis=1.25, cex.lab = 1.5)
 plot(sites,ylab="PC 2 (18.4 %)", xlab="PC 1 (34.5%)",type='n',
      xlim=c(-2.25, 2.25), ylim=c(-2.25, 2.25))
 abline(v=0, lty='dashed')
 abline(h=0, lty='dashed')
-ordihull(hyuck_rda, groups = hyuck_shade$plant_group, lwd=2,draw='polygon',
+ordihull(hyuck_rda, groups = hyuck_shade$plant_group, lwd=2.5,draw='polygon',
          col=pgcols,alpha=50, border = pgcols)
-arrows(0, 0, len * spp[, 1],  len * spp[, 2], length = 0.05, lwd=1.5)
-points(sites,cex=1.75, bg=hyuck_id$plantcols, pch=21)
-text(spp,labels=pcalabs,cex=1)
+arrows(0, 0, len * spp[, 1],  len * spp[, 2], length = 0.05, lwd=3.5)
+points(sites,cex=2, bg=hyuck_id$plantcols, pch=21)
+text(spp,labels=pcalabs,cex=1.5)
 legend("topleft", legend= c("Angiosperms", "Ferns", "Lycophytes"),
-       pch=21, inset=0.01, bty='n', cex=1,pt.cex=1.25, pt.bg=pgcols)
+       pch=21, inset=0.01, bty='n', cex=1,pt.cex=1.5, pt.bg=pgcols)
 
 # dev.copy2pdf(file= "output/pca_shade1.pdf")
  # dev.off()
@@ -124,4 +124,4 @@ legend("topleft", legend= c("Angiosperms", "Ferns", "Lycophytes"),
 #        col=c(rep("black",3),pgcols[1],"black"))
 # 
 # dev.copy2pdf(file= "output/pca_shade2.pdf")
-# dev.off()
+dev.off()

### 2021-09-27 manuscript/manuscript.Rmd
ran lettuce here, need to create new project!

@@
 Rates of g~s~ for angiosperms in open habitats increased nearly two-fold compared to angiosperms in closed habitats (*P* = 0.0341, t=-2.973), while rates of g~s~ for ferns were statistically similar across habitat types (Figure `r interfig`B).
-Neither ferns nor angiosperms differed in WUE across the open and closed habitats, and WUE was  statistically similar bewteen the two plant groups.
+Neither ferns nor angiosperms differed in WUE across the open and closed habitats, and WUE was  statistically similar between the two plant groups.
 
@@
 In closed canopy habitats, the LCP of spore-bearing plant species were equivalent and 161.9 % higher than angiosperm species (*P* = 0.0195, Figure `r seedsporefig`C).
-Photosynthetic quantum was statistically similar between each plant group, but lycophytes and ferns were broadly higher than angiosperms.
+Photosynthetic quantum yield was statistically similar between each plant group, but lycophytes and ferns were broadly higher than angiosperms.
 Spore-bearing plant species had equivalent R~d~, and R~d~ was nearly 4-fold lower than in angiosperms (*P* = 0.0178, Figure `r seedsporefig`D)

### 2021-09-27 scripts/lettuce.R

@@ -1,0 +1,21 @@
+##light response
+
+lrc <- read.csv("raw_data/lightresponse_lettuce.csv")
+lrc$id <- paste(lrc$species, lrc$individual, sep="-")
+
+source("functions/light_response_function.R")
+
+lettuce_aqua_lrc <- fit_AQ_curve(lrc, 
+                          Photo = "Photo", 
+                          PARi = "PARi", 
+                          group_id = "id") 
+
+diagnostic_AQ_plot(lrc, lettuce_aqua_lrc,
+                   Photo = "Photo", 
+                   PARi = "PARi", 
+                   group_id = "id",
+                   save_to_pdf = TRUE,
+                   save_path = "output",
+                   file_name = "LRC_lettuce_plots")
+
+write.csv(lettuce_aqua_lrc, "calculated_data/AQ_params_lettuce.csv", row.names = FALSE)
